### 2013-10-15 doc/proposal/bios/benzi.tex

%%---------------------------------------------------------------------------%%
%% benzi.tex
%%---------------------------------------------------------------------------%%

\subsection{Biographical Sketch for Michele Benzi}

% This section is not called out in the RFP, so it could be omitted if
% you are tight on space.
%\vspace*{-1ex}
\subsubsection*{Address}
\begin{tabular}{l}
  Department of Mathematics and Computer Science \\
  Emory University \\
  400 Dowman Drive \\
  Atlanta, GA 30322 \\
  Email: {\tt benzi@mathcs.emory.edu}
\end{tabular}

%%---------------------------------------------------------------------------%%

\vspace*{-1ex}
\subsubsection*{Education and Training}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex}
  xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxx \kill
  University of Bologna, Italy \> Mathematics \>  B.S., 1987 \\
  North Carolina State University \> Mathematics \> M.S., 1991 \\
  North Carolina State University \> Applied Mathematics \> Ph.D., 1993
\end{tabbing}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Research and Professional Experience}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex} 
  %% 
  \= 2012 -- present \hspace*{2ex} \= Samuel Candler Dobbs Professor, Emory
  University\\
  %%
  \> 2006 -- 2012 \> Professor, Emory University\\ 
  %% 
  \> 2000 -- 2006 \> Associate Professor, Emory University \\
  %%
  \> 1998 -- 2000 \> Staff member, Scientific Computing Group, LANL \\ 
  %%
  \> 1997 -- 1998 \> Director-funded postdoctoral associate, Scientific 
  Computing Group, LANL\\
  %%
  \> 1996 -- 1997 \> Postdoctoral Researcher, Parallel Algorithms Group,
  CERFACS, Toulouse, France\\
  %%
  \> 1993 -- 1996 \>  Researcher, Department of Mathematics, University of Bologna,
  Italy
\end{tabbing}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Related Publications}
\begin{enumerate}
\parskip = -2pt

\item M.~Benzi and V.~Kuhlemann, ``Restricted additive Schwarz methods
  for Markov chains'', Numer.~Linear Algebra Appl., {\bf 18} (2011), 
  pp.~1011--1029.

\item M.~Benzi, M.~K.~Ng, W.~Niu and Z.~Wang, ``A relaxed dimensional
  factorization preconditioner for the Navier--Stokes equations'', 
  J.~Comp.~Phys., {\bf 230} (2011), pp.~6185--6202.

\item S.~P.~Hamilton, M.~Benzi and E.~Haber, ``New multigrid
  smoothers for the Oseen Problem'', Numer.~Linear Algebra Appl., {\bf 17}
  (2010), pp.~557--576.

\item S.~P.~Hamilton, M.~Benzi and J.~S.~Warsa, ``Negative-flux fixups in 
  discontinuous finite element $S_N$ transport'', 
  Proceedings of the International Conference on Mathematics, 
  Computational Methods \& Reactor Physics 2009 (M\&C 2009), 
  American Nuclear Society, Vol. 4 (2009), pp. 2529-2538, 
  Saratoga Springs, NY, 2009.

\item M.~Noskov, M.~Benzi, and M.~D.~Smooke, ``An implicit compact 
  scheme solver for two-dimensional multicomponent flows'', Computers 
  \& Fluids, {\bf 36} (2007), pp.~376--397.

\item M.~Benzi, G.~H.~Golub, and J.~Liesen, ``Numerical solution of 
  saddle point problems'', Acta Numerica, {\bf 14} (2005), pp.~1--137.

\item J.~Warsa, M.~Benzi, T.~Wareing and J.~Morel, ``Preconditioning a 
  mixed discontinuous finite element method for radiation diffusion'', 
  Numer.~Linear Algebra Appl., {\bf 11} (2004), pp.~795--811.

\item M.~Benzi, ``Preconditioning techniques for large linear systems:
  a survey'', J.~Comp.~Phys., {\bf 182} (2002), pp.~418--477.

\item M.~Benzi and M.~Tuma, 
  ``A parallel solver for large-scale Markov chains'',
  Appl.~Numer.~Math., {\bf 41} (2002), pp.~135--153.

\item M.~Benzi, J.~Marin and M.~Tuma, ``A two-level parallel preconditioner
  based on sparse approximate inverses'', in Iterative Methods in Scientific 
  Computation IV, D. R. Kincaid and A. C. Elster, eds., IMACS Series in 
  Computational and Applied Mathematics, Vol.~5, IMACS, New Brunswick, 
  NJ (1999), pp.~167--178.

\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Synergistic Activities}
\vspace{-1ex}
%
\begin{enumerate}
  \parskip = -2pt
\item Chair, SIAM Activity Group on Linear Algebra (2010-present); 
\item Member of the SIAM Council (2009-present);
\item SIAM Fellow (Class of 2012); 
\item Program Committeee Co-chair, 2012 SIAM Applied Linear Conference
  (Valencia, Spain, June 18-22, 2012);
\item Program Committeee Co-chair, SIAM Annual Meeting (Minneapolis, MN, July
  9-13, 2012).
\item Editor-in-Charge, SISC Copper Mountain Special Issue (2012-present).
\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Co-Authors and Collaborators (last 48 months)}
\vspace{-1ex}

{\parindent 0.2in \narrower 
E.~Agichtein (Emory), Z.-Z.~Bai (CAS, Beijing), 
P.~Boito (Limoges), M.~Challacombe (LANL), F.~Chen (Beijing),
E.~Estrada (Strathclyde), L.~Ferragut (Saragoza), 
X.-P.~Guo (Shanghai), E.~Haber (UBC, Canada)
S.~Hamilton (ORNL), N.~Hatano (Tokyo), M.~K.~Ng (Hong Kong), 
Q.~Niu (Zuhai, PRC), M.~Pennacchio (Pavia), M.~Olshanskii (Moscow), 
N.~Razouk (Emory), L.~Rebholz (Clemson), V.~Simoncini 
(Bologna), Y.~Wang (Emory), Z.~Wang (ORNL), Z.-Q.~Wang (Shanghai),
J.~Warsa (LANL).
}

%%\vspace*{1ex}
\subsubsection*{Graduate and Postdoctoral Advisors and Advisees}
% List all advisors and advisees for the past 5 years
\vspace*{-1ex}
%%{\parindent 0.2in \narrower 
Advisors:
\begin{enumerate}
\item Wayne Joubert (Postdoc, LANL)
\item Iain Duff (Postdoc, CERFACS)
\item Carl Meyer (PhD, NCSU)
\end{enumerate} 
\vspace*{-1ex}
Advisees: 
\begin{enumerate}
\item Paola Boito (Postdoc) 
\item Bora Ucar (Postdoc) 
\item Jia Liu (PhD)
\item Nader Razouk (PhD)
\item Lauren Taralli (PhD)
\item Steven Hamilton (PhD)
\item Zhen Wang (PhD)
\item Christine Klymko (PhD)
\item Verena Kuhlemann (PhD)
\item Cheng-yi Zhang (PhD)
\end{enumerate}

\pagebreak
\endinput

%%---------------------------------------------------------------------------%%
%% end of benzi.tex
%%---------------------------------------------------------------------------%%