### 2013-10-15 doc/proposal/bios/engelmann.tex

%%---------------------------------------------------------------------------%%
%% engelmann.tex
%%---------------------------------------------------------------------------%%

\subsection{Biographical sketch for Christian Engelmann}

\begin{tabular*}{\textwidth}
{@{}l@{\extracolsep{\fill}}r@{}}
Christian Engelmann                       &
(865) 574-3132                            \\
Computer Science Research Group           &
engelmannc@ornl.gov                       \\
Computer Science and Mathematics Division &
www.christian-engelmann.info              \\
Oak Ridge National Laboratory             &
P.O. Box 2008, Oak Ridge, TN 37831-6173   \\
\end{tabular*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Accomplishments
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Professional Accomplishments}
\vspace*{-1ex}

\begin{tabular*}{\textwidth}
{@{}r@{~}
    l@{~~~}
    r@{~}
    l@{~~~}
    r@{~}
    l@{}}
8  & Research grants (2 as Lead)      &
6  & Peer-reviewed journal articles   &
35 & Invited talks and seminars       \\
8  & Co-advised M.Sc. theses          &
33 & Peer-reviewed conference papers  &
70 & Committees at 34 conferences     \\
2  & Mentored summer faculty          &
27 & Peer-reviewed workshop papers    &
20 & Article \& book proposal reviews \\
10 & Direct reports over 7 years      &
8  & Peer-reviewed conference posters &
11 & Conference booth exhibitions     \\
   & Erd\"{o}s number of 5            &
   & 860+ Total publication citations &
   & H-index of 16 / G-index of 27    \\
\end{tabular*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Experience
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Professional Experience}
\vspace*{-1ex}

\begin{compactitem}[0000--Present :]
  \item[2009--Present\hfill :] \textbf{R\&D Staff,
                                       Oak Ridge National Laboratory }
                            \begin{compactitem}
                            \item HPC hard-/software co-design:
                                  performance/resilience/power modeling \&
                                  simulation
                            \item Soft-error injection for vulnerability
                                  analysis of scientific applications and
                                  CMOS logic
                            \item HPC resiliency system software for monitoring,
                                  fault prediction, and fault avoidance
                            \item HPC checkpoint storage virtualization and
                                  MPI-level redundancy for HPC applications
                            \item HPC hard-/software co-design:
                                  performance modeling \& simulation
                            \item Light-weight simulation of extreme scale HPC
                                  architectures ($\sim$100,000,000 cores)
                            \end{compactitem}
  \item[2004--2009\hfill :] \textbf{Associate R\&D Staff,
                                    Oak Ridge National Laboratory }
                            \begin{compactitem}
                            \item Fault tolerance support for MPI: Scalable
                                  membership, job pause, and process migration
                            \item 99.9997\% high availability for HPC
                                  head/service-node services: Torque and PVFS MDS
                            \item Ph.D. thesis research: Symmetric active/active
                                  high availability for HPC system services
                            \item Virtual system environments for
                                  ``plug-and-play'' supercomputing with HPC
                                  hypervisors
                            \item Enhancing application development productivity
                                  via a common view across platforms
                            \end{compactitem}
  \item[2001--2004\hfill :] \textbf{Post-Master's Research Associate,
                                    Oak Ridge National Laboratory }
                            \begin{compactitem}
                            \item Harness Distributed Virtual Machine: Pluggable, 
                                  lightweight, adaptive, and fault tolerant
                            \item Light-weight simulation of HPC architectures
                                  at large scale ($\sim$1,000,000 cores)
                            \end{compactitem}
  \item[2000--2001\hfill :] \textbf{Software Developer,
                                    Oak Ridge National Laboratory }
                            \begin{compactitem}
                            \item M.Sc. thesis research: Distributed peer-to-peer 
                                  control for Harness (a resilient runtime)
                            \end{compactitem}
\end{compactitem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Education
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Education}
\vspace*{-1ex}

\begin{compactitem}[0000]
  \item[2008] Ph.D. in Computer Science,
              University of Reading, UK
  \item[2001] M.Sc. in Computer Science,
              University of Reading, UK
  \item[2001] Dipl.-Ing. (FH) in Computer Systems Engineering,
              University of Applied Sciences Berlin, Germany
\end{compactitem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Important Peer-reviewed Publications
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Important Peer-reviewed Publications}
\vspace*{-1ex}

\begin{compactenum}[{[}1{]}]
\item D.~Fiala, F.~Mueller, C.~Engelmann, K.~Ferreira, R.~Brightwell, and R.~Riesen. Detection and correction of silent data corruption for large-scale high-performance computing. In \emph{Intl. Conf. on High Performance Computing, Networking, Storage and Analysis (SC)}, 2012.
\item J.~Elliott, K.~Kharbas, D.~Fiala, F.~Mueller, K.~Ferreira, and C.~Engelmann. Combining partial redundancy and checkpointing for HPC. In \emph{Intl. Conf. on Distributed Computing Systems (ICDCS)}, pages 615--626, 2012.
\item C.~Wang, F.~Mueller, C.~Engelmann, and S.~Scott. Proactive process-level live migration and back migration in HPC environments. \emph{J. of Parallel and Distributed Computing (JPDC)}, 72\penalty0 (2):\penalty0 254--267, 2012.
\item C.~Wang, S.~Vazhkudai, X.~Ma, F.~Meng, Y.~Kim, and C.~Engelmann. NVMalloc: Exposing an aggregate SSD store as a memory partition in extreme-scale machines. In \emph{Intl. Parallel and Distributed Processing Symp. (IPDPS)}, pages 957--968, 2012.
\item S.~B{\"o}hm and C.~Engelmann. xSim: The extreme-scale simulator. In \emph{Intl. Conf. on High Performance Computing and Simulation (HPCS)}, pages 280--286, 2011.
\item M.~Li, S.~Vazhkudai, A.~Butt, F.~Meng, X.~Ma, Y.~Kim, C.~Engelmann, and G.~Shipman. Functional partitioning to optimize end-to-end performance on many-core architectures. In \emph{ Intl. Conf. on High Performance Computing, Networking, Storage and Analysis (SC)}, pages 1--12, 2010.
\item X.~He, L.~Ou, C.~Engelmann, X.~Chen, and S.~Scott. Symmetric active/active metadata service for high availability parallel file systems. \emph{J. of Parallel and Distributed Computing (JPDC)}, 69\penalty0 (12):\penalty0 961--973, 2009.
\item A.~Nagarajan, F.~Mueller, C.~Engelmann, and S.~Scott. Proactive fault tolerance for HPC with Xen virtualization. In \emph{Intl. Conf. on Supercomputing (ICS)}, pages 23--32, 2007.
\item C.~Wang, F.~Mueller, C.~Engelmann, and S.~Scott. A job pause service under {LAM/MPI+BLCR} for transparent fault tolerance. In \emph{Intl. Parallel and Distributed Processing Symp. (IPDPS)}, pages 1--10, 2007.
\item J.~Varma, C.~Wang, F.~Mueller, C.~Engelmann, and S.~Scott. Scalable, fault-tolerant membership for MPI tasks on HPC systems. In \emph{Intl. Conf. on Supercomputing (ICS)}, pages 219--228, 2006.
\end{compactenum}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Important Professional Activities
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Important Professional Activities}
\vspace*{-1ex}

\begin{compactitem}[0000--0000 :]
  \item[2010--2012\hfill :]
    PC chair:
    Workshop on Latest Advances in Scalable Algorithms
    for Large-Scale Systems (ScalA) at the Intl.
    Conf. on High Perf. Computing, Networking,
    Storage and Analysis (SC)
  \item[2009--2012\hfill :]
    PC chair:
    Workshop on Resiliency in
    High-Perf. Computing (Resilience) at the
    Intl. Symp. on
    High Perf, Distributed Computing (HPDC), the
    Intl. Symp. on
    Cluster Computing and the Grid (CCGrid), and the
    European Conf. on
    Parallel and Distributed Computing (Euro-Par)
  \item[2006--2010\hfill :]
    PC chair:
    HPC Resiliency Summit at the
    Los Alamos CS Symp. (LACSS),
    and the
    High Availability and Perf. Workshop (HAPCW)
    at the Los Alamos CS Inst. (LACSI) Symp.
  \item[2012\hfill :]
    PC vice-chair:
    IEEE Intl. Symp. on Parallel and Distributed
    Processing and Apps. (ISPA)
  \item[2009\hfill :]
    PC vice-chair:
    IEEE Intl. Conf. on Networking, Architecture, and
    Storage (NAS)
\end{compactitem}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Professional Memberships
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Professional Memberships}
\vspace*{-1ex}

USENIX,
ACM + SIGHPC/SIGOPS,
IEEE + ComSoc/CS/RL,
IEEE CS TCFT/TCDP/TCPP/TCSC

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Collaborators and Co-editors
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Collaborators and Co-editors in Past 48 Months
          (Excl. Advisors, Advisees, Junior Pers., and Non-US)}
\vspace*{-1ex}

\textbf{ANL:}
  P. Beckman,
  A. Chien,
  R. Gupta,
  R. Ross,
  M. Snir;
\textbf{Cray:}
  L. Kaplan;
\textbf{DOE:}
  L. Nowell;
\textbf{Emory U:}
  M. Benzi;
\textbf{IBM:}
  M. Elnozahy;
\textbf{Intel:}
  S. Borkar;
\textbf{ISI:}
  B. Lucas;
\textbf{LA Tech:}
  C. Leangsuksun;
\textbf{LANL:}
  N. DeBardeleben;
\textbf{LBNL:}
  P. Hargrove,
  E. Roman,
  J. Wu;
\textbf{LLNL:}
  M. Schulz,
  B. Still;
\textbf{NCSU:}
  X. Ma,
  F. Mueller;
\textbf{ORNL:}
  S. Atchley,
  D. Bernholdt,
  D. Dillow,
  T. Evans,
  S. Hamilton,
  Y. Kim,
  T. Naughton,
  B. Park,
  G. Shipman,
  G. Vall\'ee,
  S. Vazhkudai;
\textbf{PNNL:}
  S. Krishnamoorthy,
  A. Vishnu;
\textbf{SNL:}
  A. Gentile,
  J. Brandt,
  R. Brightwell,
  R. Clay,
  K. Ferreira,
  M. Heroux;
\textbf{TN Tech:}
  S. Gafoor;
\textbf{UTK:}
  G. Bosilca,
  J. Dongarra;
\textbf{VA Tech:} 
  A. Butt

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Graduate and Postdoctoral Advisors
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Graduate and Postdoctoral Advisors}
\vspace*{-1ex}

Prof. V. Alexandrov, BSC, Spain;
A. Geist, ORNL;
Prof. S. Scott, TN Tech \& ORNL;
Prof. U. Metzler, U of Applied Sciences Berlin, Germany

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Graduate and Postdoctoral Advisees
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace*{-2ex}
\subsubsection*{Graduate and Postdoctoral Advisees}
\vspace*{-1ex}

R. Baumann, TI, Germany;
S. B{\"o}hm, ORNL;
I. Jones, Ocado, UK;
F. Lauer, UTK;
A. Litvinova, Gresham Computing, UK;
B. K{\"o}nning, TU of Berlin;
K. Uhlemann, Coca Cola, Germany;
M. Weber, TU of Dresden

\pagebreak
\endinput

%%---------------------------------------------------------------------------%%
%% end of hamilton.tex
%%---------------------------------------------------------------------------%%