### 2013-10-15 doc/proposal/bios/hamilton.tex

%%---------------------------------------------------------------------------%%
%% hamilton.tex
%%---------------------------------------------------------------------------%%

\subsection{Biographical sketch for Steven P. Hamilton}

\subsubsection*{Address}
\begin{tabular}{l}
  Oak Ridge National Laboratory \\
  Radiation Transport Group\\
  Reactor and Nuclear Systems Division\\
  One Bethel Valley Road\\
  P.O.~Box 2008, MS-6170\\
  Phone:  (865) 576-3535, Fax:    (865) 574-9619	\\
  Email:  \href{mailto:hamiltonsp@ornl.gov}{\texttt{hamiltonsp@ornl.gov}}
\end{tabular}

%%---------------------------------------------------------------------------%%

\vspace*{-1ex}
\subsubsection*{Education and Training}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex} 
  xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxxxxxxxxxxxxxxxxxx \= xxxxxxxxx \kill
  Georgia Institute of Technology \> Nuclear Engineering       \> B.S. 2006 \\
  Georgia Institute of Technology \> Nuclear Engineering       \> M.S. 2007 \\
  Emory University                \> Computational Mathematics \> PhD  2011
\end{tabbing}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Research and Professional Experience}
\vspace*{-1ex}
\begin{tabbing}
  \hspace*{1ex} 
  %%
  \= 2011 -- present \hspace*{2ex} \= R\&D Staff, Oak Ridge National Laboratory\\
  %%
  \> 2009 -- 2010 \> Instructor,Mathematics and Computer Science Department,
  Emory University\\
  %%
  \> 2008 \> Research Practicum, Computational Physics (CCS-2), Los Alamos
  National Laboratory\\
  %%
  \> 2006 \> Intern, Reactor Physics Group, Oak Ridge National Laboratory
\end{tabbing}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Related Publications}
\begin{enumerate}
\parskip = -2pt

\item S. Hamilton, K. Clarno, B. Philip, M. Berrill, R. Sampath and S. Allu.
   Integrated Radiation Transport and Nuclear Fuel Performance for
   Assembly-Level Simulations.  \textit{Proc. PHYSOR}, 2012.

\item S. Hamilton, K. Clarno, B. Philip, M. Berrill, R. Sampath and S. Allu.
   Coupled Radiation Transport and Thermomechanics using the AMP and Denovo
   Codes.  \textit{Proc. 11th Copper Mtn. Conference on Iterative Methods}, 
   2012.

\item S. Hamilton and K. Clarno.  Mathematical Framework for Coupling the AMP 
   and Denovo Codes.  \textit{Trans. Am. Nucl. Soc.}, {\bf 105}, 515--517, 2011.

\item S. Hamilton and M. Benzi.  A Davidson Method for the k-Eigenvalue Problem.
   \textit{Trans. Am. Nucl. Soc.}, {\bf 105}, 432--424, 2011.

\item S. Hamilton and M. Benzi.  Eigensolvers for Radiation Transport
   Applications.  \textit{Proc. 11th Copper Mountain Conference on 
   Iterative Methods}, 2010.

\item S. Hamilton, M. Benzi and J. Warsa.  Negative Flux Fixups in 
   Discontinuous Finite Element SN Transport.  \textit{Proc. International
   Conference on Mathematics, Computational Methods and Reactor Physics},
   2009.

\item S. Hamilton, M. Benzi and E. Haber.  New Multigrid Smoothers for the Oseen 
   Problem. Num. Lin. Alg. with Apps. {\bf 17}, 557--576, 2010.

\item S. Hamilton, M. Benzi and E. Haber.  New Smoothers for the Oseen 
   Problem. \textit{Proc. 14th Copper Mtn. Conference on Multigrid Methods},
   2009.

\item S. Hamilton, K. Clarno and C. de Oliveira.  Error Control in a 
   Time-Dependent Slice-Balance Method.  \textit{Trans. Am. Nucl. Soc.},
   {\bf 97}, 533--535, 2007.

\item W. Stacey et. al.  Advances in the Subcritical,
   Gas-Cooled Transmutation Reactor Concept.  Nucl. Tech. {\bf 159}, 2007.
\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Synergistic Activities}
\vspace{-1ex}
%
\begin{enumerate}
\item Developer, Denovo parallel $S_N$ transport code and the Advanced
  Multi-Physics (AMP) Nuclear Fuel Performance package at Oak Ridge National
  Laboratory.
\item Alumnus, DOE Computational Science Graduate Fellowship program.
\item Member, American Nuclear Society.
\end{enumerate}

%%---------------------------------------------------------------------------%%

\vspace*{-3ex}
\subsubsection*{Co-Authors and Collaborators (last 48 months)}
\vspace{-1ex}

{\bf LANL:} James Warsa
{\bf ORNL:} Srikanth Allu, Mark Baird, Kevin Clarno, Greg Davidson, Tom Evans,
Josh Jarrell, Wayne Joubert, Bobby Philip, Rahul Sampath,
{\bf Univ. Arizona}: Barry Ganapol,
{\bf Univ. British Columbia}: Eldad Haber,
{\bf Univ. New Mexico}: Cassiano de Oliveira,
{\bf Univ. Tennessee}: James Banfield

\vspace*{-2ex}
\subsubsection*{Graduate Advisor}
\vspace{-1ex}
Michele Benzi, {\bf Emory University}.

\pagebreak
\endinput

%%---------------------------------------------------------------------------%%
%% end of hamilton.tex
%%---------------------------------------------------------------------------%%