### 2013-10-15 doc/proposal/emory/budget.tex

\section*{Budget Justification }

%\bigskip

\noindent {\bf Salary}

One month summer salary in each year of the project
is requested to free
the PI from teaching in order to conduct a significant portion of
the proposed research. Fringe benefits for faculty members at Emory
are currently 30\% of salary. A 3\% raise per year is projected.

The proposed research provides an excellent opportunity 
for a motivated PhD student working on a dissertation in
large-scale scientific computing on emerging architectures. 
The PhD program in Computational 
Mathematics at Emory University has a strong history of attracting
top graduate students, many of whom went on to successful careers
in academia, government, and industry, including several females.
Funds to support a full-time (academic year) PhD student 
are requested for each of the three years of the project.
Student salary is \$19,500 for the first year (plus 
7.65\% FICA), with a 3\% raise per year.
Additionally, funds to cover the student's 
enrollment fees 
totalling \$2,500/year are being requested.

\medskip
\par\noindent
{\bf Travel}

Travel support to attend meetings with the other project
investigators plus one relevant conference per year,
for a total of \$4,000/year, is requested.

\end{document}