### 2013-10-15 doc/proposal/emory/support-benzi.tex

\subsubsection{Michele Benzi}

\subsubsubsection{Pending}
\begin{itemize}

\item \FundingWithTotal
{DOE ASCR}

{MCREX: Using Monte Carlo algorithms to achieve resiliency 
and performance at scale for linear and non-linear solver
applications (this proposal)}
{8.33}
{FY13--16}
{\$???}

\item \FundingWithTotal
{DOE ASCR}

{A Forest for the Trees: An Ecosystem of Generalized
N-Body Solvers}
{8.33}
{FY13--16}
{\$???}

\subsubsubsection{Current}
\begin{itemize}

\item \FundingWithTotal
{NSF}

{Numerical Linear Algebra Tools for the Analysis of Complex Networks}
{14}
{08/01/11--07/31/14}
{\$303,045}

\end{itemize}