### 2013-10-15 doc/proposal/facilities.tex

%%---------------------------------------------------------------------------%%
%% facilities.tex
%%---------------------------------------------------------------------------%%

\section{Facilities \& resources}

Due to ORNL's world class resources for computing, ORNL will be the prinicpal
provider of facilities and resources for this collaborative proposal.
However, facility descriptions are provided for each institution in the
proposal. 
%
\nopagebreak
\subsection{Oak Ridge National Laboratory}
\begin{center}
\includegraphics[scale=0.965]{supp/ORNL_fac1}
\end{center}

\includepdf[pages=-,pagecommand={\thispagestyle{fancy}},noautoscale=true,scale=0.99]
{supp/ORNL_Facility}

\subsection{Emory University}
\input{supp/emory}

\pagebreak
\endinput

%%---------------------------------------------------------------------------%%
%% end of facilities.tex
%%---------------------------------------------------------------------------%%