### 2013-10-15 doc/proposal/templates/support_template.tex

%%---------------------------------------------------------------------------%%
%% Pending and support for DOE Proposals
%%---------------------------------------------------------------------------%%

\documentclass[11pt]{article}

%%---------------------------------------------------------------------------%%

\pagestyle{empty}

\setlength{\textheight}{9.5in}
\setlength{\textwidth}{6.5in}
\setlength{\footskip}{0.5in}
\setlength{\headheight}{0.0in}
\setlength{\topmargin}{0.0in}
\setlength{\headsep}{0.0in}

\setlength{\oddsidemargin}{0in}
\setlength{\evensidemargin}{0in}

\setlength{\parindent}{0in}

%%---------------------------------------------------------------------------%%

\begin{document}

%%---------------------------------------------------------------------------%%

{\bf Thomas M. Evans}\\

%%---------------------------------------------------------------------------%%
{\bf Current Support}\\
%%---------------------------------------------------------------------------%%

Funding Organization: Department of Energy (NE)\\
Project Title: CASL Nuclear Modeling and Simulation Hub\\
Dates of the project: 6/1/10--6/1/15\\
Annual funding: \$25M\\
Level of effort: 0.66 FTE\\

%%---------------------------------------------------------------------------%%

Funding Organization: Oak Ridge National Laboratory (LDRD)\\
Project Title: Revolutionary Radiation Transport for Next-Generation
Predictive Multi-Physics Modeling and Simulation\\
Dates of the project:  10/1/09--10/1/11\\
Annual funding:  \$250K\\
Level of effort: 0.34 FTE\\

%%---------------------------------------------------------------------------%%
{\bf Pending Support}\\
%%---------------------------------------------------------------------------%%

Funding Organization:  U.S. Department of Energy, (ASCR/SciDAC)\\
Project Title: Isotope production toolkit (this proposal\\
Dates of the project: 10/1/11--9/30/16\\
Annual funding:  \$3M\\
Level of effort:  0.5 FTE\\

Funding Organization:  U.S. Department of Energy, (ASCR/SciDAC)\\
Project Title: Institute for Computational Design and Analysis at the extreme
scale (CODA) (SciDAC Institute)\\
Dates of the project: 10/1/11--9/30/16\\
Annual funding:  \$4.5M
Level of effort:  0.5 FTE\\

Funding Organization:  Oak Ridge National Laboratory (LDRD)\\
Project Title: Creating a virtual HFIR: A complete modeling and simulation
toolkit for isotope production in HFIR \\
Dates of the project: 10/1/11--10/1/13\\
Annual funding:  \$450K\\
Level of effort:  0.25 FTE\\

\end{document}

%%---------------------------------------------------------------------------%%
%% end of pending and support
%%---------------------------------------------------------------------------%%