### 2014-06-30 doc/mrb_mc/kmc_presentation/krylov_mc.tex
adding krylov monte carlo notebook

\documentclass{beamer}
\usetheme{Warsaw}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage{listings}
\usepackage{color}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm} 
\usepackage{amsmath} 
\usepackage{tmadd,tmath}
\usepackage[mathcal]{euscript} 
\usepackage{color}
\usepackage{textcomp}
\usepackage{algorithm,algorithmic}
\definecolor{listinggray}{gray}{0.9}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{
  backgroundcolor=\color{lbcolor},
  tabsize=4,
  rulecolor=,
  language=c++,
  basicstyle=\scriptsize,
  upquote=true,
  aboveskip={1.5\baselineskip},
  columns=fixed,
  showstringspaces=false,
  extendedchars=true,
  breaklines=true,
  prebreak =
  \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
  frame=single,
  showtabs=false,
  showspaces=false,
  showstringspaces=false,
  identifierstyle=\ttfamily,
  keywordstyle=\color[rgb]{0,0,1},
  commentstyle=\color[rgb]{0.133,0.545,0.133},
  stringstyle=\color[rgb]{0.627,0.126,0.941},
}

%% colors
\setbeamercolor{boxheadcolor}{fg=white,bg=black}
\setbeamercolor{boxbodycolor}{fg=black,bg=white}

%% slide numbers

%%---------------------------------------------------------------------------%%
\author[Stuart Slattery]{Stuart Slattery\\
  \bigskip
  \href{mailto:slatterysr@ornl.gov}{\texttt{slatterysr@ornl.gov}} \\
  \bigskip
  Oak Ridge National Laboratory}

\date{\today} 
\title[Krylov Monte Carlo \hspace{1mm}
  \insertframenumber/\inserttotalframenumber]{Stochastic Krylov
  Subspace Approximation}
\begin{document}
\maketitle

%%---------------------------------------------------------------------------%%
\begin{frame}{Projection Methods}

  \medskip \medskip
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Search Subspace $\mathcal{K}$} Extract the solution to
    $\ve{A}\ve{x}=\ve{b}$ from the search subspace:
    \[
    \tilde{\ve{x}} = \ve{x}_0 +
    \boldsymbol{\delta},\ \boldsymbol{\delta} \in \mathcal{K}
    \]
  \end{beamerboxesrounded}

  \medskip \medskip
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Constraint Subspace $\mathcal{L}$} Constrain the extraction with
    the constraint subspace by asserting orthogonality with the
    residual ($\ve{r} = \ve{b}-\ve{A}\ve{x}$):
    \[
    \langle \tilde{\ve{r}},\ve{w} \rangle = 0,\ \forall \ve{w} \in
    \mathcal{L}
    \]  
  \end{beamerboxesrounded}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{The Orthogonality Constraint}

  \[
  \tilde{\ve{r}} = \ve{r}_0 - \ve{A}\boldsymbol{\delta}
  \]

  \begin{figure}[htpb!]
    \begin{center}
      \scalebox{1.1}{ \input{orthogonal_residual.pdftex_t} }
    \end{center}
    \caption{\textbf{Orthogonality constraint of the new residual with
        respect to $\mathcal{L}$.}}
  \end{figure}

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Minimization Property}

    The residual of the system is \textit{minimized} with respect to
    the constraints
    \[
    ||\tilde{\ve{r}}||_2 \leq ||\ve{r}_0||_2,\ \forall \ve{r}_0 \in
    \mathbb{R}^N
    \]
  \end{beamerboxesrounded}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Polynomials}

  The Krylov subspace looks a lot like our Monte Carlo process which
  estimates successive matrix-vector products:
  \[
  \ve{x} = \sum_k \ve{H}^k\ve{b} = \ve{b} + \ve{H}{b} + \ve{H}^2{b} +
  \dots + \ve{H}^k{b}
  \]
  Better yet, we are just constructing a general matrix polynomial:
  \[
  \ve{x} = \sum_k \alpha_k \ve{v}_k
  \]
  with $\ve{v}_k = \ve{H}^k\ve{b}$ and all $\alpha_k=1$. Each step of
  the random walk estimates a component of the polynomial basis:
  \[
  \ve{b} \rightarrow \ve{H}{b} \rightarrow \ve{H}^2{b} \rightarrow
  \dots \rightarrow \ve{H}^k{b}
  \]

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Stochastic Krylov Subspace Approximation}

  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Krylov Subspace}

    \[
    \mathcal{K}_m(\ve{A},\ve{r}_0) = span\{\ve{r}_0, \ve{A}\ve{r}_0,
    \ve{A}^2\ve{r}_0, \dots, \ve{A}^{m-1}\ve{r}_0\}
    \]
    For GMRES: $\mathcal{L} = \ve{A} \mathcal{K}_m(\ve{A},\ve{r}_0)$
  \end{beamerboxesrounded}

  We typically extract a correction from this subspace:
  \[
  \boldsymbol{\delta} = \sum_k \alpha_k^m \ve{A}^k \ve{r}_0
  \]
  which is just a polynomial in $\ve{A}$. Use the Monte Carlo process
  to stochastically construct $\mathcal{K}_m$. Apply linearly
  constrained least squares to extract $\boldsymbol{\delta}$ via
  orthogonality:
  \[
  minimize\ ||\ve{b} - \ve{A}(\ve{x_0}+\boldsymbol{\delta})||_2
  \]
  such that:
  \[
  \sum_{b=1}^B \alpha_b = 1
  \]

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}[fragile]{The Algorithm}

  \begin{algorithm}[H]
    \caption{Krylov Method}
    \begin{algorithmic}[1]
      \WHILE{$||(\ve{r})||/||(\ve{b})|| > \epsilon$}
      \STATE $\ve{r} = \ve{b} - \ve{A}\ve{x}$
      \STATE $\ve{V}$ = monteCarloPolynomial($\ve{A}$,$\ve{r}$,$N_p$,$m$)
      \STATE $\boldsymbol{\alpha}$ = constrainedLeastSquaresSolve($\ve{A}$,$\ve{x}$,$\ve{b}$,$\ve{V}$)
      \STATE $\boldsymbol{\delta} = \sum_k \alpha_k^m \ve{v}_k$
      \STATE $\ve{x} = \ve{x} + \boldsymbol{\delta}$
      \ENDWHILE
    \end{algorithmic}
  \end{algorithm}

  \begin{itemize}
  \item Expect most of the work to happen in polynomial construction
    in order to get good statistics
  \item Finding $\boldsymbol{\alpha}$ should be cheap except for dot
    products
  \item Parallel Monte Carlo strategies including threading and domain
    decomposition apply to polynomial construction
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Deterministic Results}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Results}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Summary}

  \begin{itemize}
  \item Stochastic computation of Krylov subspace - replace sequence
    of matrix-vector multiplies with a random walk of fixed length
    ($\mathcal{K}$ is defined stochastically)
  \item Petrov-Galerkin condition applied (correction extraction from
    $\mathcal{L}$ is defined deterministically
  \item Works for $\rho(\ve{H}) > 1$!
  \item Left/right preconditioning is now tractable given row-wise
    access to preconditioning matrix entries
  \item If Krylov subspace is computed exactly there are some nice
    optimality properties
  \item The bigger the subspace the better the convergence behavior
    (to a point)
  \item Right now subspace approximation is slow to converge with only
    diagonal preconditioning - more histories gives better convergence
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%

\end{document}