### 2015-03-10 doc/siam_cse_2015/siam_cse_2015.tex
adding siam cse 2015 talk

%%---------------------------------------------------------------------------%%
%% SIAM CSE 2015 Presentation
%%---------------------------------------------------------------------------%%

\documentclass{beamer}
\usefonttheme[onlymath]{serif}
\usepackage{colortbl}
\usepackage{longtable,ltcaption,array}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm} \usepackage{amsmath} \usepackage{tmadd,tmath}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{caption}
\usepackage[labelformat=empty]{subcaption}
\usepackage{tabularx}

%%---------------------------------------------------------------------------%%
%% THEME SETUP

\usetheme{CambridgeUS}
\usecolortheme{spruce}

%%---------------------------------------------------------------------------%%
%% SETUP STUFF
%%---------------------------------------------------------------------------%%

\logo{\includegraphics[width=2cm]{new_logo}}

\setbeamercolor{item}{fg=MSUgreen}
\setbeamertemplate{headline}{}
\setbeamertemplate{navigation symbols}{}

\newlength{\DUtablewidth} % internal use in tables

%%---------------------------------------------------------------------------%%
%% MATH STUFF
%%---------------------------------------------------------------------------%%

\newcommand{\Pn}{\textit{P}$\negthinspace_N$}
\newcommand{\SPn}{\textit{SP}$\negthinspace_N$}

\newcommand{\vOmega}{\ensuremath{\ve{\Omega}}}
\newcommand{\hOmega}{\ensuremath{\hat{\ve{\Omega}}}}

\newcommand{\sigs}{\ensuremath{\sigma_{\text{s}}}}
\newcommand{\sigf}{\ensuremath{\sigma_{\text{f}}}}
\newcommand{\sigsm}{\ensuremath{\sigma_{\text{s}m}}}
\newcommand{\sigsn}{\ensuremath{\sigma_{\text{s}n}}}

\newcommand{\phig}{\ensuremath{\Phi}}
\newcommand{\Sigmag}{\ensuremath{\mathbf{\Sigma}}}
\newcommand{\Dg}{\ensuremath{\mathbb{D}}}
\newcommand{\Fg}{\ensuremath{\mathbb{F}}}
\newcommand{\Qg}{\ensuremath{\mathbb{Q}}}
\newcommand{\ug}{\ensuremath{\mathbb{U}}}
\newcommand{\Ag}{\ensuremath{\mathbb{A}}}
\newcommand{\Bg}{\ensuremath{\mathbb{B}}}
\newcommand{\Cg}{\ensuremath{\mathbb{C}}}
\newcommand{\Jg}{\ensuremath{\mathbb{J}}}
\newcommand{\sg}{\ensuremath{\mathbb{S}}}

\newcommand{\aphig}{\ensuremath{\Phi^{\dagger}}}
\newcommand{\aSigmag}{\ensuremath{\mathbf{\Sigma}^{\dagger}}}

\newcommand{\apsi}[1]{\ensuremath{\psi^{\dagger\,#1}}}
\newcommand{\aphi}[1]{\ensuremath{\phi^{\dagger\,#1}}}
\newcommand{\aq}[1]{\ensuremath{q^{\dagger\,#1}}}
\newcommand{\ak}{\ensuremath{k^{\dagger}}}

\newcommand{\normal}{\ensuremath{\hat{\ve{n}}}}

% Phantom minus sign for helping with alignment of matrices
\newcommand{\phmin}{\ensuremath{\phantom{-}}}

\DeclareMathOperator{\diag}{diag}

%%---------------------------------------------------------------------------%%
%% TITLE

\title[Parallel Monte Carlo Solvers]{Parallel Algorithms for Monte Carlo
  Linear Solvers}
\author[Slattery, Hamilton, Evans]{Stuart Slattery
  \\ Steven Hamilton \\ Tom Evans}
\date[3/17/15]{March 17, 2015}
\institute[ORNL]{\small Oak Ridge National Laboratory}
\titlegraphic{\includegraphics[width=1.5in]{ORNL_Balls.pdf}}

\defbeamertemplate{footline}{titlefoot}{
    \vspace{-0.5cm}
    \hspace*{0.25cm}
    \includegraphics[width=2cm]{doe_logo}
    \hfill
    \includegraphics[width=3.25cm]{WordMarkLeaf}
    \hspace*{0.5cm}
}

\setbeamertemplate{title page}
{
  \begin{tabular}{cr}
    \begin{minipage}{4.7cm}
      {\bf \Large\textcolor{MSUgreen}{\inserttitle}}\\

      \vspace{2\baselineskip}
      \insertauthor\\

      \insertinstitute\\

      \insertdate
    \end{minipage}
    &
    \begin{minipage}{5cm}
      \raggedright
      \vspace{-0.75cm}
      \includegraphics[height=9.8cm]{combined_background}
    \end{minipage}
  \end{tabular}
}

%%---------------------------------------------------------------------------%%
\begin{document}

%%---------------------------------------------------------------------------%%

{
\setbeamertemplate{logo}{}
\setbeamertemplate{footline}[titlefoot]
\begin{frame}
\titlepage
\end{frame}
}

%%---------------------------------------------------------------------------%%
\begin{frame}{Acknowledgements}

  This material is based upon work supported by the U.S. Department of
  Energy, Office of Science, Advanced Scientific Computing Research
  program.

  \vfill

  This research used resources of the Oak Ridge Leadership Computing
  Facility at the Oak Ridge National Laboratory, which is supported by the
  Office of Science of the U.S. Department of Energy under Contract No.
  DE-AC05-00OR22725.

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Motivation}
  \vfill
\begin{itemize}
  \item As we move towards exascale computing, the rate of errors is expected
  to increase dramatically
  \vfill
  \begin{itemize}
  \item The probability that a compute node will fail during the course
    of a large scale calculation may be near 1
  \end{itemize}
  \vfill
\item Algorithms need to not only have increased concurrency/scalability
  but have the ability to recover from hardware faults
  \begin{itemize}
  \item Lightweight machines
  \item Heterogeneous machines
  \item Both characterized by low power and high concurrency
  \end{itemize}
  \vfill
\end{itemize}
\vfill
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Towards Exascale Concurrency and Resiliency}
  \begin{itemize}
    \item Two basic strategies:
      \vfill
      \begin{enumerate}
        \item State with current ``state of the art'' methods and make
          incremental modifications to improve scalability and fault
          tolerance
          \begin{itemize}
            \item Many efforts are heading in this direction, attempting
              to find additional concurrency to exploit
          \end{itemize}
        \vfill
        \item Start with methods having natural scalability and resiliency
          aspects and work at improving performance (e.g. Monte Carlo)
          \begin{itemize}
            \item Soft failures a component of the tally variance
            \item Hard failures potentially mitigated by replication
            \item Concurrency enabled by several levels of parallelism
          \end{itemize}
      \end{enumerate}
  \end{itemize}
  \vfill
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Outline}

  \begin{itemize}
  \item Monte Carlo Linear Solvers
    \vfill
  \item Domain Decomposition and Replication
    \vfill
  \item Scaling Studies
    \vfill
  \item Matrix-Free and Stochastic Approximate Inverse Algorithms
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Monte Carlo Methods

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo for Linear Systems}
  \begin{itemize}
    \item Suppose we want to solve $\mathbf{Ax}=\mathbf{b}$
    \vfill
    \item If $\rho(\mathbf{I-A})<1$, we can write the solution using the
      Neumann series
      \begin{equation*}
        \mathbf{x} = \sum_{n=0}^{\infty} (\mathbf{I-A})^n \mathbf{b}
         = \sum_{n=0}^{\infty} \mathbf{H}^n \mathbf{b}
      \end{equation*}
      where $\mathbf{H} \equiv ( \mathbf{I-A} )$ is the Richardson
      iteration matrix 
      \vfill
    \item Build the Neumann series stochastically
  \end{itemize}

  \[
  x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}
  \]

  \begin{itemize}
  \item Define a sequence of state transitions
  \end{itemize}
  \vspace*{-0.1in}
  \[
  \nu = i \rightarrow i_1 \rightarrow \cdots \rightarrow i_{k-1}
  \rightarrow i_{k}
  \]

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Forward Monte Carlo}
\begin{itemize}
  \item Choose a row-stochastic matrix $\mathbf{P}$ and weight matrix
    $\mathbf{W}$ such that $\mathbf{H} = \mathbf{P} \circ \mathbf{W}$
  \item Typical choice (Monte Carlo Almost-Optimal):
    \begin{equation*}
      \mathbf{P}_{ij} = \frac{| \mathbf{H}_{ij}| }
      {\sum_{j=1}^{N} | \mathbf{H}_{ij} | }
    \end{equation*}
  \item To compute solution component $\mathbf{x}_i$:
    \begin{itemize}
      \item Start a history in state $i$ (with initial weight of 1)
      \item Transition to new state $j$ based probabilities determined by
        $\mathbf{P}_i$
      \item Modify history weight based on corresponding entry in
        $\mathbf{W}_{ij}$
      \item Add contribution to $\mathbf{x}_i$ based on current history weight
        and value of $\mathbf{b}_j$
    \end{itemize}
  \item A given random walk can only contribute to a single component of
    the solution vector
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Sampling Example (Forward Monte Carlo)}
  \begin{itemize}
    \item Suppose
  \begin{equation*}
    \mathbf{A} = \begin{bmatrix}
      \phmin 1.0 & -0.2 & -0.6 \\
      -0.4 & \phmin 1.0 & -0.4 \\
      -0.1 & -0.4 & \phmin 1.0 \end{bmatrix} \to
    \mathbf{H} \equiv (\mathbf{I-A}) = \begin{bmatrix}
       0.0 &  0.2 &  0.6 \\
       0.4 &  0.0 &  0.4 \\
       0.1 &  0.4 &  0.0 \end{bmatrix}
  \end{equation*}
    then
  \begin{equation*}
    \mathbf{P} = \begin{bmatrix}
       0.0 & 0.25 & 0.75 \\
       0.5 &  0.0 & 0.5 \\
       0.2 &  0.8 & 0.0 \end{bmatrix}, \quad
    \mathbf{W} = \begin{bmatrix}
       0.0 &  0.8 &  0.8 \\
       0.8 &  0.0 &  0.8 \\
       0.5 &  0.5 &  0.0 \end{bmatrix}
  \end{equation*}
    \vfill
    \item If a history is started in state $3$, there is a $20\%$ chance of
      it transitioning to state $1$ and an $80\%$ chance of moving to state
      $2$
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Solving the Heat Equation: Forward Method}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics<1>[width=4in]{direct_1.png}
      \includegraphics<2>[width=4in]{direct_10.png}
      \includegraphics<3>[width=4in]{direct_100.png}
      \includegraphics<4>[width=4in]{direct_1000.png}
    \end{center}
    \caption{
      \only<1>{\textbf{Forward solution.}
        \textit{\sn{2.5}{3} total histories.} }
      \only<2>{\textbf{Forward solution.}
        \textit{\sn{2.5}{4} total histories.} }
      \only<3>{\textbf{Forward solution.}
        \textit{\sn{2.5}{5} total histories.} }
      \only<4>{\textbf{Forward solution.}
        \textit{\sn{2.5}{6} total histories.} }
    }
  \end{figure}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Synthetic Acceleration}
  \begin{itemize}
  \item Devised by Evans and Mosher in the 2000's as an acceleration
    scheme for radiation diffusion problems (LANL)
    \vfill
  \item Can be abstracted as a general linear solver with Monte Carlo as a
    preconditioner, $M_{MC}$
    \vfill
  \item Combine with Richardson iteration as a ``smoother'' in between
    Monte Carlo steps:
    \begin{align*}
      \mathbf{r}^k &= \mathbf{b} - \mathbf{Ax}^k \\
      \mathbf{x}^{k+1/2} &= \mathbf{x}^k + \mathbf{r}^k \\
      \mathbf{r}^{k+1/2} &= \mathbf{b} - \mathbf{Ax}^{k+1/2} \\
      \mathbf{\delta} &= \mathbf{M_{MC}} \mathbf{r}^{k+1/2} \\
      \mathbf{x}^{k+1} &= \mathbf{x}^{k+1/2} + \delta
    \end{align*}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Domain Decomposition and Replication

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Domain Decomposed Monte Carlo}

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Each parallel process owns a piece of the domain (linear
        system)
        \bigskip
      \item Random walks must be transported between adjacent domains
        through parallel communication
        \bigskip
      \item Domain decomposition determined by the input system
        \bigskip
      \item Load balancing straightforward for the forward method
      \end{itemize}
    \end{column}

    \begin{column}{0.5\textwidth}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.75}{ \input{ddnu_example.pdftex_t} }
        \end{center}
        \caption{\small Domain decomposition example illustrating
          how domain-to-domain transport creates communication costs.}
      \end{figure}
    \end{column}
  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Asynchronous Monte Carlo Transport Kernel}

  \begin{columns}

    \begin{column}{0.5\textwidth}

      \small
      \begin{itemize}
      \item General extension of the Milagro algorithm (LANL)
      \item Asynchronous nearest neighbor communication of histories
      \item System-tunable communciation parameters of buffer size and check
        frequency (performance impact)
      \item Need an asynchronous strategy for exiting the transport loop
        without a collective
      \end{itemize}

      \vspace{-0.25in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{domain_to_domain.pdftex_t} }
        \end{center}
      \end{figure}

    \end{column}

    \begin{column}{0.5\textwidth}

      \vspace{-0.3in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.85}{ \input{transport_loop.pdftex_t} }
        \end{center}
      \end{figure}

    \end{column}

  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Stopping the Asynchronous Kernel}

  \vspace{-0.25in}

  \begin{figure}[htpb!]
    \begin{center}
      \scalebox{0.4}{ \input{master_comm_tree.pdftex_t} }
    \end{center}
    \caption{Master/Slave}
  \end{figure}

  \begin{columns}

    \begin{column}{0.5\textwidth}
      \vspace{-0.75in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.4}{ \input{binary_comm_tree.pdftex_t} }
        \end{center}
        \caption{Binary Tree}
      \end{figure}

    \end{column}

    \begin{column}{0.5\textwidth}

      \vspace{-0.5in}

      \begin{figure}[t!]
        \begin{center}
          \includegraphics[width=0.99\textwidth]{titan_weak_bvsm.pdf}
        \end{center}
        \caption{Weak scaling absolute efficiency}
        \label{fig:titan_weak_bvsm}
      \end{figure}

    \end{column}

  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Replication}

  Different batches of Monte Carlo samplescan be combined in summation via
  superposition if they have different random number streams. For two
  different batches:

  \[
  \mathbf{M_{MC}} \mathbf{x} =
  \frac{1}{2}(\mathbf{M_1}+\mathbf{M_2})\mathbf{x} 
  \]

  Consider each of these batches independent \textit{subsets} of a Monte Carlo
  operator where now the operator can be formed as a general additive
  decomposition of $N_S$ subsets:

  \[
  \mathbf{M_{MC}} = \frac{1}{N_S}\sum_{n=1}^{N_s}{\mathbf{M_n}}
  \]

  We replicate the linear problem and form each subset on a different group of
  parallel processes. Applying the subsets to a vector requires an AllReduce
  to form the sum. Each subset is domain decomposed.

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Scaling Studies

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallel Test Application -- Nuclear Reactor Analysis}
  \small
  The simplified $P_N$ ($SP_N$) equations are an approximation
  to the Boltzmann neutron transport equation used to simulate nuclear
  reactors

  \vspace*{-0.1in}

  \begin{multline}
    \hat{\Omega} \cdot \vec{\nabla} \psi(\vec{r},\hat{\Omega},E) +
    \sigma(\vec{r},E) \psi(\vec{r},\hat{\Omega},E) = \\ \iint
    \sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}' \cdot \hat{\Omega})
    \psi(\vec{r},\hat{\Omega}',E') d\Omega' dE' +
    q(\vec{r},\hat{\Omega},E)
  \end{multline}

  \begin{multline}
    -\nabla \cdot \Bigg[\frac{n}{2n+1}\frac{1}{\Sigma_{n-1}} \nabla
      \Big(\frac{n-1}{2n-1} \phi_{n-2} + \frac{n}{2n-1}\phi_n \Big) \\+
      \frac{n+1}{2n+1}\frac{1}{\Sigma_{n+1}} \nabla
      \Big(\frac{n+1}{2n+3}\phi_n + \frac{n+2}{2n+3}\phi_{n+2}\Big)
      \Bigg] \\+ \Sigma_n \phi_n = q \delta_{n0}\ \ \ \ \ \ \ \ \ n =
    0,2,4,\cdots,N
  \end{multline}

  \begin{equation}
  \hspace*{-0.75in}
    -\nabla \cdot \mathbb{D}_n \nabla \mathbb{U}_n + \sum_{m=1}^4
    \mathbb{A}_{nm} \mathbb{U}_m = \frac{1}{k} \sum_{m=1}^4
    \mathbb{F}_{nm} \mathbb{U}_m\ \ \ \ \ \ \ n = 1,2,3,4
  \end{equation}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{$SP_N$ Assembly Problem}
\small
  \begin{columns}
    \begin{column}{0.4\textwidth}
      Test problem -- $3 \times 3$ array of fuel assemblies with
      control rod in center location (Profugus)
      \begin{itemize}
      \item 23 energy groups, 2 angular moments, 25M degrees of freedom
      \item 1,000 computational cores via domain decomposition
      \item We are interested in solving generalized eigenvalue problem,
        for this study we use Arnoldi as the eigensolver and compare
        different methods for solving linear systems
      \end{itemize}
    \end{column}
    \begin{column}{0.6\textwidth}
      \vspace*{-0.5in}
      \begin{figure}
        \centering
        \includegraphics[width=3in]{prob4}
      \end{figure}
    \end{column}
  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallel $SP_N$ Results}
\begin{table}
\centering
\begin{tabular}{cccc}
\toprule
%Method & Total GMRES Iterations & Setup Time (s) & Solve Time (s) \\
\multirow{2}{*}{\bfseries Method} &
\bfseries Total Linear & \bfseries Setup & \bfseries Solve \\
& \bfseries Iteration & \bfseries Time (s) & \bfseries Time (s) \\
\midrule
GMRES-ILUT      & 1675 & 0.7  & 18.4 \\
GMRES-AMG       & 626  & 0.7  & 46.0 \\
GMRES-MGE       & 498  & 1.5  & 33.7 \\
Richardson-AINV & 5208 & 20.6 & 52.0 \\
MCSA-AINV       & 1268 & 25.5 & 46.6 \\
\bottomrule
\end{tabular}
\end{table}
\begin{itemize}
  \item ILUT preconditioning is winner here, but known to have issues
    with parallel scaling on large core counts
  \item Solve times for MCSA are competitive, but setup times are very large
    due to construction of sparse approximate inverse
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Matrix-Free and Stochastic Approximate Inverse Algorithms

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Alternative Parallelism -- Additive Schwarz}
  \begin{itemize}
    \item Instead of performing Monte Carlo on full problem, another
      possibility is to apply Monte Carlo as an additive Schwarz approach
    \vfill
    \item Decompose problem into (possibly overlapping) domains
    \vfill
    \item Perform Monte Carlo on individual subdomains
      \begin{itemize}
        \item No communication costs in Monte Carlo problem!
      \end{itemize}
    \vfill
    \item With domain decomposed Monte Carlo, iteration counts are effectively
      independent of the number of processors
    \vfill
    \item In an additive Schwarz approach, the preconditioner will become
      less effective as processor counts grow -- algorithmic scalability
      may be an issue
      \vfill
    \item \textbf{Replication for resiliency and performance}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{More Parallelism -- Threading}
  \begin{itemize}
    \item Within a Monte Carlo solve, every history is independent of other
      histories -- great potential for highly concurrent hardware
      (GPU, Xeon Phi)
    \vfill
    \item Polynomial formulation enables a priori determination of
      operation counts per thread
      \vfill
    \item Memory locality an issue due to random access via random
      walks (block formulation?)
      \vfill
    \item Early experiments using the Trilinos Kokkos library show promising
      performance for multi-core CPUs
    \vfill
    \item Team members recently took part in OLCF ``Hackathon'' in
      late October to begin implementing computation kernels in
      OpenACC to allow for GPU capability on Titan with early results
      indicating \textbf{1.3-9.4x} speedup for MCSA (largely dependent
      on random number generation)
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Conclusions}
  \begin{itemize}
  \item Monte Carlo methods offer great potential for both resilient and
    highly parallel solvers
    \vfill
  \item For certain classes of problems, Monte Carlo methods can be
    competitive with leading modern solvers
    \vfill
    \item Extending methods to broader problem areas is significant challenge
      and an attractive area for continued research
      \vfill
    \item Performance modeling and resiliency simulations this FY
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\end{document}
%%---------------------------------------------------------------------------%%

%%---------------------------------------------------------------------------%%
%% end of pres.tex
%%---------------------------------------------------------------------------%%