### 2015-04-13 doc/ned2015/response_to_reviewer.txt
updating ned paper for reviewer comments

Response to reviewer's comments:

------------------------------------------------------------------------------
1. Since the underlying analysis seems to be graphical, they should use as
many graphs to explain what is happening.  Following words and moving
information is not so easy for a reader.

We agree that without visual aids the discussion is complex and is very much
based on the underlying graph of the matrix. We have added a figure (Figure 1)
in section 3.1 and discussion the random walk in terms of the sparse graph
underneath the matrices that should help clarify the explanation, particularly
for the n_s and n_i parameters in our diffusion analysis. In addition, we have
expanded our description of the algorithm in section 2 and have provided more
definitions for clarity.

------------------------------------------------------------------------------
2. Abstract: The term "measured computational results" is not so clear at
first.  The authors may want to find another name for this quantity.

We agree and have clarified that leakage fractions computed via numerical
experiments show good agreement with the equations we have derived.

------------------------------------------------------------------------------
3. Page 2: Last sentence of the second paragraph, it is not clear what is
meant by "specifically, have poor algorithmic strong scaling."

This statement has been clarified to indicate that when an algorithm is strong
scaled, the amount of on-process work decreases while the amount of
communication increases due to increasing processor count and decreasing local
problem size.

------------------------------------------------------------------------------
4. Top of Page 3: "adjoint method Monte Carlo method", is it necessary to use
'method' twice?  Seems a bit awkward.

This is a typo and it has been corrected.

------------------------------------------------------------------------------
5. Should indicate after Eq(2) .. yet to be determined adjoint source…

We agree and have indicated that these quantities will be determined in the
derivation.

------------------------------------------------------------------------------
4.  I assume Wm is a weight vector.

Wm is a scalar quantity equal to the weight of a random walk on its m^th
step. We have clarified the definition of Wm by indicating this.

------------------------------------------------------------------------------
6. D is the diffusion FD operator.

We have clarified that D is the diffusion operator resulting from our choice
of finite difference scheme.

------------------------------------------------------------------------------
7. Why not diagonalize D when it first appears?

We agree that we could diagonalize D initially. However, for more difficult
problems diagonal preconditioning may not be sufficient to satisfy the
convergence criteria and therefore it is instructive to separate it in the
analysis to show how a preconditioner would affect both iterative performance
as well as indicate how our analysis could be extended for more complicated
preconditioners. In addition, for our model problem the preconditioner simply
results in a scaling of the operator spectrum computed in section 4.1 and
therefore lets us analyze the operator independently from the preconditioner.

------------------------------------------------------------------------------
8. I believe D can be written as a block tridiagonal matrix for the stencil
used.  Is this the case?  If not see Burlish and Stoerer.  This would simplify
the analysis greatly.

Although we agree other matrix orderings could have been used, the results of
the paper are fundamentally based on the local stencil and independent the
overall global structure of the matrix. A block tridiagonal ordering could
reduce the bandwidth of the matrix and improve computational efficiency for a
matrix-vector multiply. However, the Monte Carlo algorithm would perform
identically in both cases given the same parallel decomposition of the
problem. The addition of Figure 1 in response to the first comment should also
help with this part of the analysis.

For example, given a banded ordering like that used in this work or the
suggested block tridiagonal ordering, the probability distribution functions
would be of the same size and have the same nearest-neighbor connections
because both use the same local stencil. Therefore, both cases will yield
random walks of the same time and space complexity. In addition, if both cases
had the same domain decomposition, random walks would be communicated from
domain to domain in an identical matter because the same local stencil was
used, therefore also resulting in the same parallel analysis.

------------------------------------------------------------------------------
9. Finally, there are no flux results with estimated errors.  We need results
since, after all, they are the bottom-line.

The purpose of the paper is to provide analytic relationships for the
domain-to-domain communication of random walks, and therefore a measure of
potential bandwidth and latency issues, when the Neumann-Ulam algorithm is
subject to domain decomposition. The domain decomposition itself does not
perturb the results of the stochastic algorithm (e.g. the flux in this
case). Rather, it simply provides a means to perform random walks via the
rules of the Neumann-Ulam game where the data for the random walks (weights
and probablities) has been domain-decomposed.

Variance estimation, however, is a different story and domain decomposition
schemes to date trade inexact variance estimates for scalablity. This is due
to the fact that the order of events must be maintained to acheive a correct
variance estimate and fully asynchronous algorithms do not, in general, permit
this. The literature indicates that this order of operations is less
significant for the correctness of the flux estimate with differences in
computations induced by round-off error in floating point
computations. Furthermore, the accuracy and convergence properties of the
Monte Carlo method have been well documented in the references cited by the
paper.

Therefore, because the purpose of this work is to analyze communication costs
of the algorithm using an operator with spectral properties similar to large
scale neutronics problems and not computing the global flux and variance
estimate, we do not feel it is necessary in this paper to report flux results.

We will, however, in an upcoming paper present our full domain decomposition
algorithm for parallelizing the solver and give a thorough analysis of its
correctness and scalability.

------------------------------------------------------------------------------