### 2015-08-21 MAX_reports/MC_solvers_article/MC_solvers.tex

\documentclass[a4paper,10pt]{article}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{geometry}
\usepackage{subfig}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfig}
\usepackage{tabu}
\usepackage{multirow}
\usepackage{color}
\usepackage{listings}
\usepackage[]{algorithm2e}
\usepackage[autostyle,english=british]{csquotes}
\usepackage{minitoc}

\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{prop}{Proposition}

%opening
\title{Monte Carlo solvers for sparse linear systems}
\author{Michele Benzi, Massimiliano Lupo Pasini}
\date{}

\begin{document}

\maketitle

\begin{abstract}
We consider linear systems of equations characterized by a sparsity pattern, 
focusing in particular on problems derived from the discretization of partial 
differential equations. We propose a treatise of the current stochastic 
techniques present in literature that might be employed to solve the problems at 
hand. New theoretical results are introduced as sufficient criteria for the 
guarantee of the convergence of the methods analyzed. An analysis about viable 
preconditioners is accomplished as well. 
\end{abstract}

\section{Introduction}
For a long period the  
pursuit of a reliable solver for linear and nonlinear algebraic systems has 
been one of the most important issues in Numerical Analaysis. The 
mathematical properties of the model (e.g. likely sparsity pattern of the 
matrix) have urged scientists to look for efficient and pattern-driven 
solvers, capable to cope with curse of dimensionality and high scalability 
requests.

As scalability computing move towards exascale facilities (which stands for 
machines computing up to $10^{18}$ flops), new numerical schemes suitable for 
this 
kind of architectures are strongly demanded. In fact the purpose is to combine 
high 
fidelity of the results with an optimized use of hardware resources at hand.

Recently new algorithms that combine statistical and numerical techniques have 
been developed. Even though \textit{numerical efficiency} has been sacrificed 
sometimes, the advantages in terms of \textit{robustness} of these methods make 
it worth carrying on studies in this direction.
A class of methods that may be employed for this purpose is represented by 
Monte Carlo linear solvers.

These algorithms are the main topic covered in the work presented here, 
starting from what was developed so far by some of 
the authors who gave significant contributions (see \cite{Hal1962}, 
\cite{Hal1994}, 
\cite{DA1998}, \cite{DVA2001}, \cite{AADBTW2005},\cite{ESW2013} and 
\cite{EMSH2014}). \newline

As underlined in \cite{DA1998}, Monte Carlo methods may be split into two 
classes: \textit{direct methods} (\cite{DA1998}, \cite{DVA2001}) and 
\textit{iterative methods} (\cite{Hal1962}, 
\cite{Hal1994}, \cite{ESW2013} 
and \cite{EMSH2014}). The first are characterized by a merely stochastic 
scheme, 
therefore the provided error with respect to the exact solution is made of 
just a stochastic component. The second ones, instead, combine numerical and 
statistical schemes generating two types of error: a 
\textit{stochastic} one and a \textit{systematic} one. It does not 
mean 
that 
it will be always simple to recognize them separately. However it is important 
to 
be 
aware of this intrinsic structure, in order to target what is the part of the 
scheme that requires a refinement (e.g. increase of numerical iterations 
rather than random walks).\newline

The paper is organized as follows. In Section 2 we describe all the 
viable stochastic settings for the definition of the Markovian process. In 
Section 3 we reinterpret probabilistic facts in terms of spectral properties of 
the linear system to be solved. Section 4 is dedicated to necessary conditions 
and sufficient conditions for the convergence of the algorithms. In Section 5 
different adaptive approaches are presented, in order to truncate the random 
walks and to decide how many samplings to use. Section 6 regards the study 
of matrices for which the convergence of the stochastic schemes is facilitated. 
Then in Section 7 a hybrid scheme combining numerical and probabilistic 
techniques is described: the \textit{Monte Carlo Synthetic Acceleration}. 
Numerical results are presented in Section 8 and 9, while Section 10 is 
dedicated to 
conclusive remarks and goals for the future.

\section{Stochastic setting}

We start our treatise by introducing the direct methods.\newline

Assume to start from a linear system such as 
\begin{equation}
A \mathbf{x}=\mathbf{b},
\label{linsys}
\end{equation}
where $A\in \mathbb{R}^{n\times n}$, $\mathbf{x}$, $\mathbf{b} \in 
\mathbb{R}^n$.

By applying a left preconditioning, the system (\ref{linsys}) gets the 
following shape:
\begin{equation}
P^{-1}A \mathbf{x}=P^{-1}\mathbf{b},
\label{linsys_prec}
\end{equation}
(\ref{linsys_prec}) can be 
reinterpreted as a fixed point scheme
\begin{equation}
 \mathbf{x}=H\mathbf{x}+\mathbf{f}.
 \label{fixedpoint}
\end{equation}
where $H=I-P^{-1}A$ and $\mathbf{f}=P^{-1}\mathbf{b}$.

Assuming that the spectral radius $\rho(H)<1$, the solution to 
(\ref{fixedpoint}) can be written in terms of a power series of 
$H$ (Neumann series)
\[
\mathbf{x}=\sum_{i=0}^\infty H^i\mathbf{f}.
\]
Therefore the fixed point scheme generates a sequence of approximate solutions
$\{\mathbf{x}^{(k)}\}_{k=0}^{\infty}$ which converges to the exact solution 
regardless of the initial guess $\mathbf{x}_0$.

By restricting the attention to a single component of $\mathbf{x}$ we 
have
\begin{equation}
x_i=\sum_{\ell=0}^\infty \sum_{k_1}^n\sum_{k_2}^n\cdots \sum_{k_{\ell}=1}^n 
H_{k_0,k_1}H_{k_1,k_2}\cdots H_{k_{\ell-1}, k_{\ell}}f_{k_{\ell}}.
\label{forward}
\end{equation}
The last equation can be reinterpreted as the realization of an estimator 
defined on a random walk.

Let us start considering a random walk whose 
state space $S$ is characterized by the set of indices of the forcing term 
$\mathbf{f}$: 
\[
S=\{1,2,\cdots, n\} \subset \mathbb{N}.
\]
Each $i$-th step of the random walk has a random variable 
$k_i$ associated with it. The realization of $k_i$ represents the index of the 
component of $\mathbf{f}$ 
which is visited in the current step of the random walk.

The way the transition probability is built and the way the initial state of 
the random walk is chosen give birth to two different approaches we are 
going to use: the \textit{Forward} and \textit{Adjoint} methods.

\subsection{Forward Method}
The goal is to evaluate a functional such as 
\[
J(\mathbf{x})=(\mathbf{h},\mathbf{x})=\sum_{i=1}^n h_i x_i.
\]
where $\mathbf{h}\in \mathbb{R}^n$ is the Riesz representative in 
$\mathbb{R}^n$ of the functional $J$.
We can use it to build the initial probability $\tilde{p}: 
S\rightarrow [0,1]$ of the random walk such that
\[
\tilde{p}(k_0=i)=\tilde{p}_{k_0}=\frac{\lvert h_i\rvert}{\sum_{i=1}^n \lvert 
h_i\rvert}.
\]
It is important to stress out that the role of vector $\mathbf{h}$ is 
restricted to the construction of the initial probability. Once that is done, 
it is 
not used anymore in the definition of the stochastic process. 
A possible choice for the transition probability $P$ can be
\[
p(k_i=j \;\lvert\;k_{i-1}=i )=P_{i,j}=\frac{\lvert H_{i,j}\rvert}{\sum_{k=1}^n 
\lvert H_{i,k}\rvert}.
\]
where $\tilde{p}(\cdot,i):S\rightarrow [0,1]$ $\forall i\in S$.
A related sequence of random variables $w_{i,j}$ can be defined 
such that 
\[
w_{i,j}=\frac{H_{i,j}}{P_{i,j}}.
\]
The probability distribution of the random variables $w_{i,j}$ is represented 
by the transition matrix that rules the stochastic process. The $w_{i,j}$'s 
just introduced can be used to build one more sequence 
of random variables. 
At first we introduce quantities $W_j$ 
\[
W_{0}=\frac{h_{k_0}}{\tilde{p}_{k_0}}, \quad W_j=W_{j-1} w_{i,j}, \quad 
j=1,\cdots, i.
\]
By defining 
\[
X(\nu)=\sum_{m=0}^k W_m f_{i_m}
\]
as the random variable associated with a specific permutation $\nu$, we can 
define the estimator $\theta_i$ such as
\[
\theta=E[X]=\sum_{\nu}P_{\nu}X(\nu).
\]
The integer $n$ represents the size of the solution vector to (\ref{linsys}) 
and 
the
index 
$i$ 
is referred to the component of the solution vector we want to compute. 
$P_{\nu}$ is the probability associated with a specific permutation of the 
random walk.  
It can be proved that 
\[
E[W_i f_{k_i}]=(\mathbf{h},H^i\mathbf{f}), \quad i=0,1,2,\cdots
\]
and 
\[
\theta_i=E\bigg[\sum_{i=0}^\infty W_i f_{k_i}\bigg]=(\mathbf{h},\mathbf{x}).
\]

A possible choice for $\mathbf{h}$ is a vector of the standard basis. This 
would correspond in setting manually the initial state of the random walk, 
which turns the related initial probability into a Kronecker delta
\[
\tilde{p}(k_0=i)=\delta_{i,j}.
\]
By doing so, we have 
\begin{equation}
\theta_i=E\bigg[\sum_{\ell=0}^\infty W_{\ell} 
f_{k_{\ell}}\bigg]=x_i=\sum_{l=0}^\infty 
\sum_{k_1=1}^{n}\sum_{k_2=1}^n\cdots \sum_{k_{\ell}=1}^n 
P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1}, 
k_{\ell}}w_{k_0,k_1}w_{k_1,k_2}\cdots 
w_{k_{\ell-1}, k_{\ell}}f_{k_{\ell}}.
\label{dir_mean}
\end{equation}

As regards the variance, we remember that the following relation holds:
\begin{equation}
Var\bigg [\sum_{\ell=0}^\infty W_{\ell} 
f_{k_{\ell}}\bigg]=E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2 
f_{k_{\ell}}^2\bigg] - \bigg (E\bigg[\sum_{\ell=0}^\infty W_{\ell} 
f_{k_{\ell}}\bigg]\bigg )^2.
\label{dir_var}
\end{equation}

Hence the variance can be computed as the difference between the second 
moment of the random variable and the square of the first moment.\newline

In order to apply the Central Limit Theorem (CLT) to the estimators defined 
above, we must require that 
the estimators have both finite expected value and variance. This is 
equivalent in 
checking the finiteness of the expected value and of the second moment.
Therefore we have to impose the following conditions:

\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell} f_{k_{\ell}}\bigg]<\infty
\end{equation}

\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2 
f_{k_{\ell}}^2\bigg]<\infty
\end{equation}

The Forward method presented above, however, has the limitation of employing an 
entire set of permutations to estimate just a single entry of 
the solution at a time. Therefore, in order to estimate the entire solution 
vector for \ref{linsys_prec}, we have to employ as many sets of permutation as 
the number of entries composing the vector. Of course this might turn to be 
extremely expensive both in terms of time and resources for most of the 
challenging problems coming from applications. This is essentially the reason 
why usually it is preferred to adopt another stochastic algorithm which is 
explained here following: the \textit{Adjoint} Monte Carlo method.

\subsection{Adjoint Method}
A second Monte Carlo method can be derived by considering the linear system 
adjoint to (\ref{linsys_prec})
\begin{equation}
(P^{-1}A)^T\mathbf{y}=\mathbf{d},
\label{adj}
\end{equation}
where $\mathbf{y}$ and $\mathbf{d}$ are the adjoint solution and source term.

By reformulating the fixed point scheme, introducing initial probability, 
transition probability and weights 
sequences in the same fashion as done before, the expected value for the 
estimator becomes
\begin{equation}
\theta_j=E\bigg[\sum_{\ell=0}^\infty W_{\ell}\delta_{k_{\ell}, 
j}\bigg]=\sum_{\ell=0}^{\infty}\sum_{k_1}^n\sum_{k_2}^n\cdots\sum_{k_{\ell}}^n 
f_{k_0}P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1},K_{\ell}}w_{k_0,k_1}\cdots 
w_{k_{\ell-1},k_{\ell}}\delta_{k_{\ell},j}.
\label{adj_mean}
\end{equation}
This estimator is known in literature as \textit{collision} estimator.

We can notice now that the component of the source vector, in each 
permutation of the random walk, is associated with the initial step. The 
Kronecker delta 
at the end of the series stands for a filter. It means that if the goal is to 
estimate the $j$-th component of the solution vector, just a subset of all the 
permutations gives a contribution to this. The only permutations considered are 
those ones that currently reside on the index associated with the entry to be 
estimated.

The variance is

\begin{equation}
Var\bigg [\sum_{\ell=0}^\infty W_{\ell} 
f_{k_0}\delta_{k_{\ell},j}\bigg]=E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2 
f_{k_0}^2\delta_{k_{\ell},j}\bigg ] - \bigg (E\bigg[\sum_{\ell=0}^\infty 
W_{\ell} 
f_{k_0}\delta_{k_{\ell},j}\bigg]\bigg )^2\quad j=1,\cdots,n
\label{adj_var}.
\end{equation}

On the same lead of what done for the Forward method, there is need to impose 
finiteness of the expected value and second moment. Therefore the following 
conditions must be verified:

\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell}\delta_{k_{\ell}, 
j}\bigg]<\infty \quad j=1,\cdots,n
\end{equation}
and
\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2 
f_{k_0}^2\delta_{k_{\ell},j}\bigg]<\infty \quad j=1,\cdots,n.
\end{equation}

The main advantage of this method, compared to the Forward one, consists in the 
fact that a single set of permutation is used to estimate the entire solution 
vector. Therefore, in general, this algorithm has to be preferred to the 
previous one since it enables to reduce dramatically the computational burden.

In literature another estimator is employed along with the Adjoint Monte Carlo
method, the so called \textit{expected value} estimator. Its 
formulations is as follows:

\begin{equation}
\theta_j=E\bigg[f_j + \sum_{\ell=0}^\infty 
W_{\ell}H_{k_{\ell}, j}^T\bigg]=f_j 
+ \sum_{\ell=0}^{\infty}\sum_{k_1}^n\sum_{k_2} ^n\cdots\sum_ { k_ { \ell}}^n 
f_{k_0}P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1},K_{\ell}}w_{k_0,k_1}\cdots 
w_{k_{\ell-1},k_{\ell}}H_{k_{\ell},j}^T.
\label{adj_mean1}
\end{equation}

Differently from the collision estimator, the expected value estimator averages 
the deterministic contribution of the iteration matrix over all the potential 
states $j$ that might be reached from the current state $\ell$. The variance 
in this case becomes:

\begin{equation}
Var\bigg [\sum_{\ell=0}^\infty W_{\ell} 
f_{k_0}H_{k_{\ell},j}^T\bigg]=E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2 
f_{k_0}^2 {H_{k_{\ell},j}^T}^T\bigg ] - \bigg (E\bigg[\sum_{\ell=0}^\infty 
W_{\ell} 
f_{k_0}H_{k_{\ell},j}^T\bigg]\bigg )^2\quad j=1,\cdots,n
\label{adj_var}.
\end{equation}

In the sections dedicated to the numerical tests, both the collision estimator 
and the expected value are employed, in order to test their effectiveness.

\section{Numerical interpretation of the statistical requirements}
The requests imposed on the expected value and second moment of the estimators 
can be reformulated in a purely deterministic setting by looking at the 
spectral radii of some properly defined matrices.

For instance the condition of finiteness of the expected value can be 
reformulated by requiring 
\begin{equation}
 \rho(H)<1
 \label{rho_h},
\end{equation}

where $H$ is the iteration matrix of the fixed point scheme.
In fact both equations (\ref{dir_mean}) and (\ref{adj_mean}) can be considered 
as 
power series in terms of $H$ and the spectral radius of $H$ smaller 
than 1 is a necessary and sufficient condition for the Neumann series to 
converge.

By analogy we aim at reproducing a similar reasoning for the second moment of 
the estimator, since we want it to be finite as well. 
By looking at the equations of the variance (\ref{dir_var}) and (\ref{adj_var}) 
for 
the Forward and the Adjoint method respectively, we notice that the second 
moment can be reinterpreted as a power series with respect to the matrices
defined as follows:

\[
 \hat{H}_{i,j}=\frac{H_{i,j}^2}{P_{i,j}} \quad \text{- Forward Method}.
\]

or

\[
 \hat{H}_{i,j}=\frac{H_{j,i}^2}{P_{i,j}} \quad \text{- Adjoint Method}.
\]

Accordingly to the necessary and sufficient condition for the power series to 
converge, we require 
\begin{equation}
 \rho(\hat{H})<1.
 \label{rho_star}
\end{equation}

Condition (\ref{rho_h}) must be required for a generic fixed point scheme to 
reach convergence, whereas the extra condition (\ref{rho_star}) is typical of 
the stochastic schemes presented in this work. Having two conditions to 
verify makes the numerical treatment of the linear system via MC very 
difficult. We will discuss this in details later in the treatise. \newline

\section{Necessary conditions, sufficient conditions and choice of the 
transition matrix}

Here below some results presented in 
\cite{MASC2013} about necessary conditions and sufficient 
conditions for 
convergence are enunciated. In particular a suitable choice for constructing a 
transition probability is discussed to guarantee and speed-up convergence.

We remember that the construction of the transition probability must satisfy 
some constraints such as

\[
\begin{cases}
  P_{i,j}\ge 0 \\
 \sum_{j=1}^N P_{i,j}=1
\end{cases}
\]

plus one more restriction which varies depending on the particular Monte Carlo 
approach chosen

\begin{align*}
& \text{Forward Method:} \quad H_{i,j}\ne 0 \Rightarrow P_{i,j}\ne 0 \\ &
\text{Adjoint Method:} \quad H_{j,i}\ne 0 \Rightarrow P_{i,j}\ne 0 
\end{align*}.

The previous restrictions on the attainable values for the entries of the 
transition probability are called \textit{transition conditions}.

At first an introductory theoretical result is shown, necessary to motivate the
choices made afterwards for the choice of the transition probability.

\begin{thm}
 Consider a vector $a=(a_1,a_2,\cdots,a_N)^T$ where at least one element is 
nonzero, $a_k\ne0$ for some $k\in\{1,\cdots,N\}$.
\begin{itemize}
 \item For a probability vector $p=(p_1,p_2,\cdots,p_N)^T$ satisfying the 
transition conditions, the lower bound of $\displaystyle 
\sum_{k=0}^N\frac{a_k^2}{p_k}$ is $\bigg(\sum_{k=1}^N \lvert a_k\rvert\bigg)^2$
\item There always exists a probability vector such that $\displaystyle 
\sum_{k=0}^N\frac{a_k^2}{p_k}\ge c\ge 1$ for all $c>1$.
\end{itemize}
\label{lemma}
\end{thm}

It might be observed that the minimum of the quantity $\displaystyle 
\sum_{k=0}^N\frac{a_k^2}{p_k}$ is attained when the probability vector is 
defined such as $\displaystyle p_k=\frac{\lvert a_k\rvert}{\sum_{k=1}^N \lvert 
a_k\rvert}$. It implies that the $\infty$-norm of $\hat{H}$ is minimized as 
much 
as possible. \newline

For the sake of simplicity let us introduce now a generic random walk
truncated at a certain $k$-th step
\[
 \gamma_k:\; r_0\rightarrow r_1 \rightarrow r_2 \rightarrow \cdots \rightarrow 
r_k
\]
and the statistical estimator
\[
 X(\gamma_k)=\frac{H_{r_0,r_1}H_{r_1,r_2}\cdots 
H_{r_{k-1},r_k}}{P_{r_0,r_1}P_{r_1,r_2}\cdots P_{r_{k-1},r_k}}f_{r_k}
\]
which is associated with the Forward Method.

Then the following theorem holds
\begin{thm}\textit{(Suited to the Forward Method)}

Let $H\in \mathbb{R}^{n\times n}$ such that $\lVert H\rVert_{\infty}<1$. 
Consider $\nu_k$ as the realization of a random walk $\gamma$ truncated at the 
$k$-th step. Then, 
there always exists a 
transition matrix $P$ such that 
$Var\Big(X(\nu_k)\Big)\rightarrow 0$ and 
$Var\Big(\sum_{\nu}X(\nu_k)\Big)$ is bounded as $k\rightarrow \infty$.
\label{for_thm}
\end{thm}

If we introduce the estimator
\[
 X(\gamma_k)=\frac{H^T_{r_0,r_1}H^T_{r_1,r_2}\cdots 
H^T_{r_{k-1},r_k}}{P_{r_0,r_1}P_{r_1,r_2}\cdots 
P_{r_{k-1},r_k}}sign(f_{r_0})\lVert \mathbf{f}\rVert_1
\]
which is associated with the Adjoint Method, then we can state a 
theorem specular to \ref{for_thm}.

\begin{thm}\textit{(Suited to the Adjoint Method)}

 Let $H\in \mathbb{R}^{n\times n}$ with $\lVert H\rVert_{1}<1$. 
Consider $\nu_k$ as the realization of a random walk $\gamma$ truncated at the 
$k$-th step. Then, 
there always exists a 
transition matrix $P$ such that 
$Var\Big(X(\nu_k)\Big)\rightarrow 0$ and 
$Var\Big(\sum_{\nu}X(\nu_k)\Big)$ is bounded as $k\rightarrow \infty$.
\label{adj_thm}
\end{thm}

In particular, relying on what guaranteed by Theorem \ref{lemma}, in both 
Theorems \ref{for_thm} and \ref{adj_thm} the thesis holds by picking the MAO 
transition probability.

These theoretical results represent sufficient conditions for the convergence 
of the  
Forward and Adjoint Monte Carlo and they can be easily checked. 
However requiring $\lVert 
H\rVert_{\infty}<1$ or $\lVert H\rVert_1<1$ may be too demanding. 
Indeed, even if there exist some preconditioners able to guarantee this 
condition (e.g. approximate inverse preconditioners with a proper value of 
dropping 
tolerance), it may entail to compromise the sparsity of the 
iteration 
matrix.

Thusly, in situations where the hypotheses of Theorems \ref{for_thm} and 
\ref{adj_thm} are too demanding to be verified, the necessary and sufficient 
condition 
on the 
spectral radius of $\hat{H}$ is the only criterion to take as reference.
Nonetheless this constraint consists in 
computing the spectral radii of two matrices and the calculation 
might be very costly. Moreover the computation of the spectral radii is 
affected by resilience issues as well as 
a deterministic solver for the resolution of 
a linear system. Therefore the original difficulty is transferred without being 
actually tackled.

\section{Construction of the transition probability}
The way the transition probability is defined is extremely important. In fact 
in some circumstances it might determine whether the stochastic scheme 
converges or not. Two very simple ways of defining it are listed here 
below. These are the approaches adopted typically in literature.

\subsection*{Uniform probabilities}
With this approach transition matrix $P$ is such that all the non-zero elements 
in each row have equal probability of occurring. 
\[
\text{Forward}:\;P_{i,j}=
\begin{cases}
0 \quad \quad \quad \qquad \qquad \qquad \text{if}\quad H_{i,j}=0 \\
\frac{1}{\#(\text{non-zeros in the row)}} \quad \text{if} \quad H_{i,j}\ne 0
\end{cases}\quad 
\]
\[
\text{Adjoint}:\;P_{i,j}=
\begin{cases}
0 \quad \quad \quad \qquad \qquad \qquad \text{if}\quad H_{j,i}=0 \\
\frac{1}{\#(\text{non-zeros in the column)}} \quad \text{if} \quad H_{j,i}\ne 0
\end{cases}
\]
The Monte Carlo approach resorting to this definition of the transition matrix, 
in accordance to \cite{AADBTW2005}, is called \textit{Uniform Monte Carlo} (UM).

\subsection*{Weighted probabilities}
An alternative way of defining the transition matrix consists in attributing 
probabilities to all the entries of the transition probability accordingly 
to the magnitude of the elements. For instance we may employ the following 
definition
\[
\text{Forward}: \;
p(k_i=j \;\lvert\;k_{i-1}=i )=P_{i,j}=\frac{\lvert 
H_{i,j}\rvert^p}{\sum_{k=1}^n 
\lvert H_{i,k}\rvert^p}
\]

\[
\text{Adjoint}: \;
p(k_i=j \;\lvert\;k_{i-1}=i )=P_{i,j}=\frac{\lvert 
H_{j,i}\rvert^p}{\sum_{k=1}^n 
\lvert H_{k,i}\rvert^p}
\]

where $p\in \mathbb{N}$.
This way of proceeding, for $p=1$, still basing on the work \cite{AADBTW2005}, 
is called 
\textit{Monte Carlo Almost Optimal} (MAO).
The reason why the almost optimal probability is called this way is associated 
with a notice from Theorem \ref{lemma}. Indeed from 
the mentioned lemma we know that the quantity $\displaystyle 
\sum_{k=0}^N\frac{a_k^2}{p_k}$ is minimized when the probability vector is 
defined such as $\displaystyle p_k=\frac{\lvert a_k\rvert}{\sum_{k=1}^N \lvert 
a_k\rvert}$. This implies that the almost optimal probability minimizes the 
$\infty$-norm of $\hat{H}$ in the Forward method and the $1$-norm of $\hat{H}$ 
in the 
Adjoint method. This is very important to be pointed out since these two norms 
are so far the best practical upper bounds for $\rho(\hat{H})$ that we can 
easily 
handle. In the Section dedicated to the numerical results it will be proved 
that the Almost Optimal Probability eases the convergence of the MC solver, 
even in cases when the employment of the Uniform Probability does not provide 
convergence at all.

\subsection*{Examples of matrices for which $\rho(H)<1$ and $\rho(\hat{H})>1$}
By focusing on the iteration matrix related to the fixed point scheme, we 
provide some examples of matrices for which the traditional condition 
$\rho(H)<1$ is not mirrored by the equally necessary condition 
$\rho(\hat{H})<1$.
The Almost Optimal transition probability is employed in order to achieve 
optimal upper bounds for the spectral radius of $\hat{H}$. The iteration matrix 
is 
obtained by applying a left diagonal preconditioner to the original system 
\ref{linsys}.

The matrices we consider are taken from the Matrix Market repository. The first 
one is "JPWH\_991". As it can be noticed from Table \ref{table_nonconv}, the 
Forward MC guarantees convergence since both the the conditions on $H$ and 
$\hat{H}$ 
are verified. The same does not hold for the Adjoint method, since the condition 
on $\hat{H}$ does not hold anymore.

Another example provided is the matrix "FS\_680\_1". In this case both the 
Forward and the Adjoint methods do not converge, since the spectral radius of 
$\hat{H}$ is bigger than 1 for both the methods.

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{d.o.f.'s} &
$\rho(H)$&Forward $\rho(\hat{H})$&Adjoint $\rho(\hat{H})$\\
\hline
JPWH\_991 & 991 & $0.9797$ & $0.9797$ & $1.0505$\\ 
\hline
FS\_680\_1 & 680 & $0.9697$ & $1.2554$ & $3.5984$\\
\hline
\end{tabular}
\caption{Examples of matrices for which $\rho(\hat{H})<1$ is violated.} 
\label{table_nonconv}
\end{table}

\section{Adaptive approaches}

In formulas (\ref{dir_mean}) and (\ref{adj_mean}) the estimation of the 
solution to
the linear system (\ref{linsys_prec}) involves infinite sums. Of course this is 
not viable for 
computational issues, therefore it is necessary to resort to truncation of 
the sums through proper criteria. In details, the truncation implies the 
employment of a finite number of permutations and a finite number of 
steps for each one of the permutations computed. Therefore it is necessary to 
truncate at two different levels to decide both how many permutations to employ 
and how many steps each one of them should be kept running.

\subsection{An adaptive cut-off for the length of random walks}

At first we focus on the definition of a weight cut-off 
in order to decide where to terminate the histories. It may be reinterpreted as 
a quantity guiding us to decide how many terms of the Neumann 
series to consider. \newline

The goal is to apply the cutoff via an automatic procedure, without requiring 
the intervention of the user directly.
Therefore we are looking for an integer $m$ such that  
\[
\tilde{\theta}_i=E\bigg[\sum_{\ell=0}^m W_{\ell} 
f_{k_{\ell}}\bigg]=x_i=\sum_{\ell=0}^m 
\sum_{k_1=1}^{n}\sum_{k_2=1}^n\cdots \sum_{k_{\ell}=1}^n 
P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1}, 
k_{\ell}}w_{k_0,k_1}w_{k_1,k_2}\cdots 
w_{k_{\ell-1}, k_{\ell}}f_{k_{\ell}}
\]

and

\[
\tilde{\theta}_j=E\bigg[\sum_{\ell=0}^m W_{\ell}\delta_{k_{\ell}, 
j}\bigg]=\sum_{\ell=0}^{m}\sum_{k_1}^n\sum_{k_2}^n\cdots\sum_{k_{\ell}}^n 
f_{k_0}P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1},K_{\ell}}w_{k_0,k_1}\cdots 
w_{k_{\ell-1},k_{\ell}}\delta_{k_{\ell},j}
\]

are good approximations of (\ref{dir_mean}) and (\ref{adj_mean}) respectively.

In \cite{Slattery2013} there is a criterion for the cutoff of the random walk 
applicable to both the Forward and the Adjoint method. It requires the set up 
of a relative cutoff threshold $W_c$ and to look for a step $m$ such that
\begin{equation}
W_m \le W_c W_0.
\label{cutoff}
\end{equation}
$W_0$ is the value of the weight at the initial step of the random walk and 
$W_m$ is the value of the weight after $m$ steps.

\subsection{An a posteriori variance-based adaptivity}

We now focus on quantities that might lead to a proper 
number of permutations to compute, in order to get a prescribed order of 
accuracy.
As regards the Forward method, we know that the expression for the variance 
is defined in formula (\ref{dir_var}).

In this context, a reasonable criterion to determine the number $\tilde{N}_i$ 
of random walks to be run is setting a threshold $\varepsilon_1$ and determine
\begin{equation}
\tilde{N}_i \quad s.t.\quad \frac{\sqrt{Var[\theta_i]}}{\lvert 
E[\theta_i]\rvert}<\varepsilon_1, \quad i=1,\cdots,n.
\label{forward_adapt}
\end{equation}
The dependence of $Var[\theta_i]$ and $E[\theta_i]$ on $\tilde{N}_i$, which 
seems 
to be absent in the previous formula, is highlighted by the fact that 
$\theta_i$ 
is estimated by fixing a finite number of histories.
Therefore we are controlling the relative standard deviation requiring it not 
to be too large. In other words we are pursuing a statistical setting where the 
uncertainty factor is not dominating over the expected value.
This simple adaptive approach can be applied for the estimation of each 
component $x_i$. Hence different number of histories may be employed to 
compute different entries of the solution vector. \newline

As concerns the Adjoint method, the estimation of the variance for each entry 
is reproduced by Formula (\ref{adj_var}).

A possible adaptive selection of $\tilde{N}$, in this situation, is 
\begin{equation}
\tilde{N} \quad s.t. \quad \frac{\lVert 
\boldsymbol{\sigma}_{\tilde{N}}\rVert_1}{\lVert 
\mathbf{x}\rVert_1}<\varepsilon_1,
\label{adjoint_adapt}
\end{equation}
where $\boldsymbol{\sigma}$ is a vector whose entries are 
$\boldsymbol{\sigma}_{\tilde{N},i}=Var[\theta_i]$.

What introduced above can be exploited in order to build 
an a posteriori adaptive algorithm, capable to identify the minimal value of 
$\tilde{N}$ that verifies (\ref{forward_adapt}) or (\ref{adjoint_adapt}) 
respectively.

Below there are the pseudo-codes associated with both Forward and Adjoint Monte 
Carlo.

\begin{algorithm}[H]
 \KwData{$N$, $\varepsilon_1$}
 \KwResult{$\tilde{N}_i$, $\sigma_i$, $x_i$}
 For each entry of the solution vector (for $i=1,\cdots,n$) \;
 $\tilde{N}_i=N$\;
 compute $\theta_i$\;
 \While{$\frac{\sigma_i}{\lvert 
E[\theta_i]\rvert}<\varepsilon_1$}{
\vspace{0.1cm}
  $\tilde{N}=\tilde{N}+N$\;
  compute $E[\theta_i]=x_i$\;
 }
 return $\theta_i$, $x_i$, $\sigma_i$\; 
 \caption{A posteriori adaptive Forward Monte Carlo}
\end{algorithm}

\begin{algorithm}[H]
 \KwData{$N$, $\varepsilon_1$}
 \KwResult{$\tilde{N}$, $\boldsymbol{\sigma_i}$, $\mathbf{x}$}
 $\tilde{N}=N$\;
 compute $\theta$\;
 \While{$\frac{\lVert\boldsymbol{\sigma} \rVert}{\lVert 
E[\theta]\rVert}<\varepsilon_1$}{
\vspace{0.1cm}
  $\tilde{N}=\tilde{N}+N$\;
  compute $E[\theta]=\mathbf{x}$\;
 }
 return $\theta$, $\mathbf{x}$, $\boldsymbol{\sigma}$\; 
 \caption{A posteriori adaptive Adjoint Monte Carlo}
\end{algorithm}

\section{Sets of matrices for which the convergence of the MC linear solver is 
guaranteed}

As we have already seen, sufficient conditions for 
the convergence of the Monte Carlo linear solvers are very restrictive (see 
\cite{MASC2013}), whereas the necessary and sufficient condition requires the 
computation of 
$\rho(H)$and $\rho(\hat{H})$, which is very expensive in general. Once a 
proper 
preconditioner is picked (e.g. Approximate Inverse), $\rho(H)<1$ may 
be always guaranteed, whilst $\rho(\hat{H})<1$ is more 
problematic to be verified.

In order to bypass the cost of these checks, it is useful to pursue for sets of 
matrices for 
which the Monte Carlo linear solver always converges, maybe with the use of an 
ad-hoc preconditioner.

 \subsection{Strictly diagonally dominant matrices}
 One of the these sets is 
represented by strictly diagonally dominant matrices.

\begin{defn}
 A matrix $A\in\mathbb{R}^{n\times n}$is strictly diagonally dominant (s.d.d) 
by rows if
 \begin{equation}
    \lvert a_{ij}\rvert>\sum_{\substack{i=1\\ i\ne j}}^{i=n}\lvert 
a_{ij}\rvert
\label{sddr}
 \end{equation}
\end{defn}

\begin{defn}
 A matrix $A\in\mathbb{R}^{n\times n}$is strictly diagonally dominant (s.d.d) 
by columns if
 \begin{equation}
    \lvert a_{ij}\rvert>\sum_{\substack{j=1\\ j\ne i}}^{j=n}\lvert 
a_{ij}\rvert
\label{sddc}
 \end{equation}
\end{defn}

When (\ref{sddr}) holds we can resort to a left diagonal preconditioning and we 
get an iteration matrix $H=I-D^{-1}A$ such that $\lVert H \rVert_{\infty}<1$.

Introducing a MAO transition probability for the Forward Method we get
\[
 P_{ij}=\frac{\lvert H_{ij}\rvert}{\sum_{k=1}^n\lvert H_{ik}\rvert}
\]
therefore the entries of $\hat{H}$ are defined as follows
\[
 \hat{H}_{ij}=\frac{H^2_{ij}}{P_{ij}}=\lvert H_{ij}\rvert\bigg 
(\sum_{k=1}^n\lvert 
H_{ik}\rvert\bigg ).
\]
Therefore
\[
 \sum_{j=1}^n\lvert \hat{H}_{ij} \rvert = \sum_{j=1}^n \hat{H}_{ij} =\bigg 
(\sum_{j=1}^n \lvert H_{ij}\rvert\bigg )\bigg (\sum_{k=1}^n\lvert 
H_{ik}\rvert\bigg ) =\bigg 
(\sum_{j=1}^n \lvert H_{ij}\rvert\bigg ) ^2 <1 \quad \forall i=1,\cdots, n.
\]
This automatically implies that $\rho(\hat{H})<=\lVert 
\hat{H}\rVert_{\infty}<1$. Thus the Forward Monte Carlo converges.
However nothing a priori can be 
stated about the behavior of the Adjoint method.\newline

Instead if (\ref{sddc}) holds we can resort to a right diagonal preconditioning 
and we 
get an iteration matrix $H=I-AD^{-1}$ such that $\lVert H \rVert_{1}<1$.
In this case, by following a similar reasoning to the one made before, we 
conclude that the Adjoint Method convergences thanks to $\lVert 
\hat{H}\rVert_1<1$, 
whereas nothing can be said in advance as concerns the Forward method.

 \subsection{Generalized diagonally dominant matrices (GDDM)}

 Another set of matrices for which the convergence of 
the MC solvers is guaranteed is made of \textit{generalized diagonally 
dominant matrices}.

\begin{defn}
A square matrix $A\in\mathbb{R}^{n\times n}$ is said to be generalized 
diagonally dominant if 
\[
 \lvert a_{ii}\rvert x_i \ge \sum_{\substack{j=1\\j\ne i}}^n \lvert 
a_{ij}\rvert 
x_j, \quad i=1,\cdots,n
\]
for some positive vector $\mathbf{x}=(x_1,\cdots,x_n)^T$.
\end{defn}

A proper subset of the generalized diagonally dominant matrices is represented 
by $M$- matrices (see \cite{Ax1996}).
There are many definitions of $M$-matrices that may be found on books and all 
of them are equivalent. The one presented in \cite{Saad} states the 
following

\begin{defn}
A matrix $A\in\mathbb{R}^{n\times n}$ is said to be an $M$-matrix if it 
satisfies the following four properties:
\begin{itemize}
 \item $a_{i,i}>0$ for $i=1,\cdots,n$
 \item $a_{i,j}\le 0$ for $i\ne j$, $i,j=1,\cdots,n$
 \item $A$ is nonsingular
 \item $A^{-1}\ge 0$
\end{itemize}
\end{defn}

A very important result about $M$-matrices enables to come up with a convergent 
splitting and, therefore, a convergent fixed point scheme.

In particular Jacobi, Block Jacobi and Gauss Seidel  are 
convergent 
splitting for an $M$-matrix.
However, the construction of a splitting like this is not enough to ensure the 
convergence of 
the Monte Carlo linear solver. In fact none of these conditions is sufficient 
to guarantee $\rho(\hat{H})<1$. Because of this obstacle, other options must 
be considered.\newline

For this purpose we 
appeal to a result in \cite{Li2002}. In this paper the author presents an 
algorithm to transform a generalized diagonally dominant matrix with 
nonzero diagonal entries into strictly diagonally dominant. The 
algorithm works for a generic complex matrix $A\in\mathbb{C}^{n\times n}$.

Here below we report the algorithm at hand\newline

\begin{algorithm}[H]
 For a given complex matrix $A$, $a_{ii}\ne 0$, $i=1,\cdots, n$;\newline
 \begin{enumerate}
  \item Compute $S_{i}=\sum_{\substack{j=1 \\ j\ne i}^{n}}\lvert 
a_{ij}\rvert$, 
$i=1,2,\cdots,n$
\item Set $t=0$. For $i=1,2,\cdots, n$, if $\lvert a_{ii}\rvert>S_i$, then set 
$t=t+1$
\item If $t=0$, then print "A is not a GDDM": END
\item If $t=n$, then print "A is a GDDM": END
\item \For{i=1,n}{$d_{i}=\frac{S_i+\varepsilon}{\lvert 
a_{ii}\rvert+\varepsilon} \quad 
\varepsilon>0,\quad j=1,2,\cdots,n$\;
$a_{ji}=a_{ji}\cdot d_i$}
\item Go to step 1.
 \end{enumerate}
 \caption{Algorithm to turn a GDDM matrix into a s.d.d by rows.}
\end{algorithm}

This approach turns a generalized diagonally dominant matrix into a 
strictly diagonally dominant matrix by rows. By substituting 
$\displaystyle S_{i}=\sum^{n}_{\substack{j=1 \\ j\ne i}}\lvert a_{ij}\rvert$ 
at step 1 with 
$\displaystyle S_{i}=\sum^{n}_{\substack{i=1 \\ i\ne j}}\lvert a_{ij}\rvert$ 
and by replacing 
$a_{ji}=a_{ji}\cdot d_i$ with $a_{ji}=a_{ji}\cdot d_j$ we obtain the algorithm 
that turns a GDDM matrix into a s.d.d. by columns.\newline

Once we have applied this transformation to the matrix at hand into a s.d.d., 
we can use the Monte Carlo linear solver which is ensured to converge.

 \subsection{"Block-diagonally dominant"-like matrices}
 In this section we wonder in which situations a block diagonal preconditioning 
might come up with a convergent Monte Carlo linear solver. 
By mimicking the computations we already showed previously, the iteration 
matrix $H\in\mathbb{R}^{n\times n}$ resulting from a block diagonal 
preconditioning takes the form

\[
 H=I-D^{-1}A=\begin{bmatrix}0_{n_1\times n_1} & -[A_{11}]^{-1}A_{12} & \cdots &
\cdots & -[A_{11}]^{-1}A_{1p} \\
-[A_{22}]^{-1}A_{21} & 0_{n_2\times n_2} & -[A_{22}]^{-1}A_{23} & 
\cdots & -[A_{22}]^{-1}A_{2p}\\
\vdots & \vdots & \ddots & \vdots & \vdots\\
\vdots & \vdots & \vdots &\ddots & \vdots \\
-[A_{pp}]^{-1}A_{p1} &  \cdots & \cdots& 
-[A_{pp}]^{-1}A_{p,p-1} & 0_{n_p \times n_p}
\end{bmatrix}.
\]

For the sake of clarity, we notify that the symbol "$\%$" used in the following 
subsections represents the modulus function between two integer numbers, while 
the floor function is represented with pretty standard symbol "$\lfloor 
\cdot \rfloor$".

\subsubsection*{Forward method}
By assuming that all $n_i$'s have the same value we may define
\[
 m=n_i=\text{size of a block}=\frac{n}{p}.
\]

The MAO transition probability matrix becomes:
\[
 P_{i,j}=\frac{\lvert H_{i,j} \rvert}{\sum_{k=1}^n\lvert H_{i,k} 
\rvert}=\frac{\bigg \lvert \bigg ([A_{\lfloor 
\frac{i}{m}\rfloor \lfloor] 
\frac{i}{m}\rfloor}]^{-1} A_{\lfloor \frac{i}{m}\rfloor \lfloor 
\frac{j}{m}\rfloor}\bigg )_{i\%m,j\%m}\bigg 
\rvert}{\sum_{\substack{k=1\\k\ne i}}^n\bigg \lvert \bigg 
([A_{\lfloor 
\frac{i}{m}\rfloor \lfloor 
\frac{i}{m}\rfloor}]^{-1} A_{\lfloor \frac{i}{m}\rfloor \lfloor 
\frac{k}{m}\rfloor}\bigg )_{i\%m,k\%m}\bigg \rvert}
\]

Consequently, the $\hat{H}$ matrix is defined such that
\[
\hat{H}_{i,j} = \lvert H_{i,j}\rvert\bigg(\sum_{k=1}^n\lvert H_{ik}\rvert\bigg)=
\bigg \lvert \bigg ([A_{\lfloor \frac{i}{m}\rfloor \lfloor 
\frac{i}{m}\rfloor}]^{-1} A_{\lfloor \frac{i}{m}\rfloor \lfloor 
\frac{j}{m}\rfloor}\bigg )_{i\%m,j\%m}\bigg 
\rvert
\sum_{\substack{k=1\\k\ne i}}^n\bigg \lvert \bigg ([A_{\lfloor 
\frac{i}{m}\rfloor \lfloor 
\frac{i}{m}\rfloor}]^{-1} A_{\lfloor \frac{i}{m}\rfloor \lfloor 
\frac{k}{m}\rfloor}\bigg )_{i\%m,k\%m}\bigg \rvert
\]

By computing the sum over a generic row of $\hat{H}$ we get:
\[
 \sum_{j=1}^n \lvert \hat{H}_{i,j}\rvert=\sum_{j=1}^n \hat{H}_{i,j} = 
 \bigg ( \sum_{\substack{j=1\\j\ne i}}^n\bigg \lvert \bigg ([A_{\lfloor 
\frac{i}{m}\rfloor \lfloor 
\frac{i}{m}\rfloor}]^{-1} A_{\lfloor \frac{i}{m}\rfloor \lfloor 
\frac{j}{m}\rfloor}\bigg )_{i\%m,j\%m}\bigg \rvert \bigg ) ^2.
\]

If we focus on the norm $\lVert \hat{H}\rVert_{\infty}$, then the following 
equivalence condition holds:
\[
 \lVert \hat{H}\rVert_{\infty}<1 \Leftrightarrow \sum_{\substack{j=1\\j\ne 
i}}^n\bigg \lvert \bigg ([A_{\lfloor 
\frac{i}{m}\rfloor \lfloor 
\frac{i}{m}\rfloor}]^{-1} A_{\lfloor \frac{i}{m}\rfloor \lfloor 
\frac{j}{m}\rfloor}\bigg )_{i\%m,j\%m}\bigg \rvert <1 \quad \forall 
i=1,\cdots,n
\]
 A sufficient condition for this to happen is
 \begin{equation}
  \sum_{\substack{j=1\\j\ne i}}^p \lVert [A_{ii}]^{-1}A_{ij}\rVert_{\infty}<1.
    \label{block_cs}\quad \forall i=1,\ldots,p.
 \end{equation}

By defining a matrix $\tilde{H}\in \mathbb{R}^{p\times p}$ such that
\[
 \tilde{H}=\begin{bmatrix}0 & \lVert [A_{11}]^{-1}A_{12}\rVert_{\infty} & 
\cdots &
\cdots & \lVert [A_{11}]^{-1}A_{1p}\rVert_{\infty} \\
\lVert [A_{22}]^{-1}A_{21}\rVert_{\infty} & 0 & \lVert 
[A_{22}]^{-1}A_{23}\rVert_{\infty} & 
\cdots & \lVert [A_{22}]^{-1}A_{2p}\rVert_{\infty} \\
\vdots & \vdots & \ddots & \vdots & \vdots\\
\vdots & \vdots & \vdots &\ddots & \vdots \\
\lVert [A_{pp}]^{-1}A_{p1}\rVert_{\infty} &  \cdots & \cdots& 
\lVert [A_{pp}]^{-1}A_{p,p-1}\rVert_{\infty} & 0
\end{bmatrix}
\]

We can use the $\tilde{H}$ matrix just defined in order to introduce a 
sufficient condition for the convergence of the Forward Monet Carlo method with 
a Block Diagonal preconditioning.

\begin{equation}
 \lVert \tilde{H} \rVert_{\infty}<1 \Rightarrow \lVert \hat{H} 
\rVert_{\infty}<1.
\end{equation}

\subsubsection*{Adjoint method} 
Analogously to the Forward method, if we define 
\[
(\hat{H})^T_{i,j} = \lvert H^T_{i,j}\rvert\bigg(\sum_{k=1}^n\lvert 
H^T_{ik}\rvert\bigg)
\]
we can formulate a sufficient condition for the convergence of the Adjoint 
Monte 
Carlo method by introducing a matrix $\tilde{H}$ which in this case is such that
\[
  \tilde{H}=\begin{bmatrix}0 & \lVert [A_{11}]^{-1}A_{12}\rVert_{1} & \cdots 
&
\cdots & \lVert [A_{11}]^{-1}A_{1p}\rVert_{1} \\
\lVert [A_{22}]^{-1}A_{21}\rVert_{1} & 0 & \lVert 
[A_{22}]^{-1}A_{23}\rVert_{1} & 
\cdots & \lVert [A_{22}]^{-1}A_{2p}\rVert_{1} \\
\vdots & \vdots & \ddots & \vdots & \vdots\\
\vdots & \vdots & \vdots &\ddots & \vdots \\
\lVert [A_{pp}]^{-1}A_{p1}\rVert_{1} &  \cdots & \cdots& 
\lVert [A_{pp}]^{-1}A_{p,p-1}\rVert_{1} & 0
\end{bmatrix}.
\]
The sufficient condition assumes the form
\begin{equation}
 \lVert \tilde{H} \rVert_{1}<1 \Rightarrow \lVert \hat{H} \rVert_{1}<1.
\end{equation}

\section{Linear solvers with hybrid schemes}

Even if the standard Monte Carlo, in both its Forward and Adjoint formulation, 
has the advantage of being embarrassingly parallelizable (because of the 
independence of the samplings employed), it has a drawback as 
well. In fact its slow rate of convergence, associated with the 
$\frac{1}{\sqrt{N}}$ behavior predicted but the CLT, prevents its use for 
problems of actual interest in applied linear algebra (with many d.o.f's).

Moreover, when the spectral radius of the iteration matrix is close to the 
unit 
value, Monte Carlo direct methods reach convergence after an excessive number 
of steps for each 
permutation. In particular this aspect will be pointed out more by some 
numerical tests analyzed in the upcoming sections of this work.
This motivation makes the standard Mote Carlo paradigm even more 
unpractical for solving linear systems associated with realistic applications. 
\newline

To cope with this, in the early 2000's Evans et al. introduced a new method: 
the 
\textit{Monte Carlo Synthetic Acceleration (MCSA)} 
(\cite{ESW2013} and \cite{EMSH2014}). Earlier contributions on these direction 
are attributed also to other scientists (e.g. \textit{Sequential Method} by 
J.H. Halton). However the MCSA is the algorithm we are going to focus on more 
hereafter. If 
we think about a fixed point formulation of the problem
such as $\mathbf{x}=H\mathbf{x}+\mathbf{b}$, then the Monte Carlo Synthetic 
Acceleration assumes this form

\begin{algorithm}[H]
 \KwData{$H$, $\mathbf{b}$, $\mathbf{x}_0$}
 \KwResult{$x_{num}$}
 $\mathbf{x}^{l}=\mathbf{x}_0$\;
 \While{not reached convergence}{
  $\mathbf{x}^{l+\frac{1}{2}}=H\mathbf{x}^l+\mathbf{b}$\;
  $\mathbf{r}^{l+\frac{1}{2}}=\mathbf{b}-A\mathbf{x}^{l+\frac{1}{2}}$\;
  $\delta \mathbf{x}^{l+\frac{1}{2}}=(I-H)^{-1}\mathbf{r}^{l+\frac{1}{2}}$\;
  $\mathbf{x}^{l+1}=\mathbf{x}^{l+\frac{1}{2}}+\delta 
\mathbf{x}^{l+\frac{1}{2}}$\;
 }
 $x_{num}=x^{l+1}$\; 
 \caption{Monte Carlo Synthetic Acceleration}
\end{algorithm}
The Monte Carlo method is used to compute the updating contribution $\delta 
\mathbf{x}^{l+\frac{1}{2}}$.

The role of the first line of the iterative process has the goal to compress 
the residual of the previous iteration. This is the actual difference 
between the MCSA and the Sequential Monte Carlo. Once this done, the corrected 
residual 
is employed to compute an update $\delta x^{1+\frac{1}{2}}$ to the solution via 
one of the Monte Carlo techniques introduced before. The utility of compressing 
the residual at the very beginning of the iteration will be shown in the two 
last test cases for numerical validations.
The intent of MCSA is to designate most of the time employed by the 
computation to the accomplishment of the stochastic part of the algorithm. 
This 
is useful to decrease the probability of fault occurrence in the non-resilient 
steps, which is represented by the non-stochastic component of the numerical 
scheme presented above. In particular, the higher is the percentage of time 
spent on the computation of $\delta x^{1+\frac{1}{2}}$ via Monte Carlo, the 
better it should be in terms of resilience.

\subsection{Preconditioning techniques}
In this section we focus on the pursuit of preconditioners that facilitate the 
convergence of the stochastic schemes introduced so far. A priori, all of the 
following techniques are good candidates to improve the numerical setting. 
However issues about the fill-in effect and the verification of convergence 
conditions limit their employment for our purposes.

\subsubsection*{Block Preconditioning}
Consider a matrix $A\in\mathbb{R}^{n\times n}$ of the form
\[
 A=\begin{bmatrix}A_{11} & A_{12} & \cdots & A_{1p}\\
    A_{21} & A_{22} & \cdots & A_{2p}\\
    \vdots & \vdots & \ddots & \vdots \\
    A_{p1} & \cdots & \cdots & A_{pp}
   \end{bmatrix}
\]
where $p$ is a divisor of $n$ and such that $A_{ii} \in \mathbb{R}^{n_i 
\times n_i}$ are nonsingular square matrices.
By defining a block diagonal matrix such as 
\[
 D=\begin{bmatrix}A_{11} & 0_{12} & \cdots & 0_{1p}\\
    0_{21} & A_{22} & \cdots & 0_{2p}\\
    \vdots & \vdots & \ddots & \vdots \\
    0_{p1} & \cdots & \cdots & A_{pp}
   \end{bmatrix}
\]
it is possible to compute its inverse. 
$D^{-1}$  can be used as a preconditioner for the original matrix $A$.
In particular 
\[
 D^{-1}A=\begin{bmatrix}I_{n_1\times n_1} & A_{11}^{-1}A_{12} & \cdots &
\cdots & A_{11}^{-1}A_{1p} \\
A_{22}^{-1}A_{21} & I_{n_2\times n_2} & A_{22}^{-1}A_{23} & 
\cdots & A_{22}^{-1}A_{2p}\\
\vdots & \vdots & \ddots & \vdots & \vdots\\
\vdots & \vdots & \vdots &\ddots & \vdots \\
A_{pp}^{-1}A_{p1} &  \cdots & \cdots& 
A_{pp}^{-1}A_{p,p-1} & I_{n_p \times n_p}
\end{bmatrix}.
\]
The ultimate goal is to resort to this preconditioning technique in order to 
achieve the condition $\rho(D^{-1}A)<1$. In fact this is necessary in order to 
reformulate \ref{linsys} into \ref{fixedpoint}.

Convergence of \ref{fixedpoint} with block diagonal preconditioning, from the 
deterministic point of view, is guaranteed when $A$ is an irreducibly 
diagonally dominant matrix or when $A$ is an $M$-matrix. In fact in situation 
such as these, theoretical results guarantee that $A=D-N$ is a regular 
splitting (see \cite{Ax1996}). However the same theoretical results cannot be 
exploited to guarantee the convergence of the MC solvers, because of the 
additional check on $\rho(\hat{H})<1$ to be considered.

\subsubsection*{Approximate Inverse of a Sparse Matrix (AINV)}

This stimulated our interests into other types of preconditioners. At first we 
concentrated on Approximate Inverse preconditioners (see \cite{Saad} and 
\cite{Benzi2002}).

There are many practical techniques that may be applied in order to build the 
AINV preconditioner.
One of these is described in \cite{Tuma1996} and \cite{Tuma1998}.
This is the actual approach we applied to the construction of the 
preconditioner. The technique constructs the preconditioner via a 
bi-conjugation fashion.
The idea is to compute an approximate factorization of the form $W^T A Z=D$. 
$W$ and $Z$ are unit upper triangular matrices and $D$ is a diagonal matrix.

In order to compute the AINV preconditioner, a \texttt{FORTRAN} code 
provided by Miroslav Tuma (Academy of Sciences, Prague) is used. 
We report as follows the algorithm employed for the construction of the so 
called \textit{Right-looking Factored AINV}. The algorithm provide the user 
with a dropping tolerance $\tau$ which can be used in order to tune the fill-in 
effect 
of the factored preconditioner. \newline

\begin{algorithm}[H]
 \KwData{$p=q=(0,\cdots,0)\in \mathbb{R}^n$, $Z=[z_1,\cdots,z_n]=I_n$, 
$W=[w_1,\cdots,w_n]=I_n$}
 \KwResult{$Z$, $W$}
 \For{$k=1,\cdots, n$}{ 
 $p_k=w_k^T A e_k$, $q_k=e_k^T A z_k$ \;
     \For{$i=k+1,\cdots,n$}{
         $p_i=(w_k^T A e_k)/p_k$, $q_i=(e_k^T A z_i)/q_k$\;
         Apply a dropping rule to $p_i$, $q_i$\;
         $w_i=w_i - w_k p_i$, $z_i=z_i - z_k q_i$\\
         Apply a dropping rule to $w_{j,i}$ and $z_{j,i}$ for $j=1,\cdots,i$\;
     }\;
  }\;
 Choose diagonal entries of $D$ as the component of $p$ or $q$\;
 return $Z$, $W$\; 
 \caption{Right-looking Factored AINV}
\end{algorithm}

\subsubsection*{Incomplete LU Factorization Preconditioners (ILU)}
This set of preconditioners has the goal of computing a sparse lower triangular 
matrix $L$ and a sparse upper triangular matrix $U$ such that the residual 
matrix $R=LU-A$ respects some constraints. One of these constraints may be the 
fact that some entries must be set to zero.
For instance we can introduce a zero pattern set 
\[
 P\subset \{(i,j)\lvert i\ne j; 1\le i,j\le n\}
\]
and use this set in order to impose some constraints on the shape of the 
preconditioner built.\newline

Different variants of the 
Incomplete LU Factorization algorithm are available (see \cite{Saad}). 
However it is 
possible to prove that if all the formulation are well defined, then they are 
all equivalent. 
One of these variants is the following

\begin{algorithm}[H]
 \KwData{$P$}
 \KwResult{$A$}
 For each $(i,j)\in P$ set $a_{ij}=0$\;
 \For{$k=1,\cdot,n-1$}{
 \For{$i=k+1,n$}{
 \If{$(i,k)\notin P$}{
 $a_{ik}=\frac{a_{ik}}{a_{kk}}$
 \For{$j=k+1,\cdots, n$ and for $(i,j)\notin P$}{
  $a_{ij}=a_{ij}-a_{ik}*a_{kj}$
 }
 }
 }
 }
 return $Z$, $W$\; 
 \caption{General Static Pattern ILU}
\end{algorithm}

This kind of algorithm is guaranteed to terminate without any breakdown just 
for $M$-matrices.

Assuming that the zero pattern $P$ coincides with the zero pattern of $A$ leads 
to the $ILU(0)$. \newline

\begin{algorithm}[H]
 \KwData{$P$}
 \KwResult{$A$}
 \For{$i=2,\cdot,n$}{
 \For{$k=i,i-1$ and for $(i,k)\in NZ(A)$}{
  $a_{ik}=\frac{a_{ik}}{a_{kk}}$
  \For{$j=k+1,\cdots,n$ and for $(i,j)\in NZ(A)$}{
  $a_{ij}=a_{ij}-a_{ik}a_{kj}$}
 }
 }
 \caption{ILU(0).}
\end{algorithm}

More sophisticated variants of the algorithm enable to enlighten the constraint 
of the sparsity pattern, finding a compromise between sparsity and accuracy in 
approximating the inverse of $A$. We refer to \cite{Saad} and 
\cite{Benzi2002} for further information about it.

\section{Numerical results}

All the numerical results shown in the following sections are computed with 
\texttt{Matlab} on a standard laptop. The simulations are accomplished in a 
serial mode.
In the 
first subsection the standard Monte Carlo solver is employed, while in the 
subsequents ones results are obtained with the MCSA scheme.

\subsection{Check on the effectiveness of the adaptive criteria}
In this subsection we show some numerical tests to validate the utility of the 
adaptive approaches for the selection of the number of permutations to employ. 
For this purpose, we restrict our attention to standard Monte Carlo linear 
solvers.

At first we take into account the Forward Monte Carlo linear solver.
In particular, the adaptive threshold associated with formulas 
(\ref{forward_adapt}) and 
(\ref{adjoint_adapt}) is set to $\varepsilon_1=0.1$ and to 
$\varepsilon_1=0.01$. \newline

A set of matrices associated with problems of various nature has been collected.
All the problems treated have a small number of d.o.f's to control the 
massive computational cost of MC solvers on a standard laptop. All the analyzed 
matrices are derived from the 
discretization of other differential problems.
One of these matrices is produced by the finite volume discretization of 
a thermal equation (Marshak problem), another one is a 1D diffusion reaction 
problem discretized with finite differences, a 
2D laplacian discretized with finite differences and an advection 
diffusion 
problem discretized with quadrilateral linear finite elements (the 
\texttt{ifiss} package is used in this case). 
For all the problems at hand a left diagonal preconditioning is applied. 
Details about all these matrices are gathered in Table 
\ref{table_data}.

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{d.o.f.'s} &
$\rho(H)$&Forward $\rho(\hat{H})$&Adjoint $\rho(\hat{H})$\\
\hline
1d shifted Laplacian & 50 & $0.4991$ & $0.9827$ & $0.9827$\\ 
\hline
2d Laplacian & 900 & $0.9949$ & $0.994$ & $0.9945$\\
\hline
ifiss & 1089 &$0.9836$ & $0.983$ & $0.983$ \\
\hline
Marshak Equation & 1600 & $0.6009$ & $0.3758$ & $ 0.3815$ \\
\hline
\end{tabular}
\caption{Properties of the matrices employed for the checks on adaptivity.} 
\label{table_data}
\end{table}

For all the numerical tests results are presented both by using the Forward 
and the Adjoint Monte Carlo.

As concerns the Forward method, a safety threshold for maximal number of 
histories per 
entry is set to $10^7$. Results are presented for 
$\varepsilon_1=0.1$ and $\varepsilon_1=0.01$ in Tables \ref{tab:For_adapt} and 
\ref{tab:For_adapt2} respectively. A batch size of two is used at each 
adaptive check to verify the magnitude of the apparent relative standard 
deviation. As expected, results are aligned with the convergence rate predicted 
by the Central Limit Theorem. Indeed, decreasing of a factor of 10 the 
tolerance $\varepsilon_1$, the relative error decreases of the same order, 
employing in average one hundred times more histories.\newline

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $0.003$ & $5,220$\\
\hline 
2d Laplacian & $0.1051$ & $262,350$\\
\hline
ifiss advection diffusion & $0.022$  & $1,060,770$\\
\hline
Marshak equation & $0.0012$ & $1,456,000$\\
\hline
\end{tabular}
\caption{Forward Monte Carlo. Adaptive selection of histories. 
$\varepsilon_1=0.1$} 
\label{tab:For_adapt}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $4\cdot 10^{-4}$ & $512,094$\\
\hline 
2d Laplacian & $0.0032$ & $108,551,850$\\
\hline
ifiss advection diffusion & $0.0023$  & $105,476,650$\\
\hline
Marshak equation & $5.8 \cdot 10^{-4}$ & $144,018,700$\\
\hline
\end{tabular}
\caption{Forward Monte Carlo. Adaptive selection of histories. 
$\varepsilon_1=0.01$} 
\label{tab:For_adapt2}
\end{table}

As regards the Adjoint Monte Carlo, at each adaptive check the number of 
random walks employed is increased by ten.

A maximal number of histories equal to $10^10$ is employed as a safety 
threshold.

Here below Tables \ref{tab:Adj_adapt}, \ref{tab:Adj_adapt2} and 
\ref{tab:Adj_adapt3} show results 
for the different test cases 
considered using $\varepsilon_1=0.1$, $\varepsilon_1=0.01$ and 
$\varepsilon_1=10^{-3}$ respectively.
By comparing the two tables, it is noticed that a decrease of the threshold 
induces a reduction of the relative error of the same order of magnitude. 
Therefore this validates 
the effectiveness of the adaptive selection of histories with an error 
reduction goal. The total number of histories employed increases on average by 
two orders of magnitude and this aspect is in line too with what expected by 
theoretical results.

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $0.08$ & $1820$\\
\hline 
2d Laplacian & $0.136$ & $570$\\
\hline
ifiss advection diffusion & $0.08$  & $2400$\\
\hline
Marshak equation & $0.288$ & $880$\\
\hline
\end{tabular}
\caption{Adjoint Monte Carlo. Adaptive selection of histories. 
$\varepsilon_1=0.1$} 
\label{tab:Adj_adapt}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $0.0082$ & $185,400$\\
\hline 
2d Laplacian & $0.0122$ & $126,800$\\
\hline
ifiss advection diffusion & $0.0093$  & $219,293$\\
\hline
Marshak equation & $0.0105$ & $650,040$\\
\hline
\end{tabular}
\caption{Adjoint Monte Carlo. Adaptive selection of histories. 
$\varepsilon_1=0.01$} 
\label{tab:Adj_adapt2}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $9.56 \cdot 10^{-4}$ & 15,268,560 \\
\hline 
2d Laplacian & $0.001$ & $12,600,000$\\
\hline
ifiss advection diffusion & $0.0011$  & $23,952,000$\\
\hline
Marshak equation & $0.0011$ &  $80,236,000$\\
\hline
\end{tabular}
\caption{Adjoint Monte Carlo. Adaptive selection of histories. 
$\varepsilon_1=10^{-3}$} 
\label{tab:Adj_adapt3}
\end{table}

The same results can be reproduced by resorting to the expected value 
estimator. Results are shown in the Tables \ref{tab:Adj_adapt4}, 
\ref{tab:Adj_adapt5} and \ref{tab:Adj_adapt6} for the threshold values of 
$\varepsilon_1=0.1$, $\varepsilon_1=0.01$ and $\varepsilon_1=10^{-3}$ 
respectively. As it can be noticed, in terms of error scaling the results are 
pretty similar to the ones obtained with the collision estimator. As regards 
the number of histories as well, used to reach a prescribed accuracy, the 
orders of magnitude are the same for both the collision estimator and the 
expected value one. However the expected value estimator requires most of the 
time a smaller number permutations. This behavior is enhances as the value of 
the threshold decreases, making the computation heavier and heavier.

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $0.0463$ & $1000$\\
\hline 
2d Laplacian & $0.1004$ & $900$\\
\hline
ifiss advection diffusion & $0.0661$  & $1300$\\
\hline
Marshak equation & $0.0526$ & $2000$\\
\hline
\end{tabular}
\caption{Expected value estimator - Adjoint Monte Carlo. Adaptive selection of 
histories. 
$\varepsilon_1=0.1$} 
\label{tab:Adj_adapt4}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $0.004$ & $100,600$\\
\hline 
2d Laplacian & $0.0094$ & $83,700$\\
\hline
ifiss advection diffusion & $0.0088$  & $124,400$\\
\hline
Marshak equation & $0.0056$ & $166,000$\\
\hline
\end{tabular}
\caption{Expected value - Adjoint Monte Carlo. Adaptive selection of histories. 
$\varepsilon_1=0.01$} 
\label{tab:Adj_adapt5}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|}
\hline
\textbf{Type of problem} & \textbf{Relative Error} &\textbf{Nb. Histories}\\
\hline
1d shifted Laplacian & $0.004$ & $10,063,300$\\
\hline 
2d Laplacian & $ 9.31 \cdot 10^{-4} $ & $ 8,377,500 $\\
\hline
ifiss advection diffusion & $ 0.0013 $  & $ 12,435,000 $\\
\hline
Marshak equation & $7.79 \cdot 10^{-4}$ & $16,537,000$\\
\hline
\end{tabular}
\caption{Expected value - Adjoint Monte Carlo. Adaptive selection of histories. 
$\varepsilon_1=10^{-3}$} 
\label{tab:Adj_adapt6}
\end{table}

\subsection{A comparison between different preconditioners and their effect 
on $\rho(H)$ and $\rho(\hat{H})$}

This subsection is focused on showing the effect of different preconditioners 
for the values attained by the spectral radii $\rho(H)$ and $\rho(\hat{H})$.
For this purpose, we consider the two dimensional laplacian 
discretized with centered finite differences and the "ifiss advection 
diffusion" matrix. The properties of both the matrices have already been 
explained in previous sections.
Results in terms of spectral radii when the diagonal preconditioning is applied 
have been shown earlier as well.

Here the analysis is focused on the evaluating the effects of the 
block diagonal preconditioner and the factorized approximate sparse inverse. 
As concerns the matrix $\hat{H}$, it has been built accordingly to te 
procedure for the Adjoint Monte Carlo method. In 
the first case the size of the blocks is a tuning parameter, whilst in the 
second case the tuning parameter is the drop tolerance to manage the fill-in 
effect of the factors.

As regards the employment for the block diagonal preconditiner, in Tables 
\ref{tab:lap2d_block} and \ref{tab:ifiss_block} there is a counterintuitive 
common trend. Indeed in both the cases an increase of $\rho(H)$ and 
$\rho(\hat{H})$  occurs as the size of the blocks diminishes.

In Tables \ref{tab:lap2d_ainv} and \ref{tab:ifiss_ainv} the values of 
the spectral radii are shown for the case of the factorized AINV 
preconditioner. It is interesting to point out that for the 
two-dimensional 
laplacian, a drop tolerance $\tau=0.05$ entails $\rho(H)<1$ but the same does 
not hold for $\hat{H}$. Both the convergence conditions are restored by 
decreasing even more the drop tolerance, causing a massive fill-in effect 
though.

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Block size} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$ 
&$\rho(\hat{H})$\\
\hline
 5 & 2.6164 & $0.9915$ & $0.9837$\\
\hline 
 10 & 5.6027 & $0.9907$ & $0.9861$\\
\hline
 30 & 16.3356 & $0.9898$  & $0.9886$\\
\hline
\end{tabular}
\caption{Behavior of $\rho(H)$ and $\rho(\hat{H})$ for the two dimensional 
laplacian with block diagonal preconditioning.} 
\label{tab:lap2d_block}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Block size} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$ 
&$\rho(\hat{H})$\\
\hline
 3 & 1.4352 & $0.9804$ & $0.9531$\\
\hline 
 9 & 3.0220 & $0.9790$ & $0.9674$\\
\hline
 33 & 9.8164 & $0.9783$  & $0.9774$\\
\hline
\end{tabular}
\caption{Behavior of $\rho(H)$ and $\rho(\hat{H})$ for the ifiss advection 
diffusion matrix with block diagonal preconditioning.} 
\label{tab:ifiss_block}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Drop tolerance} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$ 
&$\rho(\hat{H})$\\
\hline
 0.05 & 8.2797 & $0.9610$ & $1.0231$\\
\hline 
 0.01 & 33.1390 & $0.8668$ & $0.8279$\\
\hline
\end{tabular}
\caption{Behavior of $\rho(H)$ and $\rho(\hat{H})$ for the two dimensional 
laplacian with factorized AINV preconditioning.} 
\label{tab:lap2d_ainv}
\end{table}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Drop tolerance} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$ 
&$\rho(\hat{H})$\\
\hline
 0.05 & 3.8392 & $0.9396$ & $0.9069$\\
\hline 
 0.01 & 15.1798 & $0.7964$ & $0.6361$\\
\hline
\end{tabular}
\caption{Behavior of $\rho(H)$ and $\rho(\hat{H})$ for the ifiss advection 
diffusion matrix with factorized AINV preconditioning.} 
\label{tab:ifiss_ainv}
\end{table}

\subsection{The Poisson equation solved with MCSA}
In the current subsection a Poisson problem is taken into account:
\begin{equation}
\begin{cases}
 -\Delta u = f \quad \text{in}\quad \Omega \\
 u=0\quad \text{on} \quad \partial\Omega
 \end{cases}
\end{equation}

where $\Omega=(0,1)\times (0,1)$, $\sigma=1$. 
A finite difference scheme is employed to discretize the problem.
As a right hand side, a 
sinusoidal distribution in both $x$ and $y$ directions is considered, so that 
\[
 f(x_i, y_i)=sin\bigg(\frac{\pi x}{N-1}\bigg)sin\bigg(\frac{\pi y}{N-1}\bigg).
\]
N is the number of node considered on each direction, in our case $N=32$. The 
resulting linear system has 900 d.o.f.'s and $h=\frac{1}{29}$.
By applying a left diagonal preconditioning on the matrix, it is possible to 
reformulate the problem as a fixed point scheme. 
The iteration matrix $H$ is characterized by the fact that $\lVert 
H\rVert_1=1$, which is not sufficient to guarantee the convergence of 
any 
stochastic scheme analyzed in this work. However, by theoretical results, it is 
well known that applying a left diagonal preconditioning to the problem at hand 
produces an iteration matrix $H$ such that
\[
 \lim_{h\rightarrow 0}\rho(H)=1^{-}
\]
where $h$ is the spatial discretization step.
This is enough to have the guarantee that the deterministic fixed point scheme 
converges without the explicit computation of the spectral radius $\rho(H)$. 
However, the smaller is the discretization step $h$, the higher is the number of 
iterations required to reach a prescribed convergence criterion. For our 
discretization setting, the iteration matrix $H$ has a 
spectral radius of $\rho(H)\approx 0.9949$.

In order to have the guarantee of the convergence of the Adjoint Monte Carlo 
method, it is necessary to look at the spectral radius of $\hat{H}$ too. If an 
almost optimal probability is used, this implies that 
the Adjoint Monte Carlo method has a $\hat{H}$ matrix such that 
\[
 \lVert \hat{H}\rVert_1\le (\rVert H\rVert_1)^2=1.
\]
This condition by itself is not enough to guarantee that 
$\rho(\hat{H})<1$.

However, it must be recalled that the iteration matrix $H$ has zero 
valued entries on the main diagonal and it has: 
\begin{itemize}
 \item four entries equal to $\frac{-1}{4}$ on the rows associated 
with a node which is not adjacent to the boundary
\item two entries equal to $\frac{-1}{4}$ on the rows associated with nodes 
adjacent to the corner of the square domain
\item three entries equal to $\frac{-1}{4}$ on the rows associated with nodes 
adjacent to the boundary on an edge of the square domain.
\end{itemize}

Recalling the definition of $\hat{H}$ entrywise in terms of the 
iteration matrix $H$ and the transition probability $P$, we get that

\[
 \hat{H}=\tilde{D}H
\]

where $\tilde{D}$ is a diagonal matrix with diagonal entries equal to
1, $\frac{1}{2}$ or $\frac{3}{4}$. 
In particular $\tilde{d}_i=diag(\tilde D)_i=1$ if the $i$-th row of the 
discretized laplacian is associated with a node which is not adjacent to the 
boundary, $\tilde{d}_i=diag(\tilde D)_i=\frac{1}{2}$ if the row is associated 
with a node of the grid adjacent to the corner of the boundary, 
$\tilde{d}_i=diag(\tilde D)_i=\frac{3}{4}$ if the associated node of the grid 
is adjacent to the boundary of the domain far from a corner.

Since $H$ is an irreducible nonnegative matrix and $\tilde{D}$ is a 
positive definite diagonal matrix, we resort to a result presented in 
\cite{FK1975} to say that
\[
 \rho(\hat{H})=\rho(\tilde{D}H)\le \rho(\tilde D)\rho(H).
\]

Since $\rho(\tilde{D})=1$, then $\rho(\hat{H})\le \rho(H)<1$. 
Therefore the diagonal preconditioning always guarantees a convergent 
setting for the Monte Carlo linear solver, once it is applied on a 
two-dimensional Laplace 
operator discretized with finite differences. By a similar reasoning, the same 
results can be obtained for a $n$-dimensional laplace operator.

The computation of the solution is realized both with the deterministic 
Synthetic Acceleration and with Adjoint MCSA to compare the effectiveness of 
the 
stochastic technique. Results are reported in Table  
\ref{Poisson_results}. For the sake of completeness, results obtained with 
the Sequential Monte Carlo are shown as well. An evident discrepancy in terms 
of numerical iterations 
is already detected with the Sequential Monte Carlo outperforming the 
preconditioned fixed point. Results are improved even more 
when 
switching to the MCSA, which employs less histories than the Sequential MC at 
each iteration to 
reach convergence. In fact the average number of histories employed by the 
Sequential 
Monte Carlo per iteration is almost one order of magnitude higher than for 
MCSA. This 
might be explained by the fact that the first row of the iterative algorithm in 
the MCSA refines the accuracy of the solution coming from the previous 
iteration, 
before using it as the input for the MC solver at the current iterate. 
In particular, since the first line of the algorithm is a deterministic step of 
the Synthetic Acceleration, its effect is to dampen the statistical noise 
produced by the Monte Carlo linear solver to estimate the update $\delta 
x^{n + \frac{1}{2}}$ at the previous iteration. Therefore
refinement provided by the Synthetic Acceleration step decreases the number of 
random walks 
to employ for a prescribed accuracy. This is validated even more by the fact 
that for the first iteration the Sequential Monte Carlo and the MCSA resort to 
the same number of histories, while their behaviours start differing from the 
second iteration on, when the statistical noise is introduced in the estimation 
of the solution to \ref{linsys_prec} (see Figures \ref{SEQ_poisson} and 
\ref{MCSA_poisson}).

\begin{table}[!h]
\centering
\hspace*{-0.8cm}
\begin{tabular}{|c|c|c|c|}
\hline
algorithm & relative err.& \# iterations & average \# histories per iteration\\
\hline
Synthetic Acceleration & $9.8081\cdot 10^{-8}$ & 3133 & - \\ 
\hline
Sequential MC & $7.9037 \cdot 10^{-8}$ & 9 & 8.264,900\\
\hline
 Adjoint MCSA & $8.0872\cdot 10^{-8}$ & 8 & 1,738,250\\
\hline
\end{tabular}
\caption{Numerical results for the Poisson problem.}
\label{Poisson_results}
\end{table}

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{SEQ_poisson.jpg}
    \caption{Sequential MC - Poisson problem. Number of random walks employed 
at each numerical iteration.}
\label{SEQ_poisson}
\end{figure}

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{MCSA_poisson.jpg}
      \caption{MCSA - Poisson problem. Number of random walks employed at each 
numerical iteration.}
\label{MCSA_poisson}
\end{figure}

\subsection{A diffusion reaction problem solved with MCSA}
By modifying the equation of the previous test case, we consider now 
an advection reaction problem

\begin{equation}
\begin{cases}
 -\Delta u +\sigma u= f \quad \text{in}\quad \Omega \\
 u=0\quad \text{on} \quad \partial\Omega
 \end{cases}
\end{equation}
where $\Omega=(0,1)\times (0,1)$, $\sigma=0.1$ and $f=1$. 
A finite difference scheme is applied to discretize the problem.
The number of nodes 
on each direction of the domain is 100, so that $h\approx 0.01$. The 
discretized problem has 9604 d.o.f's. A left 
diagonal preconditioning is still applied to 
the stiffness matrix obtained from the discretization. The 1-norm of the 
iteration matrix is $\lVert H\rVert_1\approx 0.9756$. This automatically 
guarantees the convergence of the Adjoint Monte Carlo linear solver. The 
computation of the solution to the partial differential problem is still 
accomplished with the MCSA algorithm. A threshold of $\varepsilon =10^{-8}$ is 
used as a stop criterion for the relative residual. The threshold 
for the adaptive selection of the random walks instead is set 
to $\varepsilon_1=0.1$.
In Table \ref{DR_results} a comparison between the deterministic Synthetic 
Acceleration 
scheme, Sequential MC and the MCSA is provided. 
The employment of the MCSA provides a significant reduction of the number of 
histories necessary to satisfy the convergence criterion, in line with results 
coming from the previous test case.

\begin{table}[!h]
\centering
\hspace*{-0.8cm}
\begin{tabular}{|c|c|c|c|}
\hline
algorithm & relative err.& \# iterations & average \# histories per iteration\\
\hline
 Synthetic acceleration & $9.073\cdot 10^{-8}$ & 634 & - \\ 
\hline
 Sequential MC & $8.415 \cdot 10^{-8}$ &  8 & 12,391,375\\
 \hline
 Adjoint MCSA & $6.633 \cdot 10^{-8}$ &  7 & 3,163,700\\
\hline
\end{tabular}
\caption{Numerical results for the diffusion reaction problem.}
\label{DR_results}
\end{table}

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{SEQ_diffreac.jpg}
      \caption{Sequential MC - Diffusion reaction problem. Number of random 
walks 
employed at each 
numerical iteration.}
\label{SEQ_diffreac}
\end{figure}

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{MCSA_diffreac.jpg}
      \caption{MCSA - Diffusion reaction problem. Number of random walks 
employed at each 
numerical iteration.}
\label{MCSA_diffreac}
\end{figure}

\subsection{A parabolic problem solved with MCSA}
In this subsection we restrict our treatise to the analysis of parabolic 
partial 
differential problems of the form 
\begin{equation}
\begin{cases}
 \frac{\partial u }{\partial t} + \mathcal{L}(u)=f \quad \text{in} \Omega \\
\text{B.C.} \quad \text{on}\quad \delta \Omega.
\end{cases}
\label{parabolic}
\end{equation}

The partial derivative in time is discretized with finite differences, whereas 
quadrilateral linear finite elements are employed for the space discretization. 
Naming $N_h$ the number of d.o.f.'s associated with the finite element 
discretization, the problem \ref{parabolic} turn into a fully discretized 
problem such as:

\[
 \bigg (\frac{1}{\Delta t}M+A\bigg )(\mathbf{u}^{k+1}-\mathbf{u}^k)+A[\theta 
\mathbf{u}^{k+1}+(1-\theta)\mathbf{u}^k]=\theta 
\mathbf{f}^{k+1}+(1-\theta)\mathbf{f}^k,
\]

where $k$ is an index that goes with the time dimension.

By tuning the value of the parameter $\theta$ different numerical schemes can 
be obtained with different properties of numerical stability. 
However our discussion is not involved in this kind of issues.
In particular we restrict our attention to a single generic time step 
associated with an Implicit Euler time discretization (corresponding to 
$\theta=1$). The vector of the right hand side is chosen so that the exact 
solution to the linear system for the specific time step chosen is a unit 
vector $\tilde{\mathbf{u}}=\mathbbm{1}^{N_h}$.

By referring to $h$ as the space discretization step, we pick
\[
 \Delta t \le C h.
\]

Give an open and bounded set $\Omega=(0,1)\times(0,1)$, we consider the 
following problem
\begin{equation}
 \begin{cases}\frac{\partial u}{\partial t} -\mu \Delta u 
+\boldsymbol{\beta}(\mathbf{x})\cdot \nabla u=0, \;\; \quad \quad 
\mathbf{x}\in \Omega,\quad t\in(0,T] \\
u(\mathbf{x},0) = u_0, \qquad \qquad \qquad \qquad \quad \mathbf{x}\in 
\Omega \\
u(\mathbf{x},t)=u_D(\mathbf{x}),  \; \, \, \qquad \qquad \qquad \quad  
\mathbf{x}\in \partial 
\Omega, \quad t\in(0,T],
 \end{cases}
\end{equation}

where $\mu=\frac{3}{200}$, $\boldsymbol{\beta}(\mathbf{x})=[2y(1-x^2),\; 
-2x(1-y^2)]^T$, \\$u_D=0$ on $\{\{x=0\}\times (0,1)\}$, $\{(0,1)\times 
\{y=0\}\}$, $\{(0,1)\times \{y=1\}\}$.\newline

The discretization of the problem is realized with the \texttt{IFISS} toolbox.
The value of the discretization step is $h=2^{-8}$, generating a discretized 
linear system with 66,049 d.o.f.'s.
As concerns the time discretization, the time step is set to $\Delta t = 
10h$.\newline

The sparse approximate inverse is employed for 
the linear system at hand as a right preconditioner, with a drop tolerance of 
$\tau=0.05$ for both the factors.

In this situation it is not possible to 
resort to any of the sufficient conditions to verify the convergence. Therefore 
the only viable option is the explicit computation of the spectral radii of $H$ 
and $\hat{H}$.
The spectral radius of the iteration matrix is $\rho(H)\approx 
0.9218$ and the spectral radius of $\hat{H}$ for the Adjoint Monte Carlo is 
$\rho(\hat{H})\approx 0.9148$. The transition probability employed is the 
Almost 
Optimal one. Resorting to a Uniform Probability in this case would have impeded 
the 
convergence, since in this case $\rho(\hat{H})\approx 1.8401$. 
Therefore 
this is an example demonstrating that the Almost Optimal Probability 
facilitates the performance of the stochastic algorithm, outperforming the 
Uniform one.

The fill-in effect is 
\[
 \frac{nnz(H)}{nnz(A)}=4.26,
\]
therefore the relative number of nonzero elements in $H$ is still acceptable 
in terms of sparsity pattern and memory storage.

The threshold for the check on the relative residual is set to $\varepsilon 
=10^{-8}$ is, whilst the threshold 
for the adaptive selection of the random walks is set 
to $\varepsilon_1=0.1$.
In order to test the effectiveness of employing the MCSA with respect to a 
purely deterministic scheme, a comparison with the Richardson algorithm and 
the Sequential Monte Carlo is 
accomplished. The results are shown in Table \ref{parabolic_results}. As you 
can see, the adoption of both Sequential Monte Carlo and MCSA reduces 
dramatically the number of iterations with 
respect to the fixed point method. Of course this 
advantage has the 
increase of the computational time as a payoff. Nevertheless we put off this 
discussion which is of 
interest for a parallelized implementation of the method. Even more the number 
of random walks employed by the Sequential MC is higher than the one associated 
with the MCSA, in accordance with the previous results.

\begin{table}[!h]
\centering
\hspace*{-0.8cm}
\begin{tabular}{|c|c|c|c|}
\hline
algorithm & relative err.& \# iterations & average \# histories per iteration\\
\hline
 Richardson & $6.277\cdot 10^{-7}$ & 178 & - \\
 \hline
 Sequential MC & $1.918 \cdot 10^{-9}$ & 8 & 51,535,375\\ 
\hline
 Adjoint MCSA & $1.341\cdot 10^{-9}$ & 6 & 16,244,000\\
\hline
\end{tabular}
\caption{Numerical results for the parabolic problem.}
\label{parabolic_results}
\end{table}

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{SEQ_parabolic.jpg}
      \caption{Sequential Monte Carlo - Parabolic problem. Number of random 
walks 
employed at each 
numerical iteration.}
\label{MCSA_parabolic}
\end{figure}

\begin{figure}[h!]
  \centering
    \includegraphics[width=\textwidth]{MCSA_parabolic.jpg}
      \caption{MCSA - Parabolic problem. Number of random walks 
employed at each 
numerical iteration.}
\label{MCSA_parabolic}
\end{figure}

Other preconditioners 
have 
been 
tested besides the sparse approximate inverse preconditioner, such as block 
diagonal 
and incomplete LU factorization. However the fill-in effect introduced by the 
application of 
these preconditioners compromises the sparsity of the linear system.

\section{Conclusions and future developments}
The work presented in this paper has been accomplished to provide an overview 
about MC linear solvers. In particular the goal was to combine the main 
theoretical results found in literature with additional novel contributions on 
both the theoretical and empirical sides. This made it possible to understand 
the actual limitations of the presented algorithms.

Necessary and sufficient conditions to ensure convergence are too costly for 
most of the applicative problems coming fro engineering and Computational 
Physics. On the other hand, sufficient criteria are currently viable for a
restricted set of matrices (generalized diagonally dominant).

An empirical analysis has been carries out on parabolic problems and it has 
been shown that a proper selection of the time discretization step enables the 
employment of stochastic solvers.

We reserve for future works the analysis of the plausible connections between 
$\rho(H)$, $\rho(\hat{H})$ and the number of numerical iterations necessary to 
reach convergence. Moreover a study about the resilience of the problem will be 
accomplished, resorting to fault injection techniques.

\addcontentsline{toc}{chapter}{\refname}
\bibliographystyle{alpha}
\bibliography{library}
%\nocite{*}

\end{document}