### 2015-09-16 MAX_reports/MC_solvers_article/MC_solvers.tex
Breaking paper into multiple files for easier editing/collaborating.

\documentclass[a4paper,10pt]{article}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{mathtools}
\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{fancyhdr}
\usepackage{geometry}
\usepackage{subfig}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subfig}
\usepackage{tabu}
\usepackage{multirow}
\usepackage{color}
\usepackage{listings}
\usepackage[]{algorithm2e}
\usepackage[autostyle,english=british]{csquotes}
\usepackage{minitoc}

\newtheorem{thm}{Theorem}
\newtheorem{defn}{Definition}
\newtheorem{prop}{Proposition}

%opening
\title{Monte Carlo solvers for sparse linear systems}
\author{Michele Benzi, Massimiliano Lupo Pasini}
\date{}

\begin{document}

\maketitle

\begin{abstract}
We consider linear systems of equations characterized by a sparsity pattern,
focusing in particular on problems derived from the discretization of partial
differential equations. We propose a treatise of the current stochastic
techniques present in literature that might be employed to solve the problems at
hand. New theoretical results are introduced as sufficient criteria for the
guarantee of the convergence of the methods analyzed. An analysis about viable
preconditioners is accomplished as well.
\end{abstract}

\input{intro}
\input{mcls}
\input{convergence}
\input{results}
\input{conclusion}

\addcontentsline{toc}{chapter}{\refname}
\bibliographystyle{alpha}
\bibliography{library}
%\nocite{*}

\end{document}