### 2016-01-25 MAX_reports/MC_solvers_article/mcls.tex

\section{Stochastic Linear Solvers}
\label{sec:mcls}

We consider the solution of systems of linear equations of the form
\begin{equation}
A \mathbf{x}=\mathbf{b},
\label{linsys}
\end{equation}
where $A\in \mathbb{R}^{n\times n}$ and $\mathbf{x}$, $\mathbf{b} \in
\mathbb{R}^n$.

Equation~(\ref{linsys}) can be reinterpreted as a fixed point scheme
\begin{equation}
 \mathbf{x}=H\mathbf{x}+\mathbf{f}.
 \label{fixedpoint}
\end{equation}
where $H=I-A$ and $\mathbf{f}=\mathbf{b}$.

Assuming that the spectral radius $\rho(H)<1$, the solution to
(\ref{fixedpoint}) can be written in terms of a power series of
$H$ (Neumann series)
\[
\mathbf{x}=\sum_{i=0}^\infty H^i\mathbf{f}.
\]
Therefore the fixed point scheme generates a sequence of approximate solutions
$\{\mathbf{x}^{(k)}\}_{k=0}^{\infty}$ which converges to the exact solution
regardless of the initial guess $\mathbf{x}_0$.

By restricting the attention to a single component of $\mathbf{x}$ we
have
\begin{equation}
x_i=\sum_{\ell=0}^\infty \sum_{k_1}^n\sum_{k_2}^n\cdots \sum_{k_{\ell}=1}^n
H_{k_0,k_1}H_{k_1,k_2}\cdots H_{k_{\ell-1}, k_{\ell}}f_{k_{\ell}}.
\label{forward}
\end{equation}
The last equation can be reinterpreted as the realization of an estimator
defined on a random walk.  Let us start considering a random walk whose
state space $S$ is characterized by the set of indices of the forcing term
$\mathbf{f}$:
\[
S=\{1,2,\cdots, n\} \subset \mathbb{N}.
\]
Each $i$-th step of the random walk has a random variable
$k_i$ associated with it. The realization of $k_i$ represents the index of the
component of $\mathbf{f}$
which is visited in the current step of the random walk.
The method of building the transition probabilities and the selection of
the initial state of the random walk gives rise to two different approaches:
the \textit{forward} and \textit{adjoint} methods.

\subsection{Forward Method}
\label{subsec:forward}

The goal is to evaluate a functional such as
\[
J(\mathbf{x})=(\mathbf{h},\mathbf{x})=\sum_{i=1}^n h_i x_i.
\]
where $\mathbf{h}\in \mathbb{R}^n$ is the Riesz representative in
$\mathbb{R}^n$ of the functional $J$.
We can use it to build the initial probability $\tilde{p}:
S\rightarrow [0,1]$ of the random walk such that
\[
\tilde{p}(k_0=i)=\tilde{p}_{k_0}=\frac{\lvert h_i\rvert}{\sum_{i=1}^n \lvert
h_i\rvert}.
\]
It is important to stress out that the role of vector $\mathbf{h}$ is
restricted to the construction of the initial probability and is not
used further in the definition of the stochastic process.
A possible choice for the transition probability $P$ can be
\[
p(k_i=j \;\lvert\;k_{i-1}=i )=P_{i,j}=\frac{\lvert H_{i,j}\rvert}{\sum_{k=1}^n
\lvert H_{i,k}\rvert}.
\]
where $\tilde{p}(\cdot,i):S\rightarrow [0,1]$ $\forall i\in S$.
A related sequence of random variables $w_{i,j}$ can be defined
such that
\[
w_{i,j}=\frac{H_{i,j}}{P_{i,j}}.
\]
The probability distribution of the random variables $w_{i,j}$ is represented
by the transition matrix that rules the stochastic process. The $w_{i,j}$'s
just introduced can be used to build one more sequence
of random variables.
At first we introduce quantities $W_j$
\[
W_{0}=\frac{h_{k_0}}{\tilde{p}_{k_0}}, \quad W_j=W_{j-1} w_{i,j}, \quad
j=1,\cdots, i.
\]
By defining
\[
X(\nu)=\sum_{m=0}^k W_m f_{i_m}
\]
as the random variable associated with a specific permutation $\nu$, we can
define the estimator $\theta_i$ such as
\[
\theta=E[X]=\sum_{\nu}P_{\nu}X(\nu).
\]
The integer $n$ represents the size of the solution vector to (\ref{linsys})
and
the
index
$i$
is referred to the component of the solution vector we want to compute.
$P_{\nu}$ is the probability associated with a specific permutation of the
random walk.
It can be proved that
\[
E[W_i f_{k_i}]=(\mathbf{h},H^i\mathbf{f}), \quad i=0,1,2,\cdots
\]
and
\[
\theta_i=E\bigg[\sum_{i=0}^\infty W_i f_{k_i}\bigg]=(\mathbf{h},\mathbf{x}).
\]

A possible choice for $\mathbf{h}$ is a vector of the standard basis. This
would correspond in setting manually the initial state of the random walk,
which turns the related initial probability into a Kronecker delta
\[
\tilde{p}(k_0=i)=\delta_{i,j}.
\]
By doing so, we have
\begin{equation}
\theta_i=E\bigg[\sum_{\ell=0}^\infty W_{\ell}
f_{k_{\ell}}\bigg]=x_i=\sum_{l=0}^\infty
\sum_{k_1=1}^{n}\sum_{k_2=1}^n\cdots \sum_{k_{\ell}=1}^n
P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1},
k_{\ell}}w_{k_0,k_1}w_{k_1,k_2}\cdots
w_{k_{\ell-1}, k_{\ell}}f_{k_{\ell}}.
\label{dir_mean}
\end{equation}

As regards the variance, we remember that the following relation holds:
\begin{equation}
Var\bigg [\sum_{\ell=0}^\infty W_{\ell}
f_{k_{\ell}}\bigg]=E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2
f_{k_{\ell}}^2\bigg] - \bigg (E\bigg[\sum_{\ell=0}^\infty W_{\ell}
f_{k_{\ell}}\bigg]\bigg )^2.
\label{dir_var}
\end{equation}

Hence the variance can be computed as the difference between the second
moment of the random variable and the square of the first moment.\newline

In order to apply the Central Limit Theorem (CLT) to the estimators defined
above, we must require that
the estimators have both finite expected value and variance. This is
equivalent in
checking the finiteness of the expected value and of the second moment.
Therefore we have to impose the following conditions:

\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell} f_{k_{\ell}}\bigg]<\infty
\end{equation}

\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2
f_{k_{\ell}}^2\bigg]<\infty
\end{equation}

The forward method presented above, however, has the limitation of employing an
entire set of permutations to estimate just a single entry of
the solution at a time. Therefore, in order to estimate the entire solution
vector for Eq.~\eqref{linsys}, we have to employ a separate set of
permutations for each entry in the solution vector.

\subsection{Adjoint Method}
\label{subsec:adjoint}

A second Monte Carlo method can be derived by considering the linear system
adjoint to (\ref{linsys})
\begin{equation}
A^T\mathbf{y}=\mathbf{d},
\label{adj}
\end{equation}
where $\mathbf{y}$ and $\mathbf{d}$ are the adjoint solution and source term.

\textcolor{red}{Provide some more details relating the adjoint to original problem.}

By reformulating the fixed point scheme, introducing initial probability,
transition probability and weight
sequences in the same fashion as done before, the expected value for the
estimator becomes
\begin{equation}
\theta_j=E\bigg[\sum_{\ell=0}^\infty W_{\ell}\delta_{k_{\ell},
j}\bigg]=\sum_{\ell=0}^{\infty}\sum_{k_1}^n\sum_{k_2}^n\cdots\sum_{k_{\ell}}^n
f_{k_0}P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1},K_{\ell}}w_{k_0,k_1}\cdots
w_{k_{\ell-1},k_{\ell}}\delta_{k_{\ell},j}.
\label{adj_mean}
\end{equation}
This estimator is known in literature as \textit{collision} estimator.

The forward method adds a contribution to the component of the solution
vector where the random walk began based on the value of the source vector
in the state in which the walk currently resides.  The adjoint method,
on the other hand, adds a contribution to the component of the solution
vector where the random walk currently resides based on the value of the
source vector in the state in which the walk began.
The Kronecker delta at the end of the series represents a filter, indicating
that only a subset of permutations contribute to the $j$-th component
of the solution vector.

The variance is given by
\begin{equation}
Var\bigg [\sum_{\ell=0}^\infty W_{\ell}
f_{k_0}\delta_{k_{\ell},j}\bigg]=E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2
f_{k_0}^2\delta_{k_{\ell},j}\bigg ] - \bigg (E\bigg[\sum_{\ell=0}^\infty
W_{\ell}
f_{k_0}\delta_{k_{\ell},j}\bigg]\bigg )^2\quad j=1,\cdots,n
\label{adj_var}.
\end{equation}

Along the same lines as the development for the forward method, we must
impose finiteness of the expected value and second moment.
Therefore the following
conditions must be verified:

\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell}\delta_{k_{\ell},
j}\bigg]<\infty \quad j=1,\cdots,n
\end{equation}
and
\begin{equation}
 E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2
f_{k_0}^2\delta_{k_{\ell},j}\bigg]<\infty \quad j=1,\cdots,n.
\end{equation}

The main advantage of this method, compared to the forward one, consists in the
fact that a single set of permutations is used to estimate the entire solution
vector.  Unless only a small portion of the problem is of interest, this
property often leads to the adjoint method being favored over the forward method.

In literature another estimator is employed along with the Adjoint Monte Carlo
method, the so called \textit{expected value} estimator. Its
formulations is as follows:
\begin{equation}
\theta_j=E\bigg[f_j + \sum_{\ell=0}^\infty
W_{\ell}H_{k_{\ell}, j}^T\bigg]=f_j
+ \sum_{\ell=0}^{\infty}\sum_{k_1}^n\sum_{k_2} ^n\cdots\sum_ { k_ { \ell}}^n
f_{k_0}P_{k_0,k_1}P_{k_1,k_2}\cdots P_{k_{\ell-1},K_{\ell}}w_{k_0,k_1}\cdots
w_{k_{\ell-1},k_{\ell}}H_{k_{\ell},j}^T.
\label{adj_mean1}
\end{equation}

Thus, the expected value estimator averages
the deterministic contribution of the iteration matrix over all the potential
states $j$ that might be reached from the current state $\ell$. The variance
in this case becomes:
\begin{equation}
Var\bigg [\sum_{\ell=0}^\infty W_{\ell}
f_{k_0}H_{k_{\ell},j}^T\bigg]=E\bigg[\sum_{\ell=0}^\infty W_{\ell}^2
f_{k_0}^2 {H_{k_{\ell},j}^T}^T\bigg ] - \bigg (E\bigg[\sum_{\ell=0}^\infty
W_{\ell}
f_{k_0}H_{k_{\ell},j}^T\bigg]\bigg )^2\quad j=1,\cdots,n
\label{adj_var}.
\end{equation}

\subsection{Hybrid Stochastic/Deterministic Methods}

The direct methods described in Sections~\ref{subsec:foward} and
\ref{subsec:adjoint} suffer from a slow rate of convergence due to the
$\frac{1}{\sqrt{N}}$ behavior dictated by the central limit theorem ($N$ here
is the number of random walks used to estimate the solution).  Furthermore,
when the spectral radius of the iteration matrix is close to unity, each
individual random walk may require a large number of transitions to
approximate the corresponding components in the Neumann series.
To offset the slow convergence of the central limit theorem, schemes have
been proposed which combine traditional fixed-point iterative methods with
the stochastic solvers.  The first such method, due to Halton, was termed
the Sequential Monte Carlo (SMC) method, and can be written as:
\begin{algorithm}[H]
 \KwData{$H$, $\mathbf{b}$, $\mathbf{x}_0$}
 \KwResult{$x_{num}$}
 $\mathbf{x}^{l}=\mathbf{x}_0$\;
 \While{not reached convergence}{
  $\mathbf{r}^{l}=\mathbf{b}-A\mathbf{x}^{l}$\;
  $\delta \mathbf{x}^{l+1}=(I-H)^{-1}\mathbf{r}^{l}$\;
  $\mathbf{x}^{l+1}=\mathbf{x}^{l}+\delta \mathbf{x}^{l+1}$\;
 }
 $x_{num}=x^{l+1}$\;
 \caption{Sequential Monte Carlo}
\end{algorithm}
The Monte Carlo linear solver method is used to compute the update $\delta
\mathbf{x}^{l}$.  This algorithm is equivalent to a preconditioned Richardson
iteration in which the Monte Carlo process acts as the preconditioner.
Because Monte Carlo is only applied within a single iteration, the central
limit theorem is only applicable within that iteration rather than to
the overall convergence behavior of the algorithm.

A further extension of Halton's method, termed
\textit{Monte Carlo Synthetic Acceleration} (MCSA), has been recently
introduced (\cite{ESW2013} and \cite{EMSH2014}).  The MCSA algorithm can be written as
\begin{algorithm}[H]
 \KwData{$H$, $\mathbf{b}$, $\mathbf{x}_0$}
 \KwResult{$x_{num}$}
 $\mathbf{x}^{l}=\mathbf{x}_0$\;
 \While{not reached convergence}{
  $\mathbf{x}^{l+\frac{1}{2}}=H\mathbf{x}^l+\mathbf{b}$\;
  $\mathbf{r}^{l+\frac{1}{2}}=\mathbf{b}-A\mathbf{x}^{l+\frac{1}{2}}$\;
  $\delta \mathbf{x}^{l+\frac{1}{2}}=(I-H)^{-1}\mathbf{r}^{l+\frac{1}{2}}$\;
  $\mathbf{x}^{l+1}=\mathbf{x}^{l+\frac{1}{2}}+\delta
\mathbf{x}^{l+\frac{1}{2}}$\;
 }
 $x_{num}=x^{l+1}$\;
 \caption{Monte Carlo Synthetic Acceleration}
\end{algorithm}
As with SMC, a Monte Carlo linear solver is used to compute the updating
contribution $\delta \mathbf{x}^{l+\frac{1}{2}}$.  In this approach, an
extra step of unpreconditioned Richardson iteration is added to smooth out
some of the high-frequency noise introduced by the Monte Carlo process.
This way, the deterministic and stochastic components of the algorithm
act in a complementary fashion.