### 2016-01-25 doc/NCState_seminar_2014/ncstate_seminar.tex
adding nc state seminar tex file

%%---------------------------------------------------------------------------%%
%% Shift Roadmap Presentation
%%---------------------------------------------------------------------------%%

\documentclass{beamer}
\usefonttheme[onlymath]{serif}
\usepackage{colortbl}
\usepackage{longtable,ltcaption,array}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm} \usepackage{amsmath} \usepackage{tmadd,tmath}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{caption}
\usepackage[labelformat=empty]{subcaption}
\usepackage{tabularx}

%%---------------------------------------------------------------------------%%
%% THEME SETUP

\usetheme{CambridgeUS}
\usecolortheme{spruce}

%%---------------------------------------------------------------------------%%
%% SETUP STUFF
%%---------------------------------------------------------------------------%%

\logo{\includegraphics[width=2cm]{new_logo}}

\setbeamercolor{item}{fg=MSUgreen}
\setbeamertemplate{headline}{}
\setbeamertemplate{navigation symbols}{}

\newlength{\DUtablewidth} % internal use in tables

%%---------------------------------------------------------------------------%%
%% MATH STUFF
%%---------------------------------------------------------------------------%%

\newcommand{\Pn}{\textit{P}$\negthinspace_N$}
\newcommand{\SPn}{\textit{SP}$\negthinspace_N$}

\newcommand{\vOmega}{\ensuremath{\ve{\Omega}}}
\newcommand{\hOmega}{\ensuremath{\hat{\ve{\Omega}}}}

\newcommand{\sigs}{\ensuremath{\sigma_{\text{s}}}}
\newcommand{\sigf}{\ensuremath{\sigma_{\text{f}}}}
\newcommand{\sigsm}{\ensuremath{\sigma_{\text{s}m}}}
\newcommand{\sigsn}{\ensuremath{\sigma_{\text{s}n}}}

\newcommand{\phig}{\ensuremath{\Phi}}
\newcommand{\Sigmag}{\ensuremath{\mathbf{\Sigma}}}
\newcommand{\Dg}{\ensuremath{\mathbb{D}}}
\newcommand{\Fg}{\ensuremath{\mathbb{F}}}
\newcommand{\Qg}{\ensuremath{\mathbb{Q}}}
\newcommand{\ug}{\ensuremath{\mathbb{U}}}
\newcommand{\Ag}{\ensuremath{\mathbb{A}}}
\newcommand{\Bg}{\ensuremath{\mathbb{B}}}
\newcommand{\Cg}{\ensuremath{\mathbb{C}}}
\newcommand{\Jg}{\ensuremath{\mathbb{J}}}
\newcommand{\sg}{\ensuremath{\mathbb{S}}}

\newcommand{\aphig}{\ensuremath{\Phi^{\dagger}}}
\newcommand{\aSigmag}{\ensuremath{\mathbf{\Sigma}^{\dagger}}}

\newcommand{\apsi}[1]{\ensuremath{\psi^{\dagger\,#1}}}
\newcommand{\aphi}[1]{\ensuremath{\phi^{\dagger\,#1}}}
\newcommand{\aq}[1]{\ensuremath{q^{\dagger\,#1}}}
\newcommand{\ak}{\ensuremath{k^{\dagger}}}

\newcommand{\normal}{\ensuremath{\hat{\ve{n}}}}

% Phantom minus sign for helping with alignment of matrices
\newcommand{\phmin}{\ensuremath{\phantom{-}}}

\DeclareMathOperator{\diag}{diag}

%%---------------------------------------------------------------------------%%
%% TITLE

\title[Monte Carlo Linear Solvers]{Monte Carlo Linear Solvers}
\author[Slattery, Evans, Hamilton]{Stuart Slattery \\ Tom Evans
  \\ Steven Hamilton}
\date[11/4/14]{November 4, 2014}
\institute[ORNL]{\small Oak Ridge National Laboratory}
\titlegraphic{\includegraphics[width=1.5in]{ORNL_Balls.pdf}}

\defbeamertemplate{footline}{titlefoot}{
    \vspace{-0.5cm}
    \hspace*{0.25cm}
    \includegraphics[width=2cm]{doe_logo}
    \hfill
    \includegraphics[width=3.25cm]{WordMarkLeaf}
    \hspace*{0.5cm}
}

\setbeamertemplate{title page}
{
  \begin{tabular}{cr}
    \begin{minipage}{4.7cm}
      {\bf \Large\textcolor{MSUgreen}{\inserttitle}}\\

      \vspace{2\baselineskip}
      \insertauthor\\

      \insertinstitute\\

      \insertdate
    \end{minipage}
    &
    \begin{minipage}{5cm}
      \raggedright
      \vspace{-0.75cm}
      \includegraphics[height=9.8cm]{combined_background}
    \end{minipage}
  \end{tabular}
}

%%---------------------------------------------------------------------------%%
\begin{document}

%%---------------------------------------------------------------------------%%

{
\setbeamertemplate{logo}{}
\setbeamertemplate{footline}[titlefoot]
\begin{frame}
\titlepage
\end{frame}
}

%%---------------------------------------------------------------------------%%
\begin{frame}{Acknowledgements}

  This material is based upon work supported by the U.S. Department of
  Energy, Office of Science, Advanced Scientific Computing Research
  program.

  \vfill

  This research used resources of the Oak Ridge Leadership Computing
  Facility at the Oak Ridge National Laboratory, which is supported by the
  Office of Science of the U.S. Department of Energy under Contract No.
  DE-AC05-00OR22725.

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Motivation}
  \vfill
\begin{itemize}
  \item As we move towards exascale computing, the rate of errors is expected
  to increase dramatically
  \vfil
  \begin{itemize}
  \item The probability that a compute node will fail during the course
    of a large scale calculation may be near 1
  \end{itemize}
  \vfill
\item Algorithms need to not only have increased concurrency/scalability
  but have the ability to recover from hardware faults
  \begin{itemize}
  \item Lightweight machines
  \item Heterogeneous machines
  \item Both characterized by low power and high concurrency
  \end{itemize}
  \vfill
\end{itemize}
\vfill
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Towards Exascale Concurrency and Resiliency}
  \begin{itemize}
    \item Two basic strategies:
      \vfill
      \begin{enumerate}
        \item State with current ``state of the art'' methods and make
          incremental modifications to improve scalability and fault
          tolerance
          \begin{itemize}
            \item Many efforts are heading in this direction, attempting
              to find additional concurrency to exploit
          \end{itemize}
        \vfill
        \item Start with methods having natural scalability and resiliency
          aspects and work at improving performance (e.g. Monte Carlo)
          \begin{itemize}
            \item Soft failures buried within the tally variance
            \item Hard failures mitigated by replication
            \item Concurrency enabled by several levels of parallelism
          \end{itemize}
      \end{enumerate}
  \end{itemize}
  \vfill
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Monte Carlo Methods

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo for Linear Systems}
  \begin{itemize}
    \item Suppose we want to solve $\mathbf{Ax}=\mathbf{b}$
    \vfill
    \item If $\rho(\mathbf{I-A})<1$, we can write the solution using the
      Neumann series
      \begin{equation*}
        \mathbf{x} = \sum_{n=0}^{\infty} (\mathbf{I-A})^n \mathbf{b}
         = \sum_{n=0}^{\infty} \mathbf{H}^n \mathbf{b}
      \end{equation*}
      where $\mathbf{H} \equiv ( \mathbf{I-A} )$ is the Richardson
      iteration matrix 
      \vfill
    \item Build the Neumann series stochastically
  \end{itemize}

  \[
  x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}
  \]

  \begin{itemize}
  \item Define a sequence of state transitions
  \end{itemize}
  \vspace*{-0.1in}
  \[
  \nu = i \rightarrow i_1 \rightarrow \cdots \rightarrow i_{k-1}
  \rightarrow i_{k}
  \]

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Forward Monte Carlo}
\begin{itemize}
  \item Choose a row-stochastic matrix $\mathbf{P}$ and weight matrix
    $\mathbf{W}$ such that $\mathbf{H} = \mathbf{P} \circ \mathbf{W}$
  \item Typical choice (Monte Carlo Almost-Optimal):
    \begin{equation*}
      \mathbf{P}_{ij} = \frac{| \mathbf{H}_{ij}| }
      {\sum_{j=1}^{N} | \mathbf{H}_{ij} | }
    \end{equation*}
  \item To compute solution component $\mathbf{x}_i$:
    \begin{itemize}
      \item Start a history in state $i$ (with initial weight of 1)
      \item Transition to new state $j$ based probabilities determined by
        $\mathbf{P}_i$
      \item Modify history weight based on corresponding entry in
        $\mathbf{W}_{ij}$
      \item Add contribution to $\mathbf{x}_i$ based on current history weight
        and value of $\mathbf{b}_j$
    \end{itemize}
  \item A given random walk can only contribute to a single component of
    the solution vector
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Sampling Example (Forward Monte Carlo)}
  \begin{itemize}
    \item Suppose
  \begin{equation*}
    \mathbf{A} = \begin{bmatrix}
      \phmin 1.0 & -0.2 & -0.6 \\
      -0.4 & \phmin 1.0 & -0.4 \\
      -0.1 & -0.4 & \phmin 1.0 \end{bmatrix} \to
    \mathbf{H} \equiv (\mathbf{I-A}) = \begin{bmatrix}
       0.0 &  0.2 &  0.6 \\
       0.4 &  0.0 &  0.4 \\
       0.1 &  0.4 &  0.0 \end{bmatrix}
  \end{equation*}
    then
  \begin{equation*}
    \mathbf{P} = \begin{bmatrix}
       0.0 & 0.25 & 0.75 \\
       0.5 &  0.0 & 0.5 \\
       0.2 &  0.8 & 0.0 \end{bmatrix}, \quad
    \mathbf{W} = \begin{bmatrix}
       0.0 &  0.8 &  0.8 \\
       0.8 &  0.0 &  0.8 \\
       0.5 &  0.5 &  0.0 \end{bmatrix}
  \end{equation*}
    \vfill
    \item If a history is started in state $3$, there is a $20\%$ chance of
      it transitioning to state $1$ and an $80\%$ chance of moving to state
      $2$
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Solving the Heat Equation: Forward Method}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics<1>[width=4in]{direct_1.png}
      \includegraphics<2>[width=4in]{direct_10.png}
      \includegraphics<3>[width=4in]{direct_100.png}
      \includegraphics<4>[width=4in]{direct_1000.png}
    \end{center}
    \caption{
      \only<1>{\textbf{Forward solution.}
        \textit{\sn{2.5}{3} total histories.} }
      \only<2>{\textbf{Forward solution.}
        \textit{\sn{2.5}{4} total histories.} }
      \only<3>{\textbf{Forward solution.}
        \textit{\sn{2.5}{5} total histories.} }
      \only<4>{\textbf{Forward solution.}
        \textit{\sn{2.5}{6} total histories.} }
    }
  \end{figure}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Adjoint Monte Carlo}
\begin{itemize}
  \item Choose $\mathbf{P}$ and $\mathbf{W}$ such that
    $\mathbf{H}^{T} = \mathbf{P} \circ \mathbf{W}$
  \item Typical choice (Monte Carlo Almost-Optimal):
    \begin{equation*}
      \mathbf{P}_{ij} = \frac{| \mathbf{H}_{ji}| }
      {\sum_{i=1}^{N} | \mathbf{H}_{ji} |}
    \end{equation*}
  \item To estimate solution:
    \begin{itemize}
      \item Start a history in random state $i$ by sampling from distribution
        determined by $\mathbf{b}$
      \item Transition to new state $j$ based probabilities determined by
        $\mathbf{P}_i$
      \item Modify history weight based on corresponding entry in
        $\mathbf{W}_{ij}$
      \item Add contribution to $\mathbf{x}_j$ based on current history weight
    \end{itemize}
  \item A given random walk can contribute to many different components of
    the solution vector
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Solving the Heat Equation: Adjoint Method}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics<1>[width=4in]{adjoint_1.png}
      \includegraphics<2>[width=4in]{adjoint_10.png}
      \includegraphics<3>[width=4in]{adjoint_100.png}
      \includegraphics<4>[width=4in]{adjoint_1000.png}
      \includegraphics<5>[width=4in]{adjoint_10000.png}
      \includegraphics<6>[width=4in]{adjoint_100000.png}
      \includegraphics<7>[width=4in]{adjoint_1000000.png}
      \includegraphics<8>[width=4in]{adjoint_10000000.png}
    \end{center}
    \caption{
      \only<1>{\textbf{Adjoint solution.}
        \textit{\sn{1}{0} total histories.} }
      \only<2>{\textbf{Adjoint solution.}
        \textit{\sn{1}{1} total histories.} }
      \only<3>{\textbf{Adjoint solution.}
        \textit{\sn{1}{2} total histories.} }
      \only<4>{\textbf{Adjoint solution.}
        \textit{\sn{1}{3} total histories.} }
      \only<5>{\textbf{Adjoint solution.}
        \textit{\sn{1}{4} total histories.} }
      \only<6>{\textbf{Adjoint solution.}
        \textit{\sn{1}{5} total histories.} }
      \only<7>{\textbf{Adjoint solution.}
        \textit{\sn{1}{6} total histories.} }
      \only<8>{\textbf{Adjoint solution.}
        \textit{\sn{1}{7} total histories.} }
    }
  \end{figure}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Limitations of Monte Carlo}
\begin{itemize}
  \item Solving linear systems of equations with ``pure'' Monte Carlo methods
    is generally intractable
    \begin{itemize}
      \item Central limit theorem is barrier to accurate solutions
    \end{itemize}
\end{itemize}
\begin{figure}
  \centering
  \includegraphics[width=3.5in]{Ifiss_ConvDiff}
\end{figure}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Residual Monte Carlo Methods

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Sequential Monte Carlo}
  \begin{itemize}
  \item Devised by John Halton in the 1960's as a residual method
\vfill
    \item Instead of directly solving $\mathbf{Ax}=\mathbf{b}$ with Monte
      Carlo, apply Monte Carlo to residual equation
      $\mathbf{A}\delta = \mathbf{r}$
    \vfill
    \item Preconditioned Richardson iteration using Monte Carlo as
      preconditioner:
      \begin{align*}
        \mathbf{r}^{k} &= \mathbf{b} - \mathbf{Ax}^{k} \\
        \hat{\mathbf{A}} \mathbf{\delta} &= \mathbf{r}^{k} \\
        \mathbf{x}^{k+1} &= \mathbf{x}^k + \delta
      \end{align*}
    \vfill
    \item Exponential convergence achieved by decoupling Monte Carlo
      error from solution error
  \end{itemize}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Synthetic Acceleration}
  \begin{itemize}
  \item Devised by Evans and Mosher in the 2000's as an acceleration
    scheme for radiation diffusion problems (LANL)
    \vfill
  \item Can be abstracted as a general linear solver
    \vfill
  \item Combine with Richardson iteration as a ``smoother'' in between
    Monte Carlo steps:
    \begin{align*}
      \mathbf{r}^k &= \mathbf{b} - \mathbf{Ax}^k \\
      \mathbf{x}^{k+1/2} &= \mathbf{x}^k + \mathbf{r}^k \\
      \mathbf{r}^{k+1/2} &= \mathbf{b} - \mathbf{Ax}^{k+1/2} \\
      \hat{\mathbf{A}} \mathbf{\delta} &= \mathbf{r}^{k+1/2} \\
      \mathbf{x}^{k+1} &= \mathbf{x}^{k+1/2} + \delta
    \end{align*}
  \end{itemize}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Preconditioned MCSA}
\vspace*{-0.2in}
  {
    \small
    \[
    \ve{r}^{k} = \ve{M}_L^{-1}(\ve{b}-\ve{A}\ve{M}_R^{-1}\ve{x}^{k})
    \]
    \[
    \ve{x}^{k+1/2} = \ve{x}^k + \ve{r}^k
    \]
    \[
    \ve{r}^{k+1/2} = \ve{M}_L^{-1}(\ve{b}-\ve{A}\ve{M}_R^{-1}\ve{x}^{k+1/2})
    \]
    \[
    \ve{M}_L^{-1}\ve{A}\ve{M}_R^{-1}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}
    \]
    \[
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}
    \]
  }
  \vspace*{-0.1in}
  \begin{itemize}
    \item Requires explicit construction of preconditioned matrix
      \begin{itemize}
        \item This has generally limited preconditioner selection to diagonal,
          block diagonal, sparse approximate inverse approaches
      \end{itemize}
    \vfill
    \item Sparsity in $\mathbf{A}$, $\mathbf{M}$, does not imply sparsity
      in $\mathbf{M^{-1}A}$
    \vfill
    \item Need to investigate preconditioning strategies that lead to sparse
      preconditioned systems (Michele Benzi, Emory)
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Residual Algorithm Comparison}
  \begin{figure}
    \includegraphics[width=3.5in]{seq_conv_100}
  \end{figure}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{3D Equilibrium Thermal Radiation Diffusion}
\vspace*{-0.1in}

{ \tiny
  T. Evans, S. Mosher, S. Slattery, S. Hamilton, ``A Monte Carlo
  synthetic-acceleration method for solving the thermal radiation
  diffusion equation,'' Journal of Computational Physics \textbf{258},
  pp.~338--358 (2014).  
}
\vspace*{-0.1in}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{figure}[ht!]
        \centerline{ \includegraphics[width=2in,clip]{mesh.pdf}}
      \end{figure}
    \end{column}
    \begin{column}{0.5\textwidth}
      \begin{figure}[ht!]
        \centerline{ \includegraphics[width=2in,clip]{T_multi_mat.pdf}}
      \end{figure}
    \end{column}
  \end{columns}

\vspace*{-0.15in}
{\small
  \begin{table}
    \centering
    \begin{tabular}{cccc}
      \toprule
      \multirow{2}{*}{\bfseries Solver} &
      \bfseries Total &
      \bfseries Relative Time &
      \bfseries Relative Total \\
      & \bfseries Iterations &
      \bfseries Per Iteration &
      \bfseries Time \\
      \midrule
      CG (Jacobi) & 42,881 &  2.93 & 1.02 \\
      CG (ML) & 14,097 & 26.60 & 1.94 \\
      MCSA (Jacobi) & 90,129 & 1.0 & 1.0 \\
      \bottomrule
    \end{tabular}
  \end{table}
}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Path to Harder Problems}
  \begin{itemize}
    \item Experience with radiation diffusion suggests that for problems
      where spectral radius is low ($\lesssim 0.8$),
      MCSA can be competitive with leading methods
      \begin{itemize}
        \item Very low cost per iteration to approximate Neumann series
      \end{itemize}
    \vfill
    \item What about more challenging problems?
      \begin{itemize}
        \item Neumann series converges more slowly, so histories last longer
        \item More histories may be required
        \item What if Neumann series is not convergent?
      \end{itemize}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Polynomial Formulation

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Stopping Criteria -- Weight Cutoff}
  \begin{itemize}
    \item Histories are usually terminated when the relative weight drops
      below a specified cutoff (i.e. $\frac{w}{w_0} < \tau$)
    \vfill
    \item Adjoint MCSA on JPWH\_991 matrix with diagonal preconditioning
      ($\rho ( \mathbf{I-D^{-1}A} ) \approx 0.98$):
  \end{itemize}
\begin{figure}
  \centering
  \includegraphics[width=2.5in]{WtCutoff}
\end{figure}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Weight vs. History Length Cutoff}
\begin{figure}
  \centering
  \includegraphics[width=3in]{WtVsLength}
\end{figure}
\begin{itemize}
  \item Unlike weight cutoff, history length cutoff ``saturates'' at
    some number of histories -- fixed length Neumann series has been
    effectively reproduced
  \item Saturation point is higher for longer history lengths
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo as Approximate Polynomial Preconditioning}
  \begin{itemize}
    \item Using history length cutoff, Monte Carlo process is approximating
      a fixed length Neumann series polynomial
      \[
      x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
      \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}
      \]
    \vfill
    \item As number of histories grows, iteration count becomes identical to
      using ``true'' polynomial
    \vfill
    \item Why limit ourselves to the Neumann series polynomial?
      \begin{itemize}
        \item Chebyshev or GMRES polynomials are viable alternatives
      \end{itemize}
  \end{itemize}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Neumann Series Polynomial}
\begin{table}
\caption{Adjoint MCSA with Neumann Polynomial, $1000 \times 1000$ Shifted Laplacian Matrix.
Values are MCSA iteration counts (timing in milliseconds)
\label{tab:lap_adjoint_neumann}}
\centering
\begin{tabular}{cccc}
\toprule
Histories per & \multicolumn{3}{c}{Polynomial Order} \\
\cmidrule(lr){2-4}
Iteration & 2 & 4 & 6 \\
\midrule
250 &  775(100) & 651(96) & 664(110) \\
500 &  725(122) & 502(104) & 394(95) \\
1000 & 707(174) & 482(179) & 366(144) \\
2000 & 703(280) & 471(259) & 356(251) \\
4000 & 698(494) & 467(497) & 350(458) \\
8000 & 697(923) & 464(905) & 350(873) \\
16000 & 695(1796) & 462(1768) & 347(1711) \\
\bottomrule
\end{tabular}
\end{table}
\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{Chebyshev Polynomial}
\begin{table}
\caption{Adjoint MCSA with Chebyshev Polynomial, $1000 \times 1000$ Shifted Laplacian Matrix.
Values are MCSA iteration counts (timing in milliseconds)
\label{tab:lap_adjoint_cheby}}
\centering
\begin{tabular}{cccc}
\toprule
Histories per & \multicolumn{3}{c}{Polynomial Order} \\
\cmidrule(lr){2-4}
Iteration & 1 & 2 & 3 \\
\midrule
250 & - & - & - \\
500 & - & - & - \\
1000 & - & - & - \\
2000 & - & 328(134) & - \\
4000 & - & 296(210) & - \\
8000 & - & 288(380) & 262(423) \\
16000 & 1132(1866) & 283(721) & 175(550) \\
\bottomrule
\end{tabular}
\end{table}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Polynomial Methods -- Summary}
  \begin{itemize}
    \item Using history length cutoff rather than weight cutoff can lead to
      large reductions in iteration counts
      \vfill
    \item A good approximation to a few terms in the Neumann series
      performs better than a statistically noisy approximation to
      full series
      \vfill
    \item Significant reduction in iteration counts are possible using
      alternate polynomials, but generally outweighed by increase in number
      of histories required
      \begin{itemize}
        \item May be very beneficial from resiliency and parallel efficiency
          standpoints
        \item Will be re-evaluated in future efforts
      \end{itemize}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Parallelism

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Domain Decomposed Monte Carlo}

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Each parallel process owns a piece of the domain (linear
        system)
        \bigskip
      \item Random walks must be transported between adjacent domains
        through parallel communication
        \bigskip
      \item Domain decomposition determined by the input system
      \end{itemize}
    \end{column}

    \begin{column}{0.5\textwidth}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.75}{ \input{ddnu_example.pdftex_t} }
        \end{center}
        \caption{\small Domain decomposition example illustrating
          how domain-to-domain transport creates communication costs.}
      \end{figure}
    \end{column}
  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Asynchronous Monte Carlo Transport Kernel}

  \begin{columns}

    \begin{column}{0.5\textwidth}

      \begin{itemize}
      \item General extension of the Milagro algorithm (LANL)
      \item Asynchronous nearest neighbor communication of histories
      \item Binary asynchronous communication tree for completing
        transport
      \end{itemize}

      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{domain_to_domain.pdftex_t} }
        \end{center}
      \end{figure}

    \end{column}

    \begin{column}{0.5\textwidth}

      \begin{itemize}
      \item Extensible to problems where histories may be created
        (i.e. variance reduction)
      \end{itemize}

      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.45}{ \input{binary_comm_tree.pdftex_t} }
        \end{center}
      \end{figure}

    \end{column}

  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallel Application -- Nuclear Reactor Analysis}
  \small
  The simplified $P_N$ ($SP_N$) equations are an approximation
  to the Boltzmann neutron transport equation used to simulate nuclear
  reactors

  \vspace*{-0.1in}

  \begin{multline}
    \hat{\Omega} \cdot \vec{\nabla} \psi(\vec{r},\hat{\Omega},E) +
    \sigma(\vec{r},E) \psi(\vec{r},\hat{\Omega},E) = \\ \iint
    \sigma_s(\vec{r},E' \rightarrow E,\hat{\Omega}' \cdot \hat{\Omega})
    \psi(\vec{r},\hat{\Omega}',E') d\Omega' dE' +
    q(\vec{r},\hat{\Omega},E)
  \end{multline}

  \begin{multline}
    -\nabla \cdot \Bigg[\frac{n}{2n+1}\frac{1}{\Sigma_{n-1}} \nabla
      \Big(\frac{n-1}{2n-1} \phi_{n-2} + \frac{n}{2n-1}\phi_n \Big) \\+
      \frac{n+1}{2n+1}\frac{1}{\Sigma_{n+1}} \nabla
      \Big(\frac{n+1}{2n+3}\phi_n + \frac{n+2}{2n+3}\phi_{n+2}\Big)
      \Bigg] \\+ \Sigma_n \phi_n = q \delta_{n0}\ \ \ \ \ \ \ \ \ n =
    0,2,4,\cdots,N
  \end{multline}

  \begin{equation}
  \hspace*{-0.75in}
    -\nabla \cdot \mathbb{D}_n \nabla \mathbb{U}_n + \sum_{m=1}^4
    \mathbb{A}_{nm} \mathbb{U}_m = \frac{1}{k} \sum_{m=1}^4
    \mathbb{F}_{nm} \mathbb{U}_m\ \ \ \ \ \ \ n = 1,2,3,4
  \end{equation}

\end{frame}
%%---------------------------------------------------------------------------%%
\begin{frame}{$SP_N$ Assembly Problem}
\small
  \begin{columns}
    \begin{column}{0.4\textwidth}
      Test problem -- $3 \times 3$ array of fuel assemblies with
      control rod in center location (Profugus)
      \begin{itemize}
      \item 23 energy groups, 2 angular moments, 25M degrees of freedom
      \item 1,000 computational cores via domain decomposition
      \item We are interested in solving generalized eigenvalue problem,
        for this study we use Arnoldi as the eigensolver and compare
        different methods for solving linear systems
      \end{itemize}
    \end{column}
    \begin{column}{0.6\textwidth}
      \vspace*{-0.5in}
      \begin{figure}
        \centering
        \includegraphics[width=3in]{prob4}
      \end{figure}
    \end{column}
  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Parallel $SP_N$ Results}
\begin{table}
\centering
\begin{tabular}{cccc}
\toprule
%Method & Total GMRES Iterations & Setup Time (s) & Solve Time (s) \\
\multirow{2}{*}{\bfseries Method} &
\bfseries Total Linear & \bfseries Setup & \bfseries Solve \\
& \bfseries Iteration & \bfseries Time (s) & \bfseries Time (s) \\
\midrule
GMRES-ILUT      & 1675 & 0.7  & 18.4 \\
GMRES-AMG       & 626  & 0.7  & 46.0 \\
GMRES-MGE       & 498  & 1.5  & 33.7 \\
Richardson-AINV & 5208 & 20.6 & 52.0 \\
MCSA-AINV       & 1268 & 25.5 & 46.6 \\
\bottomrule
\end{tabular}
\end{table}
\begin{itemize}
  \item ILUT preconditioning is winner here, but known to have issues
    with parallel scaling on large core counts
  \item Solve times for MCSA are competitive, but setup times are very large
    due to construction of sparse approximate inverse
\end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Alternative Parallelism -- Additive Schwarz}
  \begin{itemize}
    \item Instead of performing Monte Carlo on full problem, another
      possibility is to apply Monte Carlo as an additive Schwarz approach
    \vfill
    \item Decompose problem into (possibly overlapping) domains
    \vfill
    \item Perform Monte Carlo on individual subdomains
      \begin{itemize}
        \item No communication costs in Monte Carlo problem!
      \end{itemize}
    \vfill
    \item With domain decomposed Monte Carlo, iteration counts are effectively
      independent of the number of processors
    \vfill
    \item In an additive Schwarz approach, the preconditioner will become
      less effective as processor counts grow -- algorithmic scalability
      may be an issue
      \vfill
    \item \textbf{Replication for resiliency and performance}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{More Parallelism -- Threading}
  \begin{itemize}
    \item Within a Monte Carlo solve, every history is independent of other
      histories -- great potential for highly concurrent hardware
      (GPU, Xeon Phi)
    \vfill
    \item Polynomial formulation enables a priori determination of
      operation counts per thread
      \vfill
    \item Memory locality an issue due to random access via random
      walks (block formulation?)
      \vfill
    \item Early experiments using the Trilinos Kokkos library show promising
      performance for multi-core CPUs
    \vfill
    \item Team members recently took part in OLCF ``Hackathon'' in
      late October to begin implementing computation kernels in
      OpenACC to allow for GPU capability on Titan with early results
      indicating \textbf{1.3-9.4x} speedup for MCSA (largely dependent
      on random number generation)
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}

  \center Conclusions

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Additional Thoughts}
  \begin{itemize}
    \item Our current implementations rely on performing fixed number of
      histories per Monte Carlo solve
      \begin{itemize}
        \item How can we dynamically select/adapt the number of histories
          that should be performed?  Use variance-based stopping criteria?
      \end{itemize}
    \vfill
    \item The ``almost optimal'' selection of the probability and weight
      matrices is arbitrary (as long as
      $\mathbf{H}/\mathbf{H}^T=\mathbf{P} \circ \mathbf{W}$)
      \begin{itemize}
        \item If $\mathbf{P}$ is taken to have a uniform distribution within
          each row, then sampling from the distribution can be done in
          constant (rather than logarithmic) time
      \end{itemize}
    \vfill
    \item Exploring Anderson acceleration of MCSA (Tim Kelley)
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Monte Carlo Linear Solvers Library (MCLS)}

  \begin{itemize}
  \item Designed to be easily incorporated with production physics
    codes
    \bigskip
  \item General asynchronous MSOD MCSA implementation
    \begin{itemize}
    \item Forward and adjoint Monte Carlo with method of expected
      values
    \item Parallel row matrix/vector interface
    \item General fixed point iteration strategy
    \item Explicit algebraic preconditioner suite
    \item Planned extensions to threaded/hybrid parallelism
    \end{itemize}
    \medskip
  \item Implemented in C++
    \bigskip
  \item Heavy use of the Trilinos scientific computing libraries
    \bigskip
  \item Open-source BSD 3-clause license
    \bigskip
  \item \textbf{\url{https://github.com/sslattery/MCLS}}
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Profugus Neutronics Mini-Application}

  A mini-application provides open-source kernels that effectively
  capture the algorithmic features of full applications

  \begin{columns}
    \begin{column}{0.3\textwidth}
      \vspace*{-0.4in}
      \begin{figure}
        \centering
        \includegraphics[width=1.5in]{exnihilo.png}
      \end{figure}
    \end{column}
    \begin{column}{0.7\textwidth}
      \begin{itemize}
      \item Extracts core computational kernels from the Exnihilo transport
        code suite
        \begin{itemize}
        \item Denovo: deterministic transport solvers
        \item Shift: Monte Carlo transport
        \item Insilico: reactor neutronics (cross sections, geometry,
          input/output)
        \end{itemize}
      \item Exnihilo is export-controlled, Profugus is not
      \end{itemize}
    \end{column}
  \end{columns}

  \vspace*{0.3in}
  Profugus includes:
  \begin{itemize}
  \item Fixed source and eigenvalue problems
  \item Trilinos solvers/preconditioners
  \item Generic material and cross section libraries
  \item \textbf{\url{https://github.com/ORNL-CEES/Profugus}}
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%
\begin{frame}{Conclusions}
  \begin{itemize}
  \item Monte Carlo methods offer great potential for both resilient and
    highly parallel solvers
    \vfill
  \item For certain classes of problems, Monte Carlo methods can be
    competitive with leading modern solvers
    \vfill
    \item Extending methods to broader problem areas is significant challenge
      and an attractive area for continued research
      \vfill
    \item Performance modeling and resiliency simulations this FY
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\end{document}
%%---------------------------------------------------------------------------%%

%%---------------------------------------------------------------------------%%
%% end of pres.tex
%%---------------------------------------------------------------------------%%