### 2016-01-25 doc/ecp/pres.tex
Updated first section draft.

%%---------------------------------------------------------------------------%%
%% Beamer ORNL Presentation Template
%%---------------------------------------------------------------------------%%

\documentclass{beamer}
%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm} \usepackage{amsmath} \usepackage{tmadd,tmath}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[compatibility=false]{caption}
\usepackage[labelformat=empty]{subcaption}
\usepackage{tikz}
\usepackage{ragged2e}

\usetikzlibrary{arrows}

%%---------------------------------------------------------------------------%%
%% THEME SETUP

\usetheme{CambridgeUS}
\usecolortheme{spruce}

%%---------------------------------------------------------------------------%%
%% SETUP STUFF
%%---------------------------------------------------------------------------%%

\logo{\includegraphics[width=2cm]{new_logo}}

\setbeamercolor{item}{fg=MSUgreen}
\beamertemplatenavigationsymbolsempty
\setbeamertemplate{headline}{}
\setbeamertemplate{footline}{
  \insertframenumber/\inserttotalframenumber}
\setbeamertemplate{title page}
{
  \begin{tabular}{cr}
    \begin{minipage}{5.5cm}
      {\RaggedRight \bf \large \textcolor{MSUgreen}{\inserttitle}}\\
      \vspace{\baselineskip}\\
      \insertauthor\\
      \vspace{0.5\baselineskip}\\
      \insertinstitute\\
      \insertdate
    \end{minipage}
  \end{tabular}
}

\newlength{\DUtablewidth} % internal use in tables

%%---------------------------------------------------------------------------%%
%% TITLE

\title[MCREX]{Using Monte Carlo Algorithms to Achieve Resiliency \& Performance
  at Scale for Linear and Nonlinear Solver Applications}
\author{Thomas Evans \and Steven Hamilton}
\date[DOE, 2016]{DOE RX-Solvers \\ January 29, 2016}
\institute[ORNL]{Oak Ridge National Lab}

%%---------------------------------------------------------------------------%%
\begin{document}
%%---------------------------------------------------------------------------%%

{
  \usebackgroundtemplate{
    \includegraphics[width=\paperwidth,height=\paperheight]{title_template}}
  \setbeamertemplate{logo}{}
  \begin{frame}
    \titlepage
  \end{frame}
}

%%---------------------------------------------------------------------------%%
\section*{Outline}
%%---------------------------------------------------------------------------%%

\begin{frame}
  \tableofcontents
\end{frame}

%%---------------------------------------------------------------------------%%
\section{Project Status}
%%---------------------------------------------------------------------------%%
\subsection{Project Goals and Target Applications}
%%---------------------------------------------------------------------------%%

\begin{frame}{Project Objectives}

  The MCREX project has 3 objectives for the development of Monte Carlo solver
  methods for linear systems:
  \begin{itemize}
  \item Analyze and improve the iterative performance of Monte Carlo
    algorithms for linear systems
  \item Analyze and characterize the resilient aspects of Monte Carlo
    algorithms for linear systems
  \item Develop and model highly scalable parallel algorithms for the
    implementation of Monte Carlo algorithms for linear systems
 \end{itemize}
 The end goal is to develop a set of resilient, scalable algorithms for the
 solution of linear systems.  We have made significant progress in each of
 these areas as summarized below.

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Target Applications}

  \begin{itemize}

  \item Models defined by PDEs that result in sparse linear systems
    \begin{itemize}
    \item fluid flow
    \item radiation transport
    \item electrostatics
    \end{itemize}

  \item We have begun to analyze eigenvalue systems that are important in
    network systems analysis and data mining

  \item Much of our analysis work has used and idealized
    advection-diffusion-reaction system that has properties common to these
    application areas

  \item Extension to non-linear Navier-Stokes dropped due to budget
    constraints of funded proposal
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%
\subsection{Overview and Status}
%%---------------------------------------------------------------------------%%

\begin{frame}{Monte Carlo for Linear Systems}
  \begin{itemize}
    \item Suppose we want to solve $\mathbf{Ax}=\mathbf{b}$
    \vfill
    \item If $\rho(\mathbf{I-A})<1$, we can write the solution using the
      Neumann series
      \begin{equation*}
        \mathbf{x} = \sum_{n=0}^{\infty} (\mathbf{I-A})^n \mathbf{b}
         = \sum_{n=0}^{\infty} \mathbf{H}^n \mathbf{b}
      \end{equation*}
      where $\mathbf{H} \equiv ( \mathbf{I-A} )$ is the Richardson
      iteration matrix
      \vfill
    \item Build the Neumann series stochastically
  \end{itemize}

  \[
  x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
  \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}
  \]

  \begin{itemize}
  \item Define a sequence of state transitions
  \end{itemize}
  \vspace*{-0.1in}
  \[
  \nu = i \rightarrow i_1 \rightarrow \cdots \rightarrow i_{k-1}
  \rightarrow i_{k}
  \]

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Solving the Heat Equation}

  \begin{figure}[h!]
    \begin{center}
      \includegraphics<1>[width=4in]{../NCState_seminar_2014/adjoint_1}
      \includegraphics<2>[width=4in]{../NCState_seminar_2014/adjoint_10}
      \includegraphics<3>[width=4in]{../NCState_seminar_2014/adjoint_100}
      \includegraphics<4>[width=4in]{../NCState_seminar_2014/adjoint_1000}
      \includegraphics<5>[width=4in]{../NCState_seminar_2014/adjoint_10000}
      \includegraphics<6>[width=4in]{../NCState_seminar_2014/adjoint_100000}
      \includegraphics<7>[width=4in]{../NCState_seminar_2014/adjoint_1000000}
      \includegraphics<8>[width=4in]{../NCState_seminar_2014/adjoint_10000000}
    \end{center}
    \caption{
      \only<1>{\textbf{Adjoint solution.}
        \textit{\sn{1}{0} total histories.} }
      \only<2>{\textbf{Adjoint solution.}
        \textit{\sn{1}{1} total histories.} }
      \only<3>{\textbf{Adjoint solution.}
        \textit{\sn{1}{2} total histories.} }
      \only<4>{\textbf{Adjoint solution.}
        \textit{\sn{1}{3} total histories.} }
      \only<5>{\textbf{Adjoint solution.}
        \textit{\sn{1}{4} total histories.} }
      \only<6>{\textbf{Adjoint solution.}
        \textit{\sn{1}{5} total histories.} }
      \only<7>{\textbf{Adjoint solution.}
        \textit{\sn{1}{6} total histories.} }
      \only<8>{\textbf{Adjoint solution.}
        \textit{\sn{1}{7} total histories.} }
    }
  \end{figure}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Limitations of Monte Carlo}
  \begin{itemize}
  \item Solving linear systems of equations with ``pure'' Monte Carlo methods
    is generally intractable
    \begin{itemize}
    \item Central limit theorem is barrier to accurate solutions
    \end{itemize}
    \begin{figure}
      \centering
      \includegraphics[width=1.75in]{../NCState_seminar_2014/Ifiss_ConvDiff}
    \end{figure}
  \item There is a potentially restrictive limit on the spectral properties of
    $\mathbf{H}$
    \begin{equation}
      \rho(\mathbf{H}) < 1\:, \rho(\hat{\mathbf{H}}) < 1\:, \hat{H}_{ij} =
      \frac{H_{ij}^2}{P_{ij}}
    \end{equation}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Monte Carlo Synthetic Acceleration}
  \begin{itemize}
  \item Address issue of convergence constrained by CLF
    \vfill
  \item Can be abstracted as a general linear solver
    \vfill
  \item Combine with Richardson iteration as a ``smoother'' in between
    Monte Carlo steps:
    \begin{align*}
      \mathbf{r}^k &= \mathbf{b} - \mathbf{Ax}^k \\
      \mathbf{x}^{k+1/2} &= \mathbf{x}^k + \mathbf{r}^k \\
      \mathbf{r}^{k+1/2} &= \mathbf{b} - \mathbf{Ax}^{k+1/2} \\
      \hat{\mathbf{A}} \mathbf{\delta} &= \mathbf{r}^{k+1/2} \\
      \mathbf{x}^{k+1} &= \mathbf{x}^{k+1/2} + \delta
    \end{align*}
  \item We still must address issue of convergence constraint by spectral radius
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Preconditioned MCSA}
\vspace*{-0.2in}
  {
    \small
    \[
    \ve{r}^{k} = \ve{M}_L^{-1}(\ve{b}-\ve{A}\ve{M}_R^{-1}\ve{x}^{k})
    \]
    \[
    \ve{x}^{k+1/2} = \ve{x}^k + \ve{r}^k
    \]
    \[
    \ve{r}^{k+1/2} = \ve{M}_L^{-1}(\ve{b}-\ve{A}\ve{M}_R^{-1}\ve{x}^{k+1/2})
    \]
    \[
    \ve{M}_L^{-1}\ve{A}\ve{M}_R^{-1}\delta\ve{x}^{k+1/2} = \ve{r}^{k+1/2}
    \]
    \[
    \ve{x}^{k+1} = \ve{x}^{k+1/2} + \delta \ve{x}^{k+1/2}
    \]
  }
  \vspace*{-0.1in}
  \begin{itemize}
    \item Requires explicit construction of preconditioned matrix
      \begin{itemize}
        \item This has generally limited preconditioner selection to diagonal,
          block diagonal, sparse approximate inverse approaches
      \end{itemize}
    \vfill
    \item Sparsity in $\mathbf{A}$, $\mathbf{M}$, does not imply sparsity
      in $\mathbf{M^{-1}A}$
    \vfill
    \item Need to investigate preconditioning strategies that lead to sparse
      preconditioned systems
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Monte Carlo as Approximate Polynomial Preconditioning}
  \begin{itemize}
    \item Using history length cutoff, Monte Carlo process is approximating
      a fixed length Neumann series polynomial
      \[
      x_i = \sum_{k=0}^{\infty}\sum_{i_1}^{N}\sum_{i_2}^{N}\ldots
      \sum_{i_k}^{N}h_{i,i_1}h_{i_1,i_2}\ldots h_{i_{k-1},i_k}b_{i_k}
      \]
    \vfill
    \item As number of histories grows, iteration count becomes identical to
      using ``true'' polynomial
    \vfill
    \item Why limit ourselves to the Neumann series polynomial?
      \begin{itemize}
        \item Chebyshev or GMRES polynomials are viable alternatives
      \end{itemize}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Neumann Series Polynomial}
  \begin{table}
    \caption{Adjoint MCSA with Neumann Polynomial, $1000 \times 1000$ Shifted Laplacian Matrix.
      Values are MCSA iteration counts (timing in milliseconds)
      \label{tab:lap_adjoint_neumann}}
    \centering
    \begin{tabular}{cccc}
      \toprule
      Histories per & \multicolumn{3}{c}{Polynomial Order} \\
      \cmidrule(lr){2-4}
      Iteration & 2 & 4 & 6 \\
      \midrule
      250 &  775(100) & 651(96) & 664(110) \\
      500 &  725(122) & 502(104) & 394(95) \\
      1000 & 707(174) & 482(179) & 366(144) \\
      2000 & 703(280) & 471(259) & 356(251) \\
      4000 & 698(494) & 467(497) & 350(458) \\
      8000 & 697(923) & 464(905) & 350(873) \\
      16000 & 695(1796) & 462(1768) & 347(1711) \\
      \bottomrule
    \end{tabular}
  \end{table}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Chebyshev Polynomial}
  \begin{table}
    \caption{Adjoint MCSA with Chebyshev Polynomial, $1000 \times 1000$ Shifted Laplacian Matrix.
      Values are MCSA iteration counts (timing in milliseconds)
      \label{tab:lap_adjoint_cheby}}
    \centering
    \begin{tabular}{cccc}
      \toprule
      Histories per & \multicolumn{3}{c}{Polynomial Order} \\
      \cmidrule(lr){2-4}
      Iteration & 1 & 2 & 3 \\
      \midrule
      250 & - & - & - \\
      500 & - & - & - \\
      1000 & - & - & - \\
      2000 & - & 328(134) & - \\
      4000 & - & 296(210) & - \\
      8000 & - & 288(380) & 262(423) \\
      16000 & 1132(1866) & 283(721) & 175(550) \\
      \bottomrule
    \end{tabular}
  \end{table}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Polynomial Methods -- Summary}
  \begin{itemize}
    \item Using history length cutoff rather than weight cutoff can lead to
      large reductions in iteration counts
      \vfill
    \item A good approximation to a few terms in the Neumann series
      performs better than a statistically noisy approximation to
      full series
      \vfill
    \item Significant reduction in iteration counts are possible using
      alternate polynomials, but partially offset by increase in number
      of histories required
      \begin{itemize}
        \item May be very beneficial from resiliency and parallel efficiency
          standpoints
        \item Will be re-evaluated in future efforts
      \end{itemize}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Preconditioning}
  \begin{itemize}
    \item Preconditioning for Monte Carlo introduces nonstandard considerations
      \begin{itemize}
        \item Elements of preconditioned matrix must be explicitly available
        \item Goal of preconditioning is to reduce $\rho(\hat{\mathbf{H}})$,
          may not follow ``standard'' trends
      \end{itemize}
  \end{itemize}
  \vfill

  \begin{table}[!h]
  \centering
  \begin{tabular}{|c|c|c|c|}
  \hline
  \textbf{Block size} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$
  &$\rho(\hat{H})$\\
  \hline
   3 & 1.4352 & $0.9804$ & $0.9531$\\
  \hline
   9 & 3.0220 & $0.9790$ & $0.9674$\\
  \hline
   33 & 9.8164 & $0.9783$  & $0.9774$\\
  \hline
  \end{tabular}
  \caption{$\rho(H)$ and $\rho(\hat{H})$ for advection--diffusion
  matrix with block diagonal preconditioning.}
  \label{tab:ifiss_block}
  \end{table}
\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Preconditioning}
  \begin{itemize}
    \item Several options have been analyzed
      \begin{itemize}
        \item (Block) diagonal preconditioners, suitable for (block)
          diagonally dominant systems
        \item ILU-style preconditioners are possible, but require additional
          filtering to maintain sparsity in preconditioned system
        \item Approximate inverse preconditioning is attractive, forming
          preconditioned matrix only requires sparse matrix-matrix product
      \end{itemize}
  \end{itemize}

  \begin{table}[!h]
  \centering
  \begin{tabular}{|c|c|c|c|}
  \hline
  \textbf{Drop tolerance} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$
  &$\rho(\hat{H})$\\
  \hline
   0.05 & 8.2797 & $0.9610$ & $1.0231$\\
  \hline
   0.01 & 33.1390 & $0.8668$ & $0.8279$\\
  \hline
  \end{tabular}
  \caption{$\rho(H)$ and $\rho(\hat{H})$ for 2D
  Laplacian with AINV preconditioning.}
  \label{tab:lap2d_ainv}
  \end{table}

%%---------------------------------------------------------------------------%%

\begin{frame}{Domain Decomposed Monte Carlo}

  \begin{columns}
    \begin{column}{0.5\textwidth}
      \begin{itemize}
      \item Each parallel process owns a piece of the domain (linear
        system)
        \bigskip
      \item Random walks must be transported between adjacent domains
        through parallel communication
        \bigskip
      \item Domain decomposition determined by the input system
        \bigskip
      \item Load balancing not yet addressed
      \end{itemize}
    \end{column}

    \begin{column}{0.5\textwidth}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.75}{ \input{ddnu_example.pdftex_t} }
        \end{center}
        \caption{\small Domain decomposition example illustrating
          how domain-to-domain transport creates communication costs.}
      \end{figure}
    \end{column}
  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Asynchronous Monte Carlo Transport Kernel}

  \begin{columns}

    \begin{column}{0.5\textwidth}
      \vspace{-0.1in}
      \small
      \begin{itemize}
      \item General extension of the Milagro algorithm (LANL)
      \item Asynchronous nearest neighbor communication of histories
      \item System-tunable communication parameters of buffer size and check
        frequency (performance impact)
      \item Need an asynchronous strategy for exiting the transport loop
        without a collective (running sum)
      \end{itemize}

      \vspace{-0.15in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.4}{ \input{domain_to_domain.pdftex_t} }
        \end{center}
      \end{figure}

    \end{column}

    \begin{column}{0.5\textwidth}

      \vspace{-0.25in}
      \begin{figure}[htpb!]
        \begin{center}
          \scalebox{0.825}{ \input{transport_loop.pdftex_t} }
        \end{center}
      \end{figure}

    \end{column}

  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Parallel Test - Simplified $P_N$ ($SP_N$) Assembly Problem}
\small
\begin{columns}

    \begin{column}{0.55\textwidth}

      \vspace{-0.2in}
      \begin{figure}
        \centering
        \includegraphics[width=2.0in]{../siam_cse_2015/presentation/prob4}
        \vspace{-0.2in}
        \caption{$SP_N$ solution example}
      \end{figure}

      \vspace{-0.1in}

      The ($SP_N$) equations are an approximation to the Boltzmann neutron
      transport equation used to simulate nuclear reactors

      \vspace{-0.2in}

      \[
        \small
        -\nabla \cdot \mathbb{D}_n \nabla \mathbb{U}_n + \sum_{m=1}^4
        \mathbb{A}_{nm} \mathbb{U}_m = \frac{1}{k} \sum_{m=1}^4
        \mathbb{F}_{nm} \mathbb{U}_m
      \]

    \end{column}

    \begin{column}{0.45\textwidth}
      Scaling problem -- $1 \times 1$ to $17 \times 17$ array of fuel
      assemblies with 289 pins each resolved by a $2 \times 2$ spatial mesh
      and 200 axial zones

      \vfill

      \begin{itemize}
      \item 7 energy groups, 1 angular moment, 1.6M to 273.5M degrees of
        freedom
      \item 64 to 10,800 computational cores via domain decomposition
      \item We are usually interested in solving generalized eigenvalue
        problem - we use the operator from these problems to test the kernel
        scaling
      \end{itemize}
    \end{column}

  \end{columns}

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Domain Decomposition Scaling}

  \vspace{-0.2in}

  \begin{table}[htb!]
    \tiny
    \begin{center}
      \begin{tabular}{rrrrrrr}
        \toprule
        \multicolumn{1}{r}{Cores} &
        \multicolumn{1}{r}{DOFs} &
        \multicolumn{1}{r}{DOFs/Core} &
        \multicolumn{1}{r}{Time Min (s)} &
        \multicolumn{1}{r}{Time Max (s)} &
        \multicolumn{1}{r}{Time Ave (s)} &
        \multicolumn{1}{r}{Efficiency}
        \\ \midrule
        %%
        256 & 273\,509\,600 & 1\,068\,397 & 515.51 & 515.52 & 515.52 & 1.00 \\
        1\,024 & 273\,509\,600 & 267\,099 & 122.76 & 122.77 & 122.76 & 1.05 \\ %CACHE
        4\,096 & 273\,509\,600 & 66\,775 & 27.96 & 27.97 & 27.96 & 1.15 \\
        7\,744 & 273\,509\,600 & 35\,319 & 17.72 & 17.72 & 17.72 & 0.96 \\
        10\,816 & 273\,509\,600 & 25\,288 & 13.72 & 13.72 & 13.72 & 0.89 \\
        %%
        \bottomrule
      \end{tabular}
    \end{center}
    \vspace{-0.07in}
    \caption{\small Strong Scaling}
  \end{table}

  \vspace{-0.4in}

  \begin{table}[htb!]
    \tiny
    \begin{center}
      \begin{tabular}{rrrrrrr}
        \toprule
        \multicolumn{1}{r}{Cores} &
        \multicolumn{1}{r}{DOFs} &
        \multicolumn{1}{r}{DOFs/Core} &
        \multicolumn{1}{r}{Time Min (s)} &
        \multicolumn{1}{r}{Time Max (s)} &
        \multicolumn{1}{r}{Time Ave (s)} &
        \multicolumn{1}{r}{Efficiency}
        \\ \midrule
        %%
        64 & 1\,618\,400 & 25\,288 & 12.65 & 12.65 & 12.65 & 1.00 \\
        256 & 6\,473\,600 & 25\,288 & 12.86 & 12.86 & 12.86 & 0.98 \\
        1\,024 & 25\,894\,400 & 25\,288 & 14.59 & 14.59 & 14.59 & 0.87 \\
        4\,096 & 103\,577\,600 & 25\,288 & 14.25 & 14.25 & 14.25 & 0.89 \\
        7\,744 & 195\,826\,400 & 25\,288 & 14.75 & 14.78 & 14.75 & 0.86 \\
        10\,816 & 273\,509\,600 & 25\,288 & 13.72 & 13.72 & 13.72 & 0.92 \\
        %
        \bottomrule
      \end{tabular}
    \end{center}
        \vspace{-0.07in}
    \caption{\small Weak Scaling}
  \end{table}

  \vspace{-0.25in}

  \begin{itemize}
    \small
  \item Titan Cray XK7 was used with no GPUs - used Eos parameters
  \item Results are consistent with transport literature
  \item History length of 15 and 3 histories per DOF
  \item Consider these effective operator apply times of 15 realizations of
    $\ve{H}$
  \item MC dominates run times in solver applications
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Replication}

  \vspace{-0.1in}

  Different batches of Monte Carlo samples can be combined in summation via
  superposition if they have different random number streams.  Consider each
  of these batches independent \textit{subsets} of a Monte Carlo operator
  where now the operator can be formed as a general additive decomposition of
  $N_S$ subsets:

  \[
  \mathbf{M_{MC}} = \frac{1}{N_S}\sum_{n=1}^{N_s}{\mathbf{M_n}}
  \]

  We replicate the linear problem and form each subset on a different group of
  parallel processes. Applying the subsets to a vector requires an AllReduce
  to form the sum. Each subset is domain decomposed.
  \vfill
  This provides a mechanism to support \textit{resiliency}; the replicated
  component provides {\bf useful} work, not just redundancy

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{Replication Scaling}

  \begin{table}[htb!]
    \tiny
    \begin{center}
      \begin{tabular}{rrrrrrr}
        \toprule
        \multicolumn{1}{r}{Subsets} &
        \multicolumn{1}{r}{Cores} &
        \multicolumn{1}{r}{DOFs} &
        \multicolumn{1}{r}{Time Min (s)} &
        \multicolumn{1}{r}{Time Max (s)} &
        \multicolumn{1}{r}{Time Ave (s)} &
        \multicolumn{1}{r}{Efficiency}
        \\ \midrule
        %%
        1 & 256 & 6\,473\,600 & 12.65 & 12.65 & 12.65 & 1.00 \\
        2 & 512 & 6\,473\,600 & 6.62 & 6.80 & 6.72 & 0.93 \\
        3 & 768 & 6\,473\,600 & 4.50 & 4.73 & 4.66 & 0.89 \\
        4 & 1\,024 & 6\,473\,600 & 3.62 & 3.81 & 3.71 & 0.83 \\
        %
        \bottomrule
      \end{tabular}
    \end{center}
    \caption{\small Fixed algorithmic scaling}
  \end{table}

  \vfill

  \begin{itemize}
    \small
  \item Timings are Monte Carlo time plus AllReduce time
  \item 256 cores per subset with domain decomposition
  \item $1/N_S$ samples per subset gives flat algorithmic performance
  \item Two performance issues
    \begin{itemize}
    \item Dividing up samples creates work starvation
    \item AllReduce buffer size is constant regardless of work starvation
    \end{itemize}
  \item Research is needed for a more performant subset combination
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%

\begin{frame}{GPU Optimization}

  We have implemented a version of MCSA that uses GPUs to perform the MC
  linear solve.

  Distributed parallelism is enabled through an additive Schwartz method, all
  MC solves
  \begin{itemize}
  \item All MC random walks are local
  \item No inter-domain communication of histories
  \end{itemize}

  \begin{table}[h]
    \begin{center}
      \caption{2D parabolic problem with AINV preconditioning.}
      \begin{tabular}{|c|c|c|c|}\hline
        Node & Iterations & Time (sec) & Relative Residual \\\hline
        CPU & 6 & 992.23 & $1.23\times 10^{-7}$ \\\hline
        GPU & 6 & 29.83 & $1.53\times 10^{-7}$ \\\hline
      \end{tabular}
    \end{center}
  \end{table}

\end{frame}

%%---------------------------------------------------------------------------%%
\subsection{Research Requirements}
%%---------------------------------------------------------------------------%%

\begin{frame}

  The following important questions need to be answered before MC methods can
  be applied to linear systems as a ``standard'' solver:

  \begin{itemize}
  \item Automatically determine history termination criteria
    \begin{itemize}
    \item We have investigated adaptive schemes based on variance
    \item However, variance is not directly correlated with error
    \end{itemize}
  \item Identify classes of matrices and preconditioners that satisfy
    convergence constraints
    \begin{itemize}
    \item This has been done for SDD and GDD
    \end{itemize}
  \item Demonstrate resiliency to hard MPI errors through replicated domain
    parallelism
  \item Further investigate non-linear, time-dependent problems
    \begin{itemize}
    \item This was part of the original proposal
    \item Non-linear, sub-cycle time-stepping can result in strongly diagonally
      dominant problems for which MCSA has been shown to be effective
    \end{itemize}
  \end{itemize}
\end{frame}

%%---------------------------------------------------------------------------%%
\subsection{Software}
%%---------------------------------------------------------------------------%%

\begin{frame}{MCSA Implementations}

  {\bf MCLS} (\underline{github.com/sslattery/MCLS})
  \begin{itemize}
  \item Massively parallel (MPI) implementation of domain decomposed MCSA
    algorithm
  \item Implemented in Trilinos framework and callable through Trilinos'
    abstract numerical algorithms (Thyra, Stratimikos) interface
  \item Works with any Tpetra row matrix
  \end{itemize}
  \vfill
  {\bf Alea} (\underline{github.com/ORNL-CEES/Profugus})
  \begin{itemize}
  \item GPU-accelerated on-node MC solve
  \item Domain-local MC solves implemented for distributed systems as an
    additive Schwartz method
  \end{itemize}

\end{frame}

%%---------------------------------------------------------------------------%%
\section{Transition and Roadmap}
%%---------------------------------------------------------------------------%%

%%---------------------------------------------------------------------------%%

\begin{frame}
\vfill
\centering
\huge Questions?
\vfill
\end{frame}

%%---------------------------------------------------------------------------%%
\end{document}
%%---------------------------------------------------------------------------%%

%%---------------------------------------------------------------------------%%
%% end of pres.tex
%%---------------------------------------------------------------------------%%